### 2019-06-22 article/article-tesis.tex
adding article

\documentclass[twocolumn]{article}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{inputenc}
\usepackage{graphicx}
\usepackage{authblk}

\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

\begin{document}
\author[1]{Eric Raphael Huiza Pereyra}
\affil[1]{Department of Postgraduate Studies, Pontifical Catholic University of Peru}
\affil[]{\textit{eric.huiza@pucp.edu.pe}}

\title{\vspace{-2.0cm}\textbf{Method to detect gestures in a poorly annotated dataset}}

\maketitle

\begin{abstract}
Include abstract here \\
\keywords{video classification,gestures detection}
\end{abstract}

\section{Introduction}

The World Health Organization (WHO) stated that 466 million people world wide have disabling hearing loss and it is estimated that by 2050 over 900 million people will have disabling hearing loss and represents a global cost of 750.00 million dollars annually  \cite{deafness_and_hearing_loss_2019}.

The Peruvian Institute of Informatics and Statistics (INEI) executed a national disabilities survey with the objective to better understand which disabilities affect the Peruvian population

En el Perú, según la última encuesta nacional especializada por discapacidad del instituto peruano de Estadística e Informática INEI realizada durante el año 2012 y publicada en marzo del 2014, el 1.8\% de la población presenta limitaciones auditivas permanentes.

El día de hoy ya es completamente viable pensar en sistemas que puedan detectar y transcribir lenguajes de señas. Sin embargo los lenguajes de señas varían mucho en diferentes regiones, y es muy costoso preparar un corpus que permita entrenar los modelos de inteligencia artificial. El Grupo de Investigación de Señas Gramaticales de la Pontificia Universidad Católica del Perú (PUCP) ha hecho el esfuerzo de crear este corpus para el Lenguaje de Señas Peruano (LSP), el cual está débilmente anotado, es decir, no se tiene anotado de manera exacta el instante en que se emite una seña y su traducción.

Este estudio se ha realizado para la obtención del grado de Magister en Informática con mención en Ciencias de la Computación durante el periodo 2019 y trata de responder las siguientes preguntas:
\begin{itemize}
\item ¿Cuales son las técnicas disponibles más relevantes para el entrenamiento de un modelo de aprendizaje de máquina a partir de contenido visual y transcripción de audio y video débilmente anotado?
\item ¿Qué tan preciso y exhaustivo es el modelo descrito en la pregunta anterior en el reconocimiento de un conjunto reducido de señas a partir de un conjunto de datos débilmente anotado?
\item ¿Cual es la relación entre el número de muestras y la detección de nuevos elementos del Lenguaje de Señas Peruano en cuanto al rendimiento en cuanto a precisión y exhaustividad del método?
\end{itemize}
contribuyendo en el desarrollo de un método generalizable para la detección progresiva de señas que hará uso del corpus del LSP elaborado por la PUCP. Asimismo los resultados del estudio servirán como base a futuras investigaciones como la mejora de la accesibilidad a sistemas informáticos y la interacción humano computador para personas con discapacidades auditivas o de habla.
\section{Related Work}
\section{Method}
\section{Experimentation}
\subsection{Dataset Description}
Este estudio hace uso del corpus de la LSP elaborado por el Grupo de Investigación en Señas Gramaticales de la PUCP el cual está limitado a 24 informantes, 12 hombres y 12 mujeres, todos residentes en Lima. Todos reportaron ser sordos de nacimiento o de manera pre-locutiva, es decir, antes de la adquisición del Castellano. El corpus esta conformado por total de 718 clips de video traducidos al Castellano, los cuales se encuentran disponibles en el Archivo Digital de Lenguas Peruanas de la PUCP (http://repositorio.pucp.edu.pe/index/handle/123456789/15344)

El corpus se encuentra débilmente anotado ya que no existe una relación directa entre la traducción y el momento en que se emite la seña.

\subsection{Video Annotation}
\begin{figure}[ht!]
\includegraphics[width=\linewidth]{video-annotation-pipeline.png}
\caption{Video annotation process}
\end{figure}
\section{Conclusion}

\bibliographystyle{ieeetr}
\bibliography{References}

\end{document}

### 2019-06-24 article/article-tesis.tex
adding introduction and data set description to the paper

@@
 \usepackage{authblk}
+\usepackage{enumerate}
 
@@
 
-\title{\vspace{-2.0cm}\textbf{Method to detect gestures in a poorly annotated dataset}}
+\title{%
+	\vspace{-2.0cm}
+	\textbf{Talking with body expressions} \\	
+	\Large \textbf{A Method for detecting sings in a weakly annotated dataset}
+}
 
@@
 
-\section{Introduction}
+\section{Introduction}\label{intro}
 
@@
 
-The Peruvian Institute of Informatics and Statistics (INEI) executed a national disabilities survey with the objective to better understand which disabilities affect the Peruvian population
+The Peruvian Institute of Informatics and Statistics (INEI) executed a national disabilities survey with the objective to segment and better understand which disabilities affect the Peruvian population \cite{disabilities_survey_2012}, the survey results show that 1.8\% of the Peruvian population shows at least one partial or permanent deafness or hearing limitation.
 
-En el Perú, según la última encuesta nacional especializada por discapacidad del instituto peruano de Estadística e Informática INEI realizada durante el año 2012 y publicada en marzo del 2014, el 1.8\% de la población presenta limitaciones auditivas permanentes.
+As of today is completely viable to think on systems that are able to detect and transcribe signs languages.
+However, in the same way as spoken languages, signs languages present local variations, for example, people that lives in the Lima metro area are not expect to use the same set of signs as people leaving in different parts of the country.
+Developing a dataset that describes the Peruvian Signs Language (LSP) \cite{lsp_2015} that can be used to train an AI model is expensive.
 
-El día de hoy ya es completamente viable pensar en sistemas que puedan detectar y transcribir lenguajes de señas.
-Sin embargo los lenguajes de señas varían mucho en diferentes regiones, y es muy costoso preparar un corpus que permita entrenar los modelos de inteligencia artificial.
-El Grupo de Investigación de Señas Gramaticales de la Pontificia Universidad Católica del Perú (PUCP) ha hecho el esfuerzo de crear este corpus para el Lenguaje de Señas Peruano (LSP), el cual está débilmente anotado, es decir, no se tiene anotado de manera exacta el instante en que se emite una seña y su traducción.
+The Pontifical Catholic University of Peru (PUCP) Grammar and  Signs research group developed a LSP dataset \cite{lsp_dataset} that is weekly annotated because it does not show exactly the relation between the instant when a sign that is emitted and its corresponding translation.
 
-Este estudio se ha realizado para la obtención del grado de Magister en Informática con mención en Ciencias de la Computación durante el periodo 2019 y trata de responder las siguientes preguntas:
-\begin{itemize}
-\item ¿Cuales son las técnicas disponibles más relevantes para el entrenamiento de un modelo de aprendizaje de máquina a partir de contenido visual y transcripción de audio y video débilmente anotado?
-\item ¿Qué tan preciso y exhaustivo es el modelo descrito en la pregunta anterior en el reconocimiento de un conjunto reducido de señas a partir de un conjunto de datos débilmente anotado?
-\item ¿Cual es la relación entre el número de muestras y la detección de nuevos elementos del Lenguaje de Señas Peruano en cuanto al rendimiento en cuanto a precisión y exhaustividad del método?
-\end{itemize}
-contribuyendo en el desarrollo de un método generalizable para la detección progresiva de señas que hará uso del corpus del LSP elaborado por la PUCP.
-Asimismo los resultados del estudio servirán como base a futuras investigaciones como la mejora de la accesibilidad a sistemas informáticos y la interacción humano computador para personas con discapacidades auditivas o de habla.
-\section{Related Work}
-\section{Method}
-\section{Experimentation}
-\subsection{Dataset Description}
-Este estudio hace uso del corpus de la LSP elaborado por el Grupo de Investigación en Señas Gramaticales de la PUCP el cual está limitado a 24 informantes, 12 hombres y 12 mujeres, todos residentes en Lima.
-Todos reportaron ser sordos de nacimiento o de manera pre-locutiva, es decir, antes de la adquisición del Castellano.
-El corpus esta conformado por total de 718 clips de video traducidos al Castellano, los cuales se encuentran disponibles en el Archivo Digital de Lenguas Peruanas de la PUCP (http://repositorio.pucp.edu.pe/index/handle/123456789/15344)
+This study was developed to obtain the Masters Degree in Informatics and Computer Science and propose a novel Signs Detection Method (nSDm) that allows signs detection in a weakly annotated dataset trying to answer the following questions:
 
-El corpus se encuentra débilmente anotado ya que no existe una relación directa entre la traducción y el momento en que se emite la seña.
+\begin{enumerate}[(i)]
+\item What are most relevant techniques currently available for training an AI model from weakly annotated video and audio transcription?\label{q1}
+\item How precise and exhaustive is the model described in the above question on the recognition of a reduced set of signs in a weakly annotated dataset?\label{q2}
+\item What is the relation between the number of samples and the detection of new LSP elements in terms of nSDm precision and recall?\label{q3}
+\end{enumerate}
 
-\subsection{Video Annotation}
+nSDm will take advantage of the LSP dataset \cite{lsp_dataset} and will contribute in the generalization and progressive signs detection and could be used as a starting point for future studies including software accessibility improvements and human computer interaction for people with deafness and hearing limitations.
+
+The rest of the article is organized as follows.
+In sec \ref{relatedwork} we review the related work on video classification and body actions recognition using both Optical Flow+Two Stream CNN and 3D CNN architectures.
+In sec \ref{method} we introduce nSDm and specify its architecture and design.
+In sec \ref{experimentation} we evaluate nSDm precision and recall and provide answers for the research questions \ref{q1},\ref{q2},\ref{q3}.
+In sec \ref{datasetdesc} we describe and provide the dataset details.
+In sec \ref{videoannot} we describe the video annotation and data pre-processing techniques applied to the dataset and finally in sec \ref{conclusion} we present our conclusion and describe future work.
+\section{Related Work} \label{relatedwork}
+\section{Method}\label{method}
+\section{Experimentation}\label{experimentation}
+\subsection{Dataset Description \cite{lsp_dataset}}\label{datasetdesc}
+The LSP dataset was developed by the PUCP Grammar and Signs research Group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female (informants), all of them residents in Lima Peru.
+All of them reported to be borne with a permanent deafness condition or acquired the condition before the acquisition of Spanish.
+The dataset consists in 718 video clips recorded with a ADR-CX220 SONY HD camera which included an embedded microphone.
+The camera only focus the informant but also records questions, instructions and translations.
+
+The videos were recorded in three sessions that included three participants: The coordinator, the translator and the informant.
+
+\textbf{Recording Session 1}: A 45-60 minutes semi structured interview that included: Biographic information as well as habits, anecdotes, opinion about cultural subjects and elicitation of names, states and actions.
+
+\textbf{Recording Session 2}: The informant was presented with a set of 55 cards describing actions and were asked to choose a set of them and build a coherent story that was subsequently told by the informant.
+
+\textbf{Recording Session 3}: A LSP conversation between the informant and the translator. The conversation was facilitated by the coordinator
+
+During all the sessions a LSP translator performs a translation after a word or phrase is finished.
+The dataset is weakly annotated given there is not direct relation between the translation and the instant when the sign is emitted.
+
+\subsection{Video Annotation}\label{videoannot}
 \begin{figure}[ht!]
@@
 \end{figure}
-\section{Conclusion}
+\section{Conclusion}\label{conclusion}
 

### 2019-06-29 article/article-tesis.tex
translating the video annotation process, added state of art for actions recognition, first version of the abstract and some minior fixes

@@
 \begin{abstract}
-Include abstract here \\
-\keywords{video classification,gestures detection}
+People with deafness or hearing disabilities are aimed to use computer based systems, state of art video classification and human action recognition rely in a combination of traditional and new methods like optical flow and 3D CNN.
+In this article we show a method to detect Peruvian Sign Language elements from a weakly annotated videos dataset.
+Video annotation techniques were applied increasing the models performance, as a result we achieved XX precision and XX exhaustiveness by training a 3D CNN with optical flow inputs.
+\\
+\keywords{video classification,human actions detection, optical flow, 3D CNN.}
 \end{abstract}
@@
 
-The World Health Organization (WHO) stated that 466 million people world wide have disabling hearing loss and it is estimated that by 2050 over 900 million people will have disabling hearing loss and represents a global cost of 750.00 million dollars annually \cite{deafness_and_hearing_loss_2019}.
+The World Health Organization (WHO) stated that 466 million people world wide have disabling hearing loss, estimating that by 2050 over 900 million people will have disabling hearing loss and will represent a global cost of 750.00 million dollars annually \cite{deafness_and_hearing_loss_2019}.
 
-The Peruvian Institute of Informatics and Statistics (INEI) executed a national disabilities survey with the objective to segment and better understand which disabilities affect the Peruvian population \cite{disabilities_survey_2012}, the survey results show that 1.8\% of the Peruvian population shows at least one partial or permanent deafness or hearing limitation.
+The Peruvian Institute of Informatics and Statistics (INEI) executed a national disabilities survey with the objective to segment and better understand which disabilities affect the Peruvian population \cite{disabilities_survey_2012}, the results showed that 1.8\% of the Peruvian population present at least a partial or permanent deafness or hearing limitation.
 
-As of today is completely viable to think on systems that are able to detect and transcribe signs languages.
-However, in the same way as spoken languages, signs languages present local variations, for example, people that lives in the Lima metro area are not expect to use the same set of signs as people leaving in different parts of the country.
-Developing a dataset that describes the Peruvian Signs Language (LSP) \cite{lsp_2015} that can be used to train an AI model is expensive.
+As of today it is completely viable to think about systems able to detect and transcribe signs languages.
+However, in the same way as spoken languages, signs languages also present local variations e.g.
+people that lives in the Lima metro area are not expect to use the same set of signs as people leaving in different parts of the country.
+Developing a dataset that describes the Peruvian Sign Language (PRL) \cite{lsp_2015} that can be used to train an AI model is expensive and difficult to achieve.
 
-The Pontifical Catholic University of Peru (PUCP) Grammar and  Signs research group developed a LSP dataset \cite{lsp_dataset} that is weekly annotated because it does not show exactly the relation between the instant when a sign that is emitted and its corresponding translation.
+The Pontifical Catholic University of Peru (PUCP) Grammar and  Signs research group created a PRL dataset \cite{lsp_dataset} that is weekly annotated because it does not show exactly the relation between the instant when a sign that is emitted and its corresponding translation.
 
-This study was developed to obtain the Masters Degree in Informatics and Computer Science and propose a novel Signs Detection Method (nSDm) that allows signs detection in a weakly annotated dataset trying to answer the following questions:
+This study was developed to obtain the Masters Degree in Informatics and Computer Science and proposes a novel Signs Detection Method (nSDm) that allows signs detection in a weakly annotated dataset and tries to answer the following questions:
 
 \begin{enumerate}[(i)]
-\item What are most relevant techniques currently available for training an AI model from weakly annotated video and audio transcription?\label{q1}
+\item What are most relevant currently available techniques for training an AI model from  weakly annotated video and audio transcription?\label{q1}
 \item How precise and exhaustive is the model described in the above question on the recognition of a reduced set of signs in a weakly annotated dataset?\label{q2}
@@
 
-nSDm will take advantage of the LSP dataset \cite{lsp_dataset} and will contribute in the generalization and progressive signs detection and could be used as a starting point for future studies including software accessibility improvements and human computer interaction for people with deafness and hearing limitations.
+nSDm will take advantage of the PRL dataset \cite{lsp_dataset} and will contribute with the generalization and progressive signs detection that could be used as a starting point for future studies including software accessibility improvements and human computer interaction for people with deafness and hearing limitations.
 
@@
 \section{Related Work} \label{relatedwork}
+\subsection{Action Recognition}
+Human action recognition is a extensively studied field and available action recognition dataset: UCF101, HMDB51, THUMOS14  are available, researches tried to solve the human action recognition problem using different approaches including Optical Flow and 3D CNN \cite{qiu2017learning}.
+\textbf{Optical Flow,} is defined as the pattern obtained from the motion of objects, surfaces and edges in a visual scene caused by the relative motion between the observer and a scene, it is achieved by distributing movement velocities and brightness across frames.
+It is a  key concept in action recognition from videos \cite{wang2019hallucinating}.
+Optical flow estimation is treated as an image reconstruction problem.
+Given a frame set, the optical flow is generated and allows to reconstruct one frame from the others \cite{zhu2018hidden}.
+Formally, taking the optical flow displacement field as input and training a CNN with it, then the network should have learned useful representations of the underlying motions.
+Even though Optical Flow represents the movement between a set of frames, if camera motion is considered as an action motion, it may corrupt the action classification \cite{wang2013dense}.
+Various types of camera motion can be observed in realistic videos, e.g., zooming, tilting, rotation, etc.
+\textbf{Motion Boundary Histogram (MBH),} It is a simple an efficient way to achieve robustness during human action detection when camera movements are mixed within the recorded actions by computing derivatives separately for the horizontal and vertical components of the optical flow.
+Since MBH represents the gradient of optical flow, locally constant camera motion is removed and information about changes in the flow field is kept.
+MBH is more robust to camera motion than optical flow, thus more discriminative for action recognition.\cite{wang2013dense}.
+3D CNN are not as effective as optical flow to detect human actions on its own, 3D CNN can be trained to learn optical flow so we can avoid costly computation and storage and obtain task-specific motion representation  \cite{zhu2018hidden} and increase models performance, precision and exhaustiveness on human action recognition.
+\subsection{Video Classification}
+\textbf{Bag of Words.}
+\\
+\textbf{Fussing parallel CNN.}
 \section{Method}\label{method}
@@
 \subsection{Dataset Description \cite{lsp_dataset}}\label{datasetdesc}
-The LSP dataset was developed by the PUCP Grammar and Signs research Group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female (informants), all of them residents in Lima Peru.
-All of them reported to be borne with a permanent deafness condition or acquired the condition before the acquisition of Spanish.
+The LSP dataset was developed by the PUCP Grammar and Signs research group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female informants, all of them are Lima Peru residents and reported to be borne with a permanent deafness condition or acquired the condition before the acquisition of Spanish.
+
 The dataset consists in 718 video clips recorded with a ADR-CX220 SONY HD camera which included an embedded microphone.
-The camera only focus the informant but also records questions, instructions and translations.
+The camera focused only the informant but also recorded questions, instructions and translations.
 
-The videos were recorded in three sessions that included three participants: The coordinator, the translator and the informant.
+The video clips were recorded in three sessions with the following participants: A coordinator, a LSP \cite{lsp_2015} translator and a informant.\\
 
@@
 
-\textbf{Recording Session 2}: The informant was presented with a set of 55 cards describing actions and were asked to choose a set of them and build a coherent story that was subsequently told by the informant.
+\textbf{Recording Session 2}: The informant was presented with a set of 55 cards describing actions and were asked to choose a set of them in order to build a coherent story that was subsequently told by the informant.
 
-\textbf{Recording Session 3}: A LSP conversation between the informant and the translator. The conversation was facilitated by the coordinator
+\textbf{Recording Session 3}: A LSP \cite{lsp_2015} conversation facilitated by the coordinator happening between the informant and the translator.\\
 
-During all the sessions a LSP translator performs a translation after a word or phrase is finished.
+During all the sessions a LSP \cite{lsp_2015} translator performs a translation after a word or phrase is completed.
 The dataset is weakly annotated given there is not direct relation between the translation and the instant when the sign is emitted.

### 2019-06-29 article/article-tesis.tex
abstract

@@
 \begin{abstract}
-People with deafness or hearing disabilities are aimed to use computer based systems, state of art video classification and human action recognition rely in a combination of traditional and new methods like optical flow and 3D CNN.
+People with deafness or hearing disabilities who aim to use computer based systems, state of art video classification and human action recognition rely in a combination of traditional and new methods like optical flow and 3D CNN.
 In this article we show a method to detect Peruvian Sign Language elements from a weakly annotated videos dataset.

### 2019-06-30 article/article-tesis.tex
introduction

@@
 	\textbf{Talking with body expressions} \\	
-	\Large \textbf{A Method for detecting sings in a weakly annotated dataset}
+	\Large \textbf{A method for detecting signs in a weakly annotated dataset}
 }
@@
 
-The World Health Organization (WHO) stated that 466 million people world wide have disabling hearing loss, estimating that by 2050 over 900 million people will have disabling hearing loss and will represent a global cost of 750.00 million dollars annually \cite{deafness_and_hearing_loss_2019}.
+The World Health Organization (WHO) stated that 466 million people world wide have disabling hearing loss, estimating that by 2050 over 900 million people will have disabling hearing loss and will represent a global cost of 750 million dollars annually \cite{deafness_and_hearing_loss_2019}.
 
-The Peruvian Institute of Informatics and Statistics (INEI) executed a national disabilities survey with the objective to segment and better understand which disabilities affect the Peruvian population \cite{disabilities_survey_2012}, the results showed that 1.8\% of the Peruvian population present at least a partial or permanent deafness or hearing limitation.
+The Peruvian Institute of Informatics and Statistics (INEI) conducted a national disabilities survey with the objective of segmenting and better understanding which disabilities affect the Peruvian population \cite{disabilities_survey_2012}, the results showed that 1.8\% of the Peruvian population suffer at least partial when not permanent deafness or hearing limitation.
 
-As of today it is completely viable to think about systems able to detect and transcribe signs languages.
-However, in the same way as spoken languages, signs languages also present local variations e.g.
-people that lives in the Lima metro area are not expect to use the same set of signs as people leaving in different parts of the country.
-Developing a dataset that describes the Peruvian Sign Language (PRL) \cite{lsp_2015} that can be used to train an AI model is expensive and difficult to achieve.
+As of today, thanks to deep learning models for computer vision and natural language processing, it is completely feasible to propose systems able to detect and transcribe sign languages.
+However, deep learning supervised models usually require large and well annotated datasets, and developing such a dataset for the Peruvian Sign Language (PSL) \cite{lsp_2015} is expensive and difficult to achieve.
+Furthermore, in the same way as spoken languages, signs languages also present local variations e.g.
+people who live in Lima metropolitan area are not expected to use the same set of signs as people leaving in other parts of the country.
+A need exists for approaching this task with self-supervised learning or by means of weakly labeled data.
 
-The Pontifical Catholic University of Peru (PUCP) Grammar and  Signs research group created a PRL dataset \cite{lsp_dataset} that is weekly annotated because it does not show exactly the relation between the instant when a sign that is emitted and its corresponding translation.
-
-This study was developed to obtain the Masters Degree in Informatics and Computer Science and proposes a novel Signs Detection Method (nSDm) that allows signs detection in a weakly annotated dataset and tries to answer the following questions:
+This article reports research conducted with the goal of proposing a method for detecting PSL signs using a weakly annotated dataset, and answering the following research questions:
 
@@
 
-nSDm will take advantage of the PRL dataset \cite{lsp_dataset} and will contribute with the generalization and progressive signs detection that could be used as a starting point for future studies including software accessibility improvements and human computer interaction for people with deafness and hearing limitations.
+The Pontifical Catholic University of Peru (PUCP) Grammar and  Signs research group created a PSL dataset \cite{lsp_dataset} that is weakly annotated because it does not show exactly the relation between the instant when a sign that is emitted and its corresponding translation.
+
+nSDm will take advantage of the PSL dataset \cite{lsp_dataset} and will contribute with the generalization and progressive signs detection that could be used as a starting point for future studies including software accessibility improvements and human computer interaction for people with deafness and hearing limitations.
 

### 2019-09-02 article/article-tesis.tex
adding tesis assesor as co-author and adding Bag of Words to the state of art

@@
 \author[1]{Eric Raphael Huiza Pereyra}
-\affil[1]{Department of Postgraduate Studies, Pontifical Catholic University of Peru}
+\affil[1]{Pontifical Catholic University of Peru}
 \affil[]{\textit{eric.huiza@pucp.edu.pe}}
+
+\author[2]{Cesar Augusto Olivares Poggi}
+\affil[2]{Pontifical Catholic University of Peru}
+\affil[]{\textit{cesar.olivares@pucp.edu.pe}}
 
@@
 \subsection{Action Recognition}
-Human action recognition is a extensively studied field and available action recognition dataset: UCF101, HMDB51, THUMOS14  are available, researches tried to solve the human action recognition problem using different approaches including Optical Flow and 3D CNN \cite{qiu2017learning}.
+Human action recognition is a extensively studied field, action recognition dataset like UCF101, HMDB51, THUMOS14  are available, researches tried to solve the human action recognition problem using different approaches including Optical Flow and 3D CNN \cite{qiu2017learning}.
 \textbf{Optical Flow,} is defined as the pattern obtained from the motion of objects, surfaces and edges in a visual scene caused by the relative motion between the observer and a scene, it is achieved by distributing movement velocities and brightness across frames.
@@
 3D CNN are not as effective as optical flow to detect human actions on its own, 3D CNN can be trained to learn optical flow so we can avoid costly computation and storage and obtain task-specific motion representation  \cite{zhu2018hidden} and increase models performance, precision and exhaustiveness on human action recognition.
+\subsection{Pose Estimation}
 \subsection{Video Classification}
-\textbf{Bag of Words.}
+\textbf{Bag of Words (BoW)} or \textbf{Bag of Visual Words (BoVW)} when applied to Computer Vision is a technique used for images classification by counting the number of occurrences of a set of image features.
+Scale-invariant feature transform
 \\
-\textbf{Fussing parallel CNN.}
+Feature trajectory detection are much improved using statistical methods like \textbf{Fisher Vectors} obtaining better results over traditional BoW \textbf{Fussing parallel CNN.}
 \section{Method}\label{method}

### 2019-09-08 src/main/pre-processing/file_processing_handler.py
transcription text aggregation for text analysis using Amazon Comprenhend

@@ -1,6 +1,8 @@
 import logging
 
 from audio.audio_processor import AudioProcessor
+from text_analisis.keywords_detection_processor import KeywordsDetectionProcessor
+from text_analisis.text_aggregation_processor import TextAggregationProcessor
 from transcription.transcription_processor import TranscriptionProcessor
 from video.video_splitter import VideoSplitter
 
@@ -15,7 +17,9 @@ class FileProcessingHandler:
         operations = {
             'extract-audio': lambda x: AudioProcessor().process(x),
             'video-split': lambda x: VideoSplitter().process(x),
-            'transcribe-audio': lambda x: TranscriptionProcessor().process(x)
+            'transcribe-audio': lambda x: TranscriptionProcessor().process(x),
+            'text-aggregation': lambda x: TextAggregationProcessor(x),
+            'keywords-detection': lambda x: KeywordsDetectionProcessor().process(x)
         }
 
         if operation in operations:

### 2019-09-08 src/main/pre-processing/text_analisis/keywords_detection_processor.py

@@ -1,0 +1,12 @@
+import logging
+from processor import Processor
+
+
+class KeywordsDetectionProcessor(Processor):
+    def __init__(self):
+        super().__init__()
+
+        self.logger = logging.getLogger(__name__)
+
+    def process(self, data):
+        print()

### 2019-09-08 src/main/pre-processing/text_analisis/text_aggregation_processor.py

@@ -1,0 +1,33 @@
+import logging
+import json
+
+from processor import Processor
+
+
+class TextAggregationProcessor(Processor):
+    __transcription_files_base = "transcriptions"
+
+    def __init__(self):
+        super().__init__()
+
+        self.logger = logging.getLogger(__name__)
+
+    def process(self, data):
+        transcription_folders = self._s3.list_objects_v2(
+            Bucket=self._bucketName,
+            Prefix=self.__transcription_files_base)
+
+        for folder in transcription_folders['Contents']:
+            folder_key = folder['Key'] + '/chunks/'
+            transcription_files = self._s3.list_objects_v2(Bucket=self._bucketName, Prefix=folder_key)
+
+            for file in transcription_files['Contents']:
+                file_key = file['Key']
+                file_obj = self._s3.get_object(Bucket=self._bucketName, Key=file_key)
+                file_content = file_obj['Body'].read().decode('utf-8')
+                json_content = json.loads(file_content)
+
+                transcriptions = json_content['results']['transcripts']
+                for transcription in transcriptions:
+                    with open('/transcriptions/..') as tf:
+                        tf.write(transcription['transcript'])

### 2019-09-09 article/article-tesis.tex
adding additional bag of words and fisher vectors related work

--- 2019-06-30 article/article-tesis.tex
@@
 3D CNN are not as effective as optical flow to detect human actions on its own, 3D CNN can be trained to learn optical flow so we can avoid costly computation and storage and obtain task-specific motion representation  \cite{zhu2018hidden} and increase models performance, precision and exhaustiveness on human action recognition.
+\subsection{Pose Estimation}
 \subsection{Video Classification}
-\textbf{Bag of Words.}
-\\
-\textbf{Fussing parallel CNN.}
+\textbf{Bag of Words (BoW)} or \textbf{Bag of Visual Words (BoVW)} based on natural language processing techniques is one of the simplest and oldest local descriptor encoding strategies that counts the number of occurrences of a set of image features \cite{wang2019hallucinating}, when applied to Computer Vision is a technique used to create images representations or features vectors used that can be learned by CNNs, resulting on improved images classification and video classification.
+Feature trajectory detection are much improved using statistical methods like \textbf{Fisher Vectors} obtaining better results over traditional BoW \textbf{Fussing parallel CNN.}, the Bag of Visual Words representation suffers from sparsity and high dimensionality, in the other hand representations obtained using the Fisher Vectors kernel are more compact and dense which results on better results on image and video classification problems.
 \section{Method}\label{method}

### 2019-09-10 article/article-tesis.tex
adding a better audio transcription aggregation mechanism

@@
 \subsection{Video Classification}
-\textbf{Bag of Words (BoW)} or \textbf{Bag of Visual Words (BoVW)} based on natural language processing techniques is one of the simplest and oldest local descriptor encoding strategies that counts the number of occurrences of a set of image features \cite{wang2019hallucinating}, when applied to Computer Vision is a technique used to create images representations or features vectors used that can be learned by CNNs, resulting on improved images classification and video classification.
+\textbf{Bag of Words (BoW)} or \textbf{Bag of Visual Words (BoVW)} based on natural language processing techniques is one of the simplest and oldest local descriptor encoding strategies.
+In its simplest form, it consists of (i) clustering with k-means a collection of descriptor vectors from the training set to build so-called visual vocabulary, (ii) as signing each descriptor to its nearest cluster center from the visual dictionary, and (iii) aggregating the one-hot assignment vectors via average pooling \cite{wang2019hallucinating}, when applied to Computer Vision is a technique used to create images representations or features vectors used that can be learned by CNNs, resulting on improved images classification and video classification.
 Feature trajectory detection are much improved using statistical methods like \textbf{Fisher Vectors} obtaining better results over traditional BoW \textbf{Fussing parallel CNN.}, the Bag of Visual Words representation suffers from sparsity and high dimensionality, in the other hand representations obtained using the Fisher Vectors kernel are more compact and dense which results on better results on image and video classification problems.

### 2019-09-10 src/main/pre-processing/common/io_utils.py

@@ -1,12 +1,16 @@
 import os
 import ntpath
 
-CHUNKS_FOLDER = 'chunks'
-INCOMING_QUEUE = 'incoming-queue'
-BINARY_FOLDER = 'binary'
+chunks_folder = 'chunks'
+incoming_queue_folder = 'incoming-queue'
+binary_folder = 'binary'
 
 
 def get_filename_without_extension(file_path):
+    return os.path.splitext(ntpath.basename(file_path))[0]
+
+
+def get_filename(file_path):
     return os.path.splitext(ntpath.basename(file_path))[0]
 
 
@@ -23,11 +27,11 @@ def add_prefix_to_filename(file_path, prefix, separator='-'):
 
 def get_video_chunk_path(video_path, idx):
     base_dir = os.path.dirname(video_path)
-    return os.path.join(base_dir, CHUNKS_FOLDER, add_prefix_to_filename(video_path, str(idx).zfill(2)))
+    return os.path.join(base_dir, chunks_folder, add_prefix_to_filename(video_path, str(idx).zfill(2)))
 
 
 def get_video_chunk_base_key(video_key):
-    return os.path.join(os.path.dirname(video_key), CHUNKS_FOLDER)
+    return os.path.join(os.path.dirname(video_key), chunks_folder)
 
 
 def check_path_dir(file_path):

### 2019-09-10 src/main/pre-processing/common/processor.py

@@ -13,7 +13,7 @@ class Processor:
         self._s3 = boto3.client('s3')
         self._bucketName = os.getenv('S3_BUCKET', None)
         self._work_dir = os.getenv('WORK_DIR', './')
-        self._binary_folder = os.path.join(self._work_dir, io_utils.BINARY_FOLDER)
+        self._binary_folder = os.path.join(self._work_dir, io_utils.binary_folder)
         self._source_videos_base = 'not-annotated-videos'
         self._always_upload = os.getenv('ALWAYS_UPLOAD_S3', 'True') == 'True'
 

### 2019-09-10 src/main/pre-processing/file_processing_handler.py

@@ -2,7 +2,7 @@ import logging
 
 from audio.audio_processor import AudioProcessor
 from text_analisis.keywords_detection_processor import KeywordsDetectionProcessor
-from text_analisis.text_aggregation_processor import TextAggregationProcessor
+from text_analisis.audio_transcription_aggregation_processor import AudioTranscriptionAggregationProcessor
 from transcription.transcription_processor import TranscriptionProcessor
 from video.video_splitter import VideoSplitter
 
@@ -18,7 +18,7 @@ class FileProcessingHandler:
             'extract-audio': lambda x: AudioProcessor().process(x),
             'video-split': lambda x: VideoSplitter().process(x),
             'transcribe-audio': lambda x: TranscriptionProcessor().process(x),
-            'text-aggregation': lambda x: TextAggregationProcessor(x),
+            'audio-transcript-aggregation': lambda x: AudioTranscriptionAggregationProcessor().process(x),
             'keywords-detection': lambda x: KeywordsDetectionProcessor().process(x)
         }
 

### 2019-09-10 src/main/pre-processing/main.py

@@ -31,7 +31,7 @@ def setup_logging(default_level=logging.INFO):
 
 
 def main():
-    incoming_queue = os.path.join(working_folder, io_utils.INCOMING_QUEUE)
+    incoming_queue = os.path.join(working_folder, io_utils.incoming_queue_folder)
 
     logger.info('Pre-processing is initiated with the following parameters')
     logger.info('Incoming Queue: %s', incoming_queue)

### 2019-09-10 src/main/pre-processing/text_analisis/audio_transcription_aggregation_processor.py

@@ -1,0 +1,56 @@
+import logging
+import json
+import os
+
+from processor import Processor
+from common import io_utils
+
+
+class AudioTranscriptionAggregationProcessor(Processor):
+    __transcription_files_prefix = "transcriptions/"
+    __transcriptions_folder = "transcriptions"
+
+    def __init__(self):
+        super().__init__()
+
+        self.logger = logging.getLogger(__name__)
+
+    def process(self, data):
+        transcription_folders = self._s3.list_objects_v2(
+            Bucket=self._bucketName,
+            StartAfter=self.__transcription_files_prefix)
+
+        local_file_name = data['localFileName']
+        s3_key = data['s3Key']
+
+        if self.s3_exist_object(s3_key):
+            self.logger.debug('audio transcription aggregation already available in s3 at: %s', s3_key)
+            return
+
+        if os.path.exists(local_file_name):
+            self.logger.debug('local file %s exist', local_file_name)
+            os.remove(local_file_name)
+            self.logger.debug('local file %s was deleted', local_file_name)
+
+        io_utils.check_path_dir(local_file_name)
+
+        for transcription_dict in transcription_folders['Contents']:
+            transcription_key = transcription_dict['Key']
+            if not transcription_key.endswith('.json'):
+                continue
+
+            self.logger.debug('Extracting audio transcription from: %s', transcription_key)
+
+            file_obj = self._s3.get_object(Bucket=self._bucketName, Key=transcription_key)
+            file_content = file_obj['Body'].read().decode('utf-8')
+            json_content = json.loads(file_content)
+
+            transcriptions = json_content['results']['transcripts']
+            for transcription in transcriptions:
+                with open(local_file_name, 'a+') as tf:
+                    tf.write(transcription['transcript'] + '\n')
+
+        self.logger.debug('audio transcription aggregation is available locally at: %s', local_file_name)
+
+        self.upload_file(local_file_name, s3_key)
+        self.logger.debug('audio transcription aggregation uploaded to: %s', s3_key)

### 2019-09-10 src/main/pre-processing/text_analisis/text_aggregation_processor.py

@@ -1,33 +1,0 @@
-import logging
-import json
-
-from processor import Processor
-
-
-class TextAggregationProcessor(Processor):
-    __transcription_files_base = "transcriptions"
-
-    def __init__(self):
-        super().__init__()
-
-        self.logger = logging.getLogger(__name__)
-
-    def process(self, data):
-        transcription_folders = self._s3.list_objects_v2(
-            Bucket=self._bucketName,
-            Prefix=self.__transcription_files_base)
-
-        for folder in transcription_folders['Contents']:
-            folder_key = folder['Key'] + '/chunks/'
-            transcription_files = self._s3.list_objects_v2(Bucket=self._bucketName, Prefix=folder_key)
-
-            for file in transcription_files['Contents']:
-                file_key = file['Key']
-                file_obj = self._s3.get_object(Bucket=self._bucketName, Key=file_key)
-                file_content = file_obj['Body'].read().decode('utf-8')
-                json_content = json.loads(file_content)
-
-                transcriptions = json_content['results']['transcripts']
-                for transcription in transcriptions:
-                    with open('/transcriptions/..') as tf:
-                        tf.write(transcription['transcript'])

### 2019-09-10 src/main/pre-processing/text_analisis/keywords_detection_processor.py
initial work for audio transcription keyphrases detection

@@ -1,4 +1,5 @@
 import logging
+import boto3
 from processor import Processor
 
 
@@ -7,6 +8,17 @@ class KeywordsDetectionProcessor(Processor):
         super().__init__()
 
         self.logger = logging.getLogger(__name__)
+        self._comprehend = boto3.client('comprehend')
 
     def process(self, data):
-        print()
+        input_location = data['inputLocation']
+        output_location = data['outputLocation']
+
+        self._comprehend.start_key_phrases_detection_job(
+            InputDataConfig={
+                'S3Uri': input_location
+            },
+            OutputDataConfig={
+                'S3Uri': output_location
+            }
+        )

### 2019-09-10 src/main/pre-processing/text_analisis/keywords_detection_processor.py
adding a basic get job name method

@@ -4,6 +4,8 @@ from processor import Processor
 
 
 class KeywordsDetectionProcessor(Processor):
+    __preferred_language_code = 'es'
+
     def __init__(self):
         super().__init__()
 
@@ -20,5 +22,11 @@ class KeywordsDetectionProcessor(Processor):
             },
             OutputDataConfig={
                 'S3Uri': output_location
-            }
+            },
+            JobName=self.__get_job_name(),
+            LanguageCode=self.__preferred_language_code
         )
+
+    @staticmethod
+    def __get_job_name(self):
+        return ''

### 2019-09-22 src/main/pre-processing/common/io_utils.py
adding support for RGB video frames generation

@@ -36,10 +36,14 @@ def get_video_chunk_base_key(video_key):
     return os.path.join(os.path.dirname(video_key), chunks_folder)
 
 
-def check_path_dir(file_path):
+def check_path_file(file_path):
     if not os.path.exists(os.path.dirname(file_path)):
         os.makedirs(os.path.dirname(file_path), exist_ok=True)
 
+
+def check_path_dir(dir_path):
+    if not os.path.exists(dir_path):
+        os.makedirs(dir_path, exist_ok=True)
 
 def change_extension(file_path, new_extension):
     pre, ext = os.path.splitext(file_path)

### 2019-09-22 src/main/pre-processing/common/processor.py

@@ -32,6 +32,10 @@ class Processor:
         else:
             print('the file is already available at: {}'.format(key))
 
+    def download_file(self, key, file_path):
+        self._s3.download_file(self._bucketName, key, file_path)
+        self.logger.debug('%s was downloaded to %s', key, file_path)
+
     def s3_exist_object(self, key):
         try:
             self._s3.head_object(Bucket=self._bucketName, Key=key)
@@ -59,6 +63,6 @@ class Processor:
                 continue
 
             file_local_path = os.path.join(self._work_dir, file_key)
-            io_utils.check_path_dir(file_local_path)
+            io_utils.check_path_file(file_local_path)
             self._s3.download_file(self._bucketName, file_key, file_local_path)
             self.logger.debug('%s was downloaded to %s', file_key, file_local_path)

### 2019-09-22 src/main/pre-processing/file_processing_handler.py

@@ -6,6 +6,7 @@ from text_analisis.transcription_syntax_detection_processor import Transcription
 from transcription.transcription_processor import TranscriptionProcessor
 from video.video_splitting_processor import VideoSplittingProcessor
 from text_analisis.samples_metadata_generation_processor import SamplesMetadataGenerationProcessor
+from video.rgb_samples_generation_processor import RGBSamplesGenerationProcessor
 
 
 class FileProcessingHandler:
@@ -21,7 +22,8 @@ class FileProcessingHandler:
             'transcribe-audio': lambda x: TranscriptionProcessor().process(x),
             'transcription-syntax-detection': lambda x: TranscriptionSyntaxDetectionProcessor().process(x),
             'keywords-detection': lambda x: KeywordsDetectionProcessor().process(x),
-            'samples-generation': lambda x: SamplesMetadataGenerationProcessor().process(x)
+            'samples-metadata-generation': lambda x: SamplesMetadataGenerationProcessor().process(x),
+            'rgb-samples-generation': lambda x: RGBSamplesGenerationProcessor().process(x)
         }
 
         if operation in operations:

### 2019-09-22 src/main/pre-processing/text_analisis/transcription_syntax_detection_processor.py

@@ -76,7 +76,7 @@ class TranscriptionSyntaxDetectionProcessor(Processor):
             os.remove(local_file_path)
             self.logger.debug('local file %s was deleted', local_file_path)
 
-        io_utils.check_path_dir(local_file_path)
+        io_utils.check_path_file(local_file_path)
 
         return local_file_path, s3_key
 

### 2019-09-22 src/main/pre-processing/video/rgb_samples_generation_processor.py

@@ -1,0 +1,75 @@
+import os
+import shutil
+import logging
+import csv
+
+from common.processor import Processor
+from rgb_video_frames_handler import RGBVideoFramesHandler
+from common import io_utils
+
+
+class RGBSamplesGenerationProcessor(Processor):
+    __videos_dataset_rgb_path = 'dataset/rgb'
+
+    def __init__(self):
+        super().__init__()
+
+        self.logger = logging.getLogger(__name__)
+        
+    def process(self, data):
+        metadata_file_path = os.path.join(self._work_dir, data['metadata'])
+        if not os.path.exists(metadata_file_path):
+            self.logger.debug('the %s metadata file does not exist, the process cannot continue', metadata_file_path)
+            return
+
+        reset_dataset = bool(data['reset_dataset'])
+        if reset_dataset and os.path.exists(self.__videos_dataset_rgb_path):
+            shutil.rmtree(os.path.join(self._work_dir, self.__videos_dataset_rgb_path))
+
+        delay_factor = float(data['translation_delay_factor'])
+        metadata_fields = ['token', 'video_path', 'start_time', 'end_time']
+
+        with open(metadata_file_path) as metadata_file:
+            reader = csv.DictReader(metadata_file, fieldnames=metadata_fields)
+            next(reader, None)
+            for metadata in reader:
+                self.__handle_sample_metadata(metadata, delay_factor)
+
+    def __handle_sample_metadata(self, metadata, delay_factor: float):
+        token = metadata['token']
+        video_key = metadata['video_path']
+        video_local_path = os.path.join(self._work_dir, metadata['video_path'])
+
+        original_start_time = float(metadata['start_time'])
+        original_end_time = float(metadata['end_time'])
+
+        start_time = self.__get_start_time(original_start_time, delay_factor)
+        end_time = original_end_time + 0.5
+
+        io_utils.check_path_file(video_local_path)
+        if not os.path.exists(video_local_path):
+            self.download_file(video_key, video_local_path)
+
+        self.logger.debug('Starting a RGB sample generation with the following parameters => token: %s, '
+                          'video_path: %s, start_time: %s, end_time: %s', token, video_local_path, start_time, end_time)
+
+        sample_folder_path = self.__get_sample_folder_path(token)
+        io_utils.check_path_dir(sample_folder_path)
+
+        handler = RGBVideoFramesHandler()
[…]

### 2019-09-22 src/main/pre-processing/video/rgb_video_frames_handler.py

@@ -7,14 +7,15 @@ from common.handler import Handler
 class RGBVideoFramesHandler(Handler):
     def handle(self, **kwargs):
         video_path = kwargs['VideoPath']
-        start_time = kwargs['StartTime']
-        end_time = kwargs['EndTime']
+        start_time = float(kwargs['StartTime'])
+        end_time = float(kwargs['EndTime'])
         folder_path = kwargs['FolderPath']
 
         video_capture = cv2.VideoCapture(video_path)
         fps = video_capture.get(cv2.CAP_PROP_FPS)
-        # frame_count = video_capture.get(cv2.CAP_PROP_FRAME_COUNT)
-        # duration = float(frame_count) / float(fps)
+
+        frame_count = video_capture.get(cv2.CAP_PROP_FRAME_COUNT)
+        duration = float(frame_count) / float(fps)
 
         start_frame_no = 0 if int(start_time * fps) - 1 < 0 else int(start_time * fps) - 1
         end_frame_no = 0 if int(end_time * fps) - 1 < 0 else int(end_time * fps) - 1
@@ -22,12 +23,14 @@ class RGBVideoFramesHandler(Handler):
         video_capture.set(cv2.CAP_PROP_POS_FRAMES, start_frame_no)
 
         frame_no = start_frame_no
+        img_index = 1
         while frame_no <= end_frame_no:
-            frame_file_path = os.path.join(folder_path, "frame%d.jpg" % frame_no)
+            frame_file_path = os.path.join(folder_path, "frame%d.jpg" % img_index)
             success, frame = video_capture.read()
             if not success:  # eof
                 break
             cv2.imwrite(frame_file_path, frame)
             frame_no += 1
+            img_index += 1
 
         video_capture.release()

### 2019-09-22 src/main/pre-processing/video/video_splitting_processor.py

@@ -42,7 +42,7 @@ class VideoSplittingProcessor(Downloader):
             datetime.combine(datetime.today(), end_time.time()) - datetime.combine(datetime.today(), start_time.time())
 
         chunk_path = io_utils.get_video_chunk_path(self._video_file_path, idx)
-        io_utils.check_path_dir(chunk_path)
+        io_utils.check_path_file(chunk_path)
 
         cmd = "ffmpeg -i {} -ss {} -t {} {} -loglevel panic -y".format(
             self._video_file_path, start_time.strftime('%H:%M:%S'), str(duration), chunk_path)

### 2019-09-23 article/article-tesis.tex
video annotation process updates reflected on the design, article was updated to reflect manual video clean up

@@
-\documentclass[twocolumn]{article}
+\documentclass[twocolumn,conference]{article}
 
@@
 \usepackage{enumerate}
+\usepackage{caption}
+
+\usepackage{array,booktabs,longtable,tabularx}
+\usepackage{ltablex}
+\usepackage{wrapfig,lipsum,booktabs}
+\usepackage{multirow}
 
@@
 	\vspace{-2.0cm}
-	\textbf{Talking with body expressions} \\	
+	\textbf{Talking with signs} \\	
 	\Large \textbf{A method for detecting signs in a weakly annotated dataset}
@@
 People with deafness or hearing disabilities who aim to use computer based systems, state of art video classification and human action recognition rely in a combination of traditional and new methods like optical flow and 3D CNN.
-In this article we show a method to detect Peruvian Sign Language elements from a weakly annotated videos dataset.
+In this article we show a novel method to detect Peruvian Sign Language elements (nSDm) from a weakly annotated videos dataset.
 Video annotation techniques were applied increasing the models performance, as a result we achieved XX precision and XX exhaustiveness by training a 3D CNN with optical flow inputs.
@@
 \subsection{Dataset Description \cite{lsp_dataset}}\label{datasetdesc}
-The LSP dataset was developed by the PUCP Grammar and Signs research group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female informants, all of them are Lima Peru residents and reported to be borne with a permanent deafness condition or acquired the condition before the acquisition of Spanish.
+The PSL dataset was developed by the PUCP Grammar and Signs research group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female informants, all of them are Lima Peru residents and reported to be borne with a permanent deafness condition or acquired the condition before the acquisition of Spanish.
 
@@
 
-The video clips were recorded in three sessions with the following participants: A coordinator, a LSP \cite{lsp_2015} translator and a informant.\\
+The video clips were recorded in three sessions with the following participants: A coordinator, a PSL \cite{lsp_2015} translator and a informant.\\
 
@@
 
-\textbf{Recording Session 3}: A LSP \cite{lsp_2015} conversation facilitated by the coordinator happening between the informant and the translator.\\
+\textbf{Recording Session 3}: A PSL \cite{lsp_2015} conversation facilitated by the coordinator happening between the informant and the translator.\\
 
-During all the sessions a LSP \cite{lsp_2015} translator performs a translation after a word or phrase is completed.
-The dataset is weakly annotated given there is not direct relation between the translation and the instant when the sign is emitted.
+During all the sessions a PSL \cite{lsp_2015} translator performs a translation after a word or phrase is completed.
 
 \subsection{Video Annotation}\label{videoannot}
-\begin{figure}[ht!]
-\includegraphics[width=\linewidth]{video-annotation-pipeline.png}
+The PSL dataset is weakly annotated because there is no a direct relation between the instant when a sign is emitted and when its translation to Spanish is delivered, nSDm proposes a semi automatic video annotation pipeline described in Figure \ref{fig:video-annotation-pipeline} for cleaning, pre-processing and analyzing PSL videos in order to produce an annotated PSL dataset that can be used for training the nSDm model using supervised learning.
+The pipeline is described in detail in \ref{manual-video-cleanup}, \ref{video-pre-processing}, \ref{audio-transcription-analysis} and \ref{samples-generation}
+
+\begin{figure}[hbt!]
+\includegraphics[width=\linewidth]{images/video-annotation-pipeline.png}
 \caption{Video annotation process}
+\label{fig:video-annotation-pipeline}
 \end{figure}
+
+\subsubsection{Manual Video Clean Up}\label{manual-video-cleanup}
+The PSL recordings contain a considerable amount of noise introduced during the recording sessions, it makes difficult to find video intervals that clearly show a relation between signs emitted by the informant and the translation delivered by the translator, the noise factors are listed in the following list:
+\begin{itemize}
+	\item Multiple participants speaking during the session.
+	\item Conversations between participants that are not relevant to emitted sings
+	\item High frequency of silent periods 
+\end{itemize}
+A manual video cleanup process is required to find noise free video segments (table \ref{table:noise-free-video-segments} lists a noise free video segments), it is a long process that required watching all the videos available in the PSL dataset and manually writing in a database when an informant started emitting signs and when the translator finished delivering a translation.
+\begin{table}[!htb]
+\captionsetup{font=footnotesize}
+\centering
+\begin{tabular}{lrrc}
+\toprule
+\multicolumn{1}{c}{\textbf{Video}} & 
+	\multicolumn{1}{c}{\textbf{Start}} &
+	\multicolumn{1}{c}{\textbf{End}} &
+	\multicolumn{1}{c}{\textbf{Alignment}}\\
+\midrule
+\multirow{5}{7.5em}{\textbf{consultant-02-session-01-part-01.mp4}} & 00:15 & 00:21 & center\\
+& 00:15 & 00:21 & center\\
+& 00:53 & 01:07 & center\\
+& 02:43 & 02:47 & center\\
+& 17:33 & 18:01 & left\\
+\bottomrule
+\end{tabular}
+\captionsetup{font=footnotesize}
+\caption{Noise free video segments extract} \label{table:noise-free-video-segments}
+\end{table}
+The recordings show the informant in two alignments (centered and left), the manual video clean up process also stores the informant alignment, table \ref{table:informant-alignment} shows the two available alignments, we use the alignment annotation later in the process during the video frames extraction to create the PSL annotated dataset.
+\begin{table}[!htb]
+\captionsetup{font=footnotesize}
+\centering
+\begin{tabular}{cc}
+\toprule
+\multicolumn{1}{c}{\textbf{Center Aligned}} & 
+	\multicolumn{1}{c}{\textbf{Left Aligned}}\\
+\midrule
+\includegraphics{images/informant-center-alignment.png}& \includegraphics{images/informant-left-alignment.png}\\
+\bottomrule
+\end{tabular}
+\caption{Informant Alignment} \label{table:informant-alignment}
+\end{table}
+
+\subsubsection{Video Pre-Processing}\label{video-pre-processing}
+\subsubsection{Audio Transcription and Text Analysis}\label{audio-transcription-analysis}
+\begin{table}[!htb]
+\captionsetup{size=footnotesize}
+\begin{tabular}{ p{16em} c}
+\toprule
+\multicolumn{1}{c}{\textbf{Token}} & 
+	\multicolumn{1}{c}{\textbf{Frequency}}\\
+\midrule
+\textbf{pareja}&	4\\
+\textbf{cosas}&	3\\
+\textbf{cine}&	2\\
+\textbf{noche}&	2\\
+\textbf{terror}&	1\\
+\textbf{parque}&	1\\
+\textbf{casa}&	1\\
+\textbf{montón}&	1\\
+\textbf{apariciones}&	1\\
+\textbf{fantasmas}&	1\\
+\textbf{dos}&	1\\
+\bottomrule
+\end{tabular}
+\caption{Most relevant tokens detection frequency in the PSL dataset} \label{tab:token-freq}
+\end{table}
+
+\begin{table}[!htb]
+\captionsetup{font=footnotesize}
+\centering
+\begin{tabular}{ l p{10em} r r }
+%\resizebox{\textwidth}{!}{\begin{tabular}{ llrr }
+\toprule
+\multicolumn{1}{c}{\textbf{Token}} & 
+	\multicolumn{1}{c}{\textbf{Video}} &
+	\multicolumn{1}{c}{\textbf{Start}} &
+	\multicolumn{1}{c}{\textbf{End}}\\
+\midrule
+\textbf{cine}&	consultant-02-session-01-part-01-00.mp4&	4.19&	4.75\\
+\textbf{cine}&	consultant-02-session-01-part-01-01.mp4&	1.19&	1.75\\
+\textbf{terror}&	consultant-02-session-01-part-01-01.mp4&	3.82&	4.4\\
+\textbf{parque}&	consultant-02-session-01-part-01-03.mp4&	8.97&	9.3\\
+\textbf{casa}&	consultant-02-session-01-part-01-03.mp4&	10.12&	10.57\\
+\textbf{pareja}&	consultant-02-session-01-part-01-04.mp4&	3.91&	4.36\\
+\textbf{noche}&	consultant-02-session-01-part-01-04.mp4&	4.49&	4.92\\
+\textbf{noche}&	consultant-02-session-01-part-01-04.mp4&	7.91&	8.2\\
+\textbf{mont\'on}&	consultant-02-session-01-part-01-04.mp4&	8.5&	8.78\\
+\textbf{cosas}&	consultant-02-session-01-part-01-04.mp4&	8.88&	9.38\\
+\bottomrule
+\end{tabular}
+%\end{tabular}}
+\caption{Shows metadata extacted from the PSL dataset: (1)\textit{Token} could be a noun or a number (2)\textit{Video Path} shows the video where the token was detected (3)\textit{Start Time} time when the token reproduction starts (4)\textit{End Time} time when the token reproduction ends.}
+\end{table}
+
+\subsubsection{Samples Generation}\label{samples-generation} 
 \section{Conclusion}\label{conclusion}

### 2019-10-07 article/article-tesis.tex
adding a new section for Video Annotation method on the article

@@
 \section{Method}\label{method}
-\section{Experimentation}\label{experimentation}
-\subsection{Dataset Description \cite{lsp_dataset}}\label{datasetdesc}
-The PSL dataset was developed by the PUCP Grammar and Signs research group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female informants, all of them are Lima Peru residents and reported to be borne with a permanent deafness condition or acquired the condition before the acquisition of Spanish.
-
-The dataset consists in 718 video clips recorded with a ADR-CX220 SONY HD camera which included an embedded microphone.
-The camera focused only the informant but also recorded questions, instructions and translations.
-
-The video clips were recorded in three sessions with the following participants: A coordinator, a PSL \cite{lsp_2015} translator and a informant.\\
-
-\textbf{Recording Session 1}: A 45-60 minutes semi structured interview that included: Biographic information as well as habits, anecdotes, opinion about cultural subjects and elicitation of names, states and actions.
-
-\textbf{Recording Session 2}: The informant was presented with a set of 55 cards describing actions and were asked to choose a set of them in order to build a coherent story that was subsequently told by the informant.
-
-\textbf{Recording Session 3}: A PSL \cite{lsp_2015} conversation facilitated by the coordinator happening between the informant and the translator.\\
-
-During all the sessions a PSL \cite{lsp_2015} translator performs a translation after a word or phrase is completed.
-
 \subsection{Video Annotation}\label{videoannot}
-The PSL dataset is weakly annotated because there is no a direct relation between the instant when a sign is emitted and when its translation to Spanish is delivered, nSDm proposes a semi automatic video annotation pipeline described in Figure \ref{fig:video-annotation-pipeline} for cleaning, pre-processing and analyzing PSL videos in order to produce an annotated PSL dataset that can be used for training the nSDm model using supervised learning.
+The PSL dataset is weakly annotated because there is not a direct relation between the instant when a sign is emitted and when its translation to Spanish is delivered, nSDm proposes a semi automatic video annotation pipeline described in Figure \ref{fig:video-annotation-pipeline} for cleaning, pre-processing and analyzing PSL videos in order to produce an annotated PSL dataset that can be used for training the nSDm model using supervised learning.
 The pipeline is described in detail in \ref{manual-video-cleanup}, \ref{video-pre-processing}, \ref{audio-transcription-analysis} and \ref{samples-generation}
@@
 
-\subsubsection{Manual Video Clean Up}\label{manual-video-cleanup}
-The PSL recordings contain a considerable amount of noise introduced during the recording sessions, it makes difficult to find video intervals that clearly show a relation between signs emitted by the informant and the translation delivered by the translator, the noise factors are listed in the following list:
+\subsubsection{Manual Automatic Video Clean Up}\label{manual-video-cleanup}
+The PSL recordings described on \ref{datasetdesc} contain a considerable amount of noise introduced during the recording sessions which makes difficult to easily find video intervals that clearly show a relation between signs emitted by the informant and the translation delivered by the translator, the noise factors are listed in the following list:
 \begin{itemize}
 	\item Multiple participants speaking during the session.
-	\item Conversations between participants that are not relevant to emitted sings
-	\item High frequency of silent periods 
+	\item Conversations between participants that are not relevant to emitted sings.
+	\item High frequency of silent periods.
 \end{itemize}
-A manual video cleanup process is required to find noise free video segments (table \ref{table:noise-free-video-segments} lists a noise free video segments), it is a long process that required watching all the videos available in the PSL dataset and manually writing in a database when an informant started emitting signs and when the translator finished delivering a translation.
+A manual video cleanup process is required to find noise free video intervals, it is a manual process that requires watching all the videos available in the PSL dataset and manually annotating the instant when an informant started emitting sings and the instant when the translator completed delivering a translation, table \ref{table:noise-free-video-segments} shows an annotation example.
+
+The recordings show the informant in two alignments (centered and left), the manual video clean up process also stores the informant alignment, table \ref{table:informant-alignment} shows the two available alignments, we use the alignment annotation later in the process during the video frames extraction to create the PSL annotated dataset.
+
+% Noise free video segment extract table
 \begin{table}[!htb]
@@
 	\multicolumn{1}{c}{\textbf{Alignment}}\\
+\midrule
+\multirow{5}{7.5em}{\textbf{consultant-01-session-01-part-01.mp4}} & 00:30 & 00:55 & center\\
+& 01:15 & 01:29 & center\\
+& 00:53 & 01:07 & center\\
+& 08:12 & 09:01 & center\\
 \midrule
@@
 \end{table}
-The recordings show the informant in two alignments (centered and left), the manual video clean up process also stores the informant alignment, table \ref{table:informant-alignment} shows the two available alignments, we use the alignment annotation later in the process during the video frames extraction to create the PSL annotated dataset.
+
+% Informat aligment
 \begin{table}[!htb]
@@
 \subsubsection{Video Pre-Processing}\label{video-pre-processing}
-\subsubsection{Audio Transcription and Text Analysis}\label{audio-transcription-analysis}
+Weakly PSL annotated videos require processing before any metadata can be extracted, we propose a sequence of pre-processing tasks that take advantage of the annotation generated on \ref{manual-video-cleanup}.
+A \textbf{video splitting processor} generates a set of video chunks using the \textit{ffmpeg} multimedia framework and stores the produced chunks in Amazon S3 for later usage.
+The audio within the video chunks is then transcribed by an \textbf{audio transcription processor}, using the Amazon Transcription service, we selected the Amazon Transcription service because it provides an accurate mapping between audio participants and transcribed words along with useful metadata that describes the start and end time when words are pronounced.
+
+At the moment of writing this work Amazon Transcription service only supported Spanish for Spain and the US, it caused certain words that are specific for Peruvian Spanish not being fully recognized, in order to improve transcription accuracy we provided a custom vocabulary containing Peruvian expressions which improved  Peruvian words recognition, for the matters of this work unrecognized Peruvian words were omitted and not processed.
+
+\subsubsection{Audio Transcription Analysis}\label{audio-transcription-analysis}
+Audio transcription requires additional processing in order to produce useful information that leads to a successful PSL signs detection.
+Bag of Embedding Words (BoEW) is a widely used technique on Natural Language Processing tasks providing a easy and flexible way to list the most relevant words based on frequency.
+This work is focused on detecting nouns and numbers (nSDm is a ongoing method that will be progressively improved to handle a wider set of PSL elements) assuming that nouns (numbers are a subset of nouns) suffer less variations in spoken Spanish than verbs, pronouns, adverbs and adjectives, and provide more semantic value than conjunctions, prepositions and interjections.
+
 \begin{table}[!htb]
@@
 \midrule
-\textbf{pareja}&	4\\
-\textbf{cosas}&	3\\
-\textbf{cine}&	2\\
-\textbf{noche}&	2\\
-\textbf{terror}&	1\\
-\textbf{parque}&	1\\
-\textbf{casa}&	1\\
-\textbf{montón}&	1\\
-\textbf{apariciones}&	1\\
-\textbf{fantasmas}&	1\\
-\textbf{dos}&	1\\
+\textbf{pareja}&	40\\
+\textbf{cosas}&	30\\
+\textbf{cine}&	20\\
+\textbf{noche}&	20\\
+\textbf{terror}&	10\\
+\textbf{parque}&	10\\
+\textbf{casa}&	10\\
+\textbf{montón}&	10\\
+\textbf{apariciones}&	10\\
+\textbf{fantasmas}&	10\\
+\textbf{dos}&	10\\
 \bottomrule
@@
 
+We used Amazon comprehend for text analysis, specifically the syntax detection functionality which will provide a comprehensive list of detected language elements along with a score from 0.0 to 1.0 indicating the detection accuracy, we have selected the ones that have at least a 0.8 accuracy score and omitted the rest, this process was automated using a \textbf{transcription detection processor} which uses BoEW to provide a list of most relevant nouns and numbers based on appearance frequency.
+Table \ref{tab:token-freq} shows a list of nouns and numbers and their frequencies in the PSL dataset.
+
+Once a weighted list of nouns and numbers is generated a mapping showing when nouns and numbers appear in videos is required, moving forward called \textbf{Samples Metadata}.
+Table \ref{tab:token-video-mapping} shows mapping metadata extracted from PSL.
+
+\subsubsection{Samples Generation}\label{samples-generation} 
+nSDm requires PSL elements to be represented as a set of RGB frames and a calculated Optical Flow using the Lucas-Kanade method, both representations are inputs of two different models as presented on \ref{nsdm}.
+
+\textbf{Translation Delay Factor: } The difference in time between the instant when a sign is emitted and when a translation for that given sign is delivered is uncertain, we are calling that uncertainty as the \textit{translation delay factor} which we are trying to approximate using a constant value, we have chosen a 3 seconds translation delay factor assuming that most of the translations will occur between three seconds after a sign is emitted.
+
+A \textbf{RGB Samples generation processor} uses samples metadata in combination with the translation delay factor to determine frames that represent a given PSL element.
+We use \textit{OpenCV} to extract frames and store them following a hierarchical folder structure (listing \ref{list:rgb-samples-folders}) that nSDm data loaders will use to feed data into the RGM branch on the nSDm model architecture during training and testing.
+
+A \textbf{Optical Flow Samples generation processor} uses video frames and the hierarchical folder structure generated by the RGM samples generation processor to calculate an Optical Flow representation for PSL elements and store them in a hierarchical folder structure that will also used by the nSDm data loaders to feed the optical flow branch on the nSDm model architecture during training and testing.
+We selected optical flow as a samples generation strategy because its ability to represent movement traces from previous frames, it is particular useful for representing PSL elements, a PSL element is composed on different body movements including: elbow, arms, neck, eyes, shoulders and hands, which are performed quickly, a way to detect movement traces between frames allows to generate a single image representation of all movement involved on a sign emitting (Figure \ref{fig:opticalflow-two}).
+
 \begin{table}[!htb]
@@
 \begin{tabular}{ l p{10em} r r }
-%\resizebox{\textwidth}{!}{\begin{tabular}{ llrr }
 \toprule
@@
 \caption{Shows metadata extacted from the PSL dataset: (1)\textit{Token} could be a noun or a number (2)\textit{Video Path} shows the video where the token was detected (3)\textit{Start Time} time when the token reproduction starts (4)\textit{End Time} time when the token reproduction ends.}
+\label{tab:token-video-mapping}
 \end{table}
 
-\subsubsection{Samples Generation}\label{samples-generation} 
+\subsection{Novel Signs Detection Method (nSDm)}\label{nsdm}
+\subsubsection{Architecture}\label{nsdm-architecture}
+\begin{figure}[hbt!]
+\includegraphics[width=\linewidth]{images/network-architecture.png}
+\caption{nSDm network architecture}
+\label{fig:opticalflow-two}
+\end{figure}
+\subsection{New Videos Testing}
+
+\section{Experimentation}\label{experimentation}
+\subsection{Dataset Description \cite{lsp_dataset}}\label{datasetdesc}
+The PSL dataset was developed by the PUCP Grammar and Signs research group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female informants, all of them are Lima Peru residents and reported to be borne with a permanent deafness condition or acquired the condition before the acquisition of Spanish.
+
+The dataset consists in 718 video clips recorded with a ADR-CX220 SONY HD camera which included an embedded microphone.
+The camera focused only the informant but also recorded questions, instructions and translations.
+
+The video clips were recorded in three sessions with the following participants: A coordinator, a PSL \cite{lsp_2015} translator and a informant.\\
+
+\textbf{Recording Session 1}: A 45-60 minutes semi structured interview that included: Biographic information as well as habits, anecdotes, opinion about cultural subjects and elicitation of names, states and actions.
+
+\textbf{Recording Session 2}: The informant was presented with a set of 55 cards describing actions and were asked to choose a set of them in order to build a coherent story that was subsequently told by the informant.
+
+\textbf{Recording Session 3}: A PSL \cite{lsp_2015} conversation facilitated by the coordinator happening between the informant and the translator.\\
+
+During all the sessions a PSL \cite{lsp_2015} translator performs a translation after a word or phrase is completed.
+
+\subsection{Video Annotation Results}
+The video annotation pipeline described on \ref{videoannot} produced an annotated PSL dataset suitable for using it in a supervised learning experiment.
+The annotated dataset is divided in two main parts (RGB and Optical Flow samples)
+
+\textbf{RGB Samples folder structure} is a hierarchical folder structure where each detected noun or number is represented as a first level folder, all instances of a detected sign received an identifier which is an auto incremental integer, detected instances represent the second level folders, detected signs video represent the third level, Figure \ref{fig:rgb-two}.
+\begin{lstlisting}[caption=RGB Samples Folder Structure example, basicstyle=\ttfamily\small]
+L1:dos
+	L2:1
+		L3:rgb-frame01.jpg
+		L3:rgb-frame02.jpg
+		L3:rgb-frame03.jpg
+		...
+		L3:rgb-frame08.jpg
+L1:cine
+	L2:1
+		L3:rgb-frame01.jpg
+		L3:rgb-frame02.jpg
+		L3:rgb-frame03.jpg
+		...
+		L3:rgb-frame15.jpg
+	L2:2
+		L3:rgb-frame01.jpg
+		L3:rgb-frame02.jpg
+		L3:rgb-frame03.jpg
+		...
+		L3:rgb-frame11.jpg
+\end{lstlisting}\label{list:rgb-samples-folders}
+
+\begin{figure}[hbt!]
+\includegraphics[width=\linewidth]{images/dos-rgb-frames.png}
+\caption{PSL number "Two" RGB representation}
+\label{fig:rgb-two}
+\end{figure}
+
+\textbf{Optical Flow samples folder structure} is a hierarchical folder structure based on the RGB samples folder structure, it is a more simple based on two levels instead of three, the nature of Optical Flow of tracing movement between frames allow to produce a single image for each detected PSL element instance (listing \ref{list:opticalflow-samples-folders}), Figure \ref{fig:opticalflow-two} shows an example of an optical flow generated sample
+\begin{lstlisting}[caption=Optical Flow Samples Folder Structure example, basicstyle=\ttfamily\small]
+L1:dos
+	L2:oflow-dos-01.jpg
+L1:cine
+	L2:oflow-cine-01.jpg
+	L2:oflow-cine-02.jpg
+\end{lstlisting}\label{list:opticalflow-samples-folders}
+\begin{figure}[hbt!]
+\includegraphics[width=\linewidth]{images/dos-opticalflow.jpg}
+\caption{PSL number "Two" OpticalFlow representation}
+\label{fig:opticalflow-two}
+\end{figure}
+
+\subsection{nSDm results}
+
 \section{Conclusion}\label{conclusion}

### 2019-10-07 src/main/pre-processing/video/opticalflow_handler.py

@@ -26,7 +26,7 @@ class OpticalFlowHandler(Handler):
         color = [255, 0, 128]
 
         frame_index = 1
-        no_frames = len([name for name in os.listdir(rgb_sample_folder_path)])
+        no_frames = len([name for name in os.listdir(rgb_sample_folder_path) if name != '.DS_Store'])
         frame_file_name = os.path.join(rgb_sample_folder_path, 'frame' + str(frame_index) + '.jpg')
 
         # Take first frame and find corners in it

### 2019-10-28 article/article-tesis.tex
adding pose estimation state of art

@@
 \subsection{Pose Estimation}
+Pose estimation is also an extensively studied field.
+Techniques based on key points have shown state of art results on human pose estimation.
+An approach on key points estimation \cite{tulsianiMalik} uses \textit{Point of View Determination} and \textit{Key Points Prediction} components.
+Point of View Determination is formulated by the prediction of three Euler angles (azimut, elevation and cyclotation) generating a global position estimate, then a local appearance is modeled by obtaining a heat map that corresponds to the verisimilitude spatial distribution for each key point, finally key points predictions are obtained by combining heat maps obtained in a previous stage with a conditioned verisimilitude at the point of view predicted in the previous stage.
+
+Key points detection methods based CNNs have received an special attention in Human Pose Detection problems, CNNs methods are divided in bottom-up and top-down.
+Bottom-up methods process images from low resolution  to high resolution, focusing first on detecting joints and then before associating them to human actions.
+Top-down methods process focused first on detecting humans and then estimating the human pose to predict key points.
+
+The datasets MPII and COCO have been used in state of art methods obtaining good results\cite{XiaoWuWeiSimpleBaseline} and establishing a framework for future work in combination with classic approaches like optical flow for recognizing patterns movement between frames by increasing accuracy on key points detection.
 \subsection{Video Classification}

### 2019-11-20 article/article-tesis.tex
adding article updates

@@
 	\textbf{Talking with signs} \\	
-	\Large \textbf{A method for detecting signs in a weakly annotated dataset}
+	\Large \textbf{A method for progressively detect sings in a non annotated Peruvian Sings Language corpus}
 }
@@
 \begin{abstract}
-People with deafness or hearing disabilities who aim to use computer based systems, state of art video classification and human action recognition rely in a combination of traditional and new methods like optical flow and 3D CNN.
-In this article we show a novel method to detect Peruvian Sign Language elements (nSDm) from a weakly annotated videos dataset.
-Video annotation techniques were applied increasing the models performance, as a result we achieved XX precision and XX exhaustiveness by training a 3D CNN with optical flow inputs.
+People with deafness or hearing disabilities who aim to use computer based systems relies on state of art video classification and human action recognition techniques that combine traditional movement pattern recognition and deep learning techniques.
+In this work we present a pipeline for semi automatic video annotation applied to a non annotated \textit{Peruvian Signs Language} (PSL) corpus along with a novel method for a progressive detection of PSL elements (nSDm).
+We produced a set of video annotations indicating signs appearances for a small set of nouns and numbers along with a \textit{labeled PSL dataset} (PSL dataset).
+A composite model obtained from the combination of a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade optical flow, and a RNN with LSTM cells trained with raw rgb frames extracted from the PSL dataset achieved XX precision and XX exhaustivity on sings classification, reporting state of art results over the PSL dataset.
 \\
-\keywords{video classification,human actions detection, optical flow, 3D CNN.}
+\keywords{Video Classification, Human Actions Detection, Peruvian Signs Language, Optical Flow, 2D CNN, LSTM.}
 \end{abstract}
@@
 \section{Introduction}\label{intro}
+The World Health Organization (WHO) stated that 466 million people world wide have disabling hearing loss, estimating that by 2050 over 900 million people will have disabling hearing loss that will represent a global cost of 750 million dollars annually \cite{deafness_and_hearing_loss_2019}.
 
-The World Health Organization (WHO) stated that 466 million people world wide have disabling hearing loss, estimating that by 2050 over 900 million people will have disabling hearing loss and will represent a global cost of 750 million dollars annually \cite{deafness_and_hearing_loss_2019}.
+The Peruvian Institute of Informatics and Statistics (INEI) conducted a national disabilities survey with the objective of segmenting and acquire a better understanding about disabilities that affect the Peruvian population \cite{disabilities_survey_2012}, the results showed that 1.8\% of the Peruvian population suffer of at least partial when not permanent deafness or hearing limitations.
 
-The Peruvian Institute of Informatics and Statistics (INEI) conducted a national disabilities survey with the objective of segmenting and better understanding which disabilities affect the Peruvian population \cite{disabilities_survey_2012}, the results showed that 1.8\% of the Peruvian population suffer at least partial when not permanent deafness or hearing limitation.
+As of today, thanks to Deep Learning techniques for Computer Vision and Natural Language Processing, it is completely feasible to propose and design systems that are able to detect and transcribe sign languages elements improving in systems accessibility for people with physical limitations.
+Sings detection is usually a supervised learning task that requires a large and fully labeled dataset in order to achieve satisfactory results during model training and inference tasks.
+At the time  of writing this work there were no labeled data sets available for the PSL \cite{lsp_2015}, previous work was done by the \textit{Grammar and Signs research group of the Pontifical Catholic University of Peru (PUCP)} who created the first PSL corpus \cite{lsp_dataset}, this corpus is not annotated because there is no available metadata that describes signs being emitted and
 
-As of today, thanks to deep learning models for computer vision and natural language processing, it is completely feasible to propose systems able to detect and transcribe sign languages.
-However, deep learning supervised models usually require large and well annotated datasets, and developing such a dataset for the Peruvian Sign Language (PSL) \cite{lsp_2015} is expensive and difficult to achieve.
+  on annotating a Peruvian Sign Language (PSL) \cite{lsp_2015} a dataset, configuring an expensive and difficult task to achieve.
+The Pontifical Catholic University of Peru (PUCP) Grammar and  Signs research group created a non annotated PSL dataset \cite{lsp_dataset} because it does not show exactly the relation between the instant when a sign that is emitted and its corresponding translation.
+
 Furthermore, in the same way as spoken languages, signs languages also present local variations e.g.
 people who live in Lima metropolitan area are not expected to use the same set of signs as people leaving in other parts of the country.
-A need exists for approaching this task with self-supervised learning or by means of weakly labeled data.
+A need exists for approaching this task with self-supervised learning or by means of weakly or nonexistent labeled data.
 
-This article reports research conducted with the goal of proposing a method for detecting PSL signs using a weakly annotated dataset, and answering the following research questions:
+This work reports research conducted with the goal of proposing a method for detecting PSL elements using a not annotated dataset, and answering the following research questions:
 
 \begin{enumerate}[(i)]
-\item What are most relevant currently available techniques for training an AI model from  weakly annotated video and audio transcription?\label{q1}
-\item How precise and exhaustive is the model described in the above question on the recognition of a reduced set of signs in a weakly annotated dataset?\label{q2}
+\item What are most relevant and currently available techniques for training an AI model from not annotated videos and audio transcription?\label{q1}
+\item How precise and exhaustive is the model described in the above question on the recognition of a reduced set of signs in a not annotated dataset?\label{q2}
 \item What is the relation between the number of samples and the detection of new LSP elements in terms of nSDm precision and recall?\label{q3}
 \end{enumerate}
-
-The Pontifical Catholic University of Peru (PUCP) Grammar and  Signs research group created a PSL dataset \cite{lsp_dataset} that is weakly annotated because it does not show exactly the relation between the instant when a sign that is emitted and its corresponding translation.
 

### 2019-11-20 src/main/learning/configurers/base_configurer.py

@@ -1,0 +1,19 @@
+from tensorflow.keras.optimizers import SGD
+from tensorflow.keras.models import Model
+
+
+class BaseModelConfigurer:
+    def __init__(self, model: Model):
+        self.model = model
+
+    def configure(self):
+        self.model.compile(optimizer=SGD)
+
+    def get_optimizer(self):
+        return SGD()
+
+    def get_loss_function(self):
+        pass
+
+    def get_metrics(self):
+        pass

### 2019-11-20 src/main/learning/main.py

@@ -25,6 +25,19 @@ def get_rgb_model():
     return nn
 
 
+def get_cmd_args():
+    parser = ArgumentParser()
+
+    parser.add_argument('-n', '--model', help='Model Name', required=True)
+    parser.add_argument('-o', '--operation', help='Operation', default='train')
+
+    parser.add_argument('-bs', '--batch_size', help='Batch Size', default=10)
+    parser.add_argument('-nc', '--no_classes', help='Number of classes', default=10)
+    parser.add_argument('-ne', '--no_epochs', help='Number of epochs', default=20)
+
+    return parser.parse_args()
+
+
 if __name__ == '__main__':
     working_folder = os.getenv('WORK_DIR', './')
 
@@ -32,20 +45,8 @@ if __name__ == '__main__':
     setup_logging(working_folder, 'learning-logging.yaml')
     logger = logging.getLogger(__name__)
 
-    logger.debug("test")
-
     base_dataset_path = os.path.join(working_folder, 'dataset')
-
-    parser = ArgumentParser()
-
-    parser.add_argument('-n', '--network', help='Network Name', required=True)
-    parser.add_argument('-o', '--operation', help='Operation', default='train')
-
-    parser.add_argument('-bs', '--batch_size', help='Batch Size', default=10)
-    parser.add_argument('-nc', '--no_classes', help='Number of classes', default=10)
-    parser.add_argument('-ne', '--no_epochs', help='Number of epochs', default=20)
-
-    args = parser.parse_args()
+    args = get_cmd_args()
 
     kwargs = {
         'BatchSize': args.batch_size,
@@ -53,15 +54,15 @@ if __name__ == '__main__':
         'NoEpochs': args.no_epochs,
     }
 
-    networks = {
+    models = {
         'opticalflow': get_opticalflow_model,
         'rgb': get_rgb_model,
     }
-    network = networks[args.network]()
+    model = models[args.model]()
 
     operations = {
-        'train': network.train_network,
-        'evaluate': network.evaluate_network,
-        'predict': network.predict,
+        'train': model.train_network,
+        'evaluate': model.evaluate_network,
+        'predict': model.predict,
     }
     operations[args.operation]()

### 2019-11-20 src/main/learning/models/base_model.py

@@ -12,9 +12,6 @@ class BaseModel(tf.keras.Model):
         super(BaseModel, self).__init__()
 
         self._dataset_path = dataset_path
-        self._batch_size = kwargs['BatchSize']
-        self._no_classes = kwargs['NoClasses']
-        self._epochs = kwargs['NoEpochs']
         self._model = None
 
         self.data_folder = pathlib.Path(self._dataset_path)
@@ -23,7 +20,7 @@ class BaseModel(tf.keras.Model):
         pass
 
     def configure(self):
-        pass
+        self._model.compile()
 
     def get_class_names(self):
         pass

### 2019-11-22 article/article-tesis.tex
updates on the paper and initial work for learning models

@@
 	\textbf{Talking with signs} \\	
-	\Large \textbf{A method for progressively detect sings in a non annotated Peruvian Sings Language corpus}
+	\Large \textbf{A method for progressively detect signs in a non annotated signs language corpus}
 }
@@
 We produced a set of video annotations indicating signs appearances for a small set of nouns and numbers along with a \textit{labeled PSL dataset} (PSL dataset).
-A composite model obtained from the combination of a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade optical flow, and a RNN with LSTM cells trained with raw rgb frames extracted from the PSL dataset achieved XX precision and XX exhaustivity on sings classification, reporting state of art results over the PSL dataset.
+A composite model obtained from the combination of a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade optical flow, and a RNN with LSTM cells trained with raw rgb frames extracted from the PSL dataset achieved XX precision and XX exhaustivity on sings classification tasks, reporting state of art results over the PSL dataset.
 \\
@@
 
-As of today, thanks to Deep Learning techniques for Computer Vision and Natural Language Processing, it is completely feasible to propose and design systems that are able to detect and transcribe sign languages elements improving in systems accessibility for people with physical limitations.
+With today's Deep Learning techniques for Computer Vision and Natural Language Processing it is completely feasible to propose and design systems that are able to detect and transcribe sign languages elements improving in systems accessibility for people with physical limitations.
 Sings detection is usually a supervised learning task that requires a large and fully labeled dataset in order to achieve satisfactory results during model training and inference tasks.

### 2019-11-22 src/main/learning/configurers/base_configurer.py

@@ -1,19 +1,0 @@
-from tensorflow.keras.optimizers import SGD
-from tensorflow.keras.models import Model
-
-
-class BaseModelConfigurer:
-    def __init__(self, model: Model):
-        self.model = model
-
-    def configure(self):
-        self.model.compile(optimizer=SGD)
-
-    def get_optimizer(self):
-        return SGD()
-
-    def get_loss_function(self):
-        pass
-
-    def get_metrics(self):
-        pass

### 2019-11-22 src/main/learning/main.py

@@ -14,14 +14,12 @@ rgb_network_name = 'rgb'
 def get_opticalflow_model():
     dataset_path = os.path.join(base_dataset_path, 'opticalflow')
     nn = OpticalFlowModel(dataset_path, **kwargs)
-    nn.configure()
     return nn
 
 
 def get_rgb_model():
     dataset_path = os.path.join(base_dataset_path, 'rgb')
     nn = RGBModel(dataset_path, **kwargs)
-    nn.configure()
     return nn
 
 
@@ -61,8 +59,8 @@ if __name__ == '__main__':
     model = models[args.model]()
 
     operations = {
-        'train': model.train_network,
-        'evaluate': model.evaluate_network,
+        'train': model.train,
+        'evaluate': model.evaluate,
         'predict': model.predict,
     }
     operations[args.operation]()

### 2019-11-22 src/main/learning/models/base_model.py

@@ -1,26 +1,45 @@
-import tensorflow as tf
-import pathlib
+import logging
 import os
+import pathlib
+import tensorflow as tf
 
-from tensorflow.keras import layers
 from tensorflow.keras.models import Model
-from tensorflow.keras.applications.resnet_v2 import ResNet152V2
 
 
-class BaseModel(tf.keras.Model):
-    def __init__(self, dataset_path, **kwargs):
-        super(BaseModel, self).__init__()
+class BaseModel:
+    DEFAULT_NO_CLASSES = 10
+    DEFAULT_LEARNING_RATE = 0.001
+    DEFAULT_NO_EPOCHS = 50
 
-        self._dataset_path = dataset_path
-        self._model = None
+    SAVED_MODEL_FOLDER_NAME = 'saved-models/'
 
-        self.data_folder = pathlib.Path(self._dataset_path)
+    def __init__(self, working_folder, **kwargs):
+        self.working_folder = working_folder
 
-    def call(self, inputs, training=None, mask=None):
+        self.learning_rate = kwargs['LearningRate'] or self.DEFAULT_LEARNING_RATE
+        self.no_classes = kwargs['NoClasses'] or self.DEFAULT_NO_CLASSES
+        self.no_epochs = kwargs['NoEpochs'] or self.DEFAULT_NO_EPOCHS
+
+        self.model: Model = Model()  # using a default model if not assigned
+        self.model_history = None
+        self.pre_trained_model: Model
+
+        self.logger = logging.getLogger(__name__)
+
+    def get_model(self):
         pass
 
-    def configure(self):
-        self._model.compile()
+    def show_sample(self):
+        pass
+
+    def train(self):
+        pass
+
+    def evaluate(self):
+        pass
+
+    def predict(self):
+        pass
 
     def get_class_names(self):
         pass
@@ -28,14 +47,7 @@ class BaseModel(tf.keras.Model):
     def load_dataset(self):
         pass
 
-    def configure_model(self):
-        pass
-
-    def train_network(self):
-        self.configure_model()
-
-    def evaluate_network(self):
-        pass
-
-    def predict(self):
-        pass
+    def __save_model(self):
+        saved_model_path = os.path.join(self.working_folder, self.SAVED_MODEL_FOLDER_NAME)
+        tf.saved_model.save(self.model, saved_model_path)
+        self.logger.debug('model saved to %s', saved_model_path)

### 2019-11-22 src/main/learning/models/combined_model.py

@@ -1,0 +1,48 @@
+import logging
+import os
+import tensorflow as tf
+
+from base_model import BaseModel
+from opticalflow_model import OpticalFlowModel
+from rgb_model import RGBModel
+
+
+class CombinedModel(BaseModel):
+    OPTICALFLOW_MODEL_FOLDER_NAME = 'saved-models/opticalflow/'
+    RGB_MODEL_FOLDER_NAME = 'saved-models/rgb/'
+
+    def __init__(self, working_folder, **kwargs):
+        super(CombinedModel, self).__init__(working_folder, **kwargs)
+
+        self.opticalflow_model_path = os.path.join(self.working_folder, OpticalFlowModel.SAVED_MODEL_FOLDER_NAME)
+        self.rgb_model_path = os.path.join(self.working_folder, RGBModel.SAVED_MODEL_FOLDER_NAME)
+
+        self.logger = logging.getLogger(__name__)
+
+    def load_dataset(self):
+        pass
+
+    def get_model(self):
+        if not os.path.exists(self.opticalflow_model_path):
+            self.logger.error('the pre-trained opticalflow model does not exist at %s', self.opticalflow_model_path)
+
+        # load a pre-trained opticalflow model
+        opticalflow_model = tf.saved_model.load(self.opticalflow_model_path)
+
+        if not os.path.exists(self.rgb_model_path):
+            self.logger.error('the pre-trained rgb recurrent model does not exist at %s', self.rgb_model_path)
+
+        # load a pre-trained rgb recurrent model
+        rgb_model = tf.saved_model.load(self.rgb_model_path)
+
+    def train(self):
+        self.logger.debug('Opticalflow model train started with the following parameters')
+        self.model_history = self.model.fit()
+
+    def evaluate(self):
+        self.logger.debug('Opticalflow model evaluation started')
+        self.model.evaluate()
+
+    def predict(self):
+        self.logger.debug('Opticalflow model predict started')
+        self.model.predict()

### 2019-11-22 src/main/learning/models/opticalflow_model.py

@@ -1,22 +1,71 @@
-import pathlib
+import logging
 import tensorflow as tf
+import os
+
 from models.base_model import BaseModel
+from tensorflow.keras.applications.resnet_v2 import ResNet152V2
+from tensorflow.keras.models import Model
+from tensorflow.keras.layers import Dense, GlobalAveragePooling2D
+from tensorflow.keras import Input
+from tensorflow.keras.activations import sigmoid
+from tensorflow.keras.optimizers import Adam
+from tensorflow.keras.losses import CategoricalCrossentropy
+from tensorflow.keras.metrics import Recall, AUC, Precision
 
 
 class OpticalFlowModel(BaseModel):
-    def __init__(self, dataset_name, **kwargs):
-        super().__init__(dataset_name, **kwargs),
+    DEFAULT_NO_CHANNELS = 3  # Color images
+    DEFAULT_IMG_WIDTH = 224  # Imagenet default image width
+    DEFAULT_IMG_HEIGHT = 224  # Imagenet default image height
 
-    def configure(self):
+    SAVED_MODEL_FOLDER_NAME = 'saved-models/opticalflow/'
+
+    def __init__(self, working_folder, **kwargs):
+        super(OpticalFlowModel, self).__init__(working_folder, **kwargs)
+
+        self.no_channels = kwargs['NoChannels'] or self.DEFAULT_NO_CHANNELS
+        self.img_width = kwargs['ImageWidth'] or self.DEFAULT_IMG_WIDTH
+        self.img_height = kwargs['ImageHeight'] or self.DEFAULT_IMG_HEIGHT
+
+        # use a ResNet152 pre trained model as a base model, we are not including the top layer for maximizing the
+        # learned features
+        self.pre_trained_model: Model = ResNet152V2(include_top=False, weights='imagenet')
+
+        self.logger = logging.getLogger(__name__)
+
+    def load_dataset(self):
         print()
 
-    def get_class_names(self):
-        pass
-        # self.data_folder.glob("*").
+    def get_model(self):
+        # Freeze pre-trained base model weights
+        self.pre_trained_model.trainable = False
 
-    def configure_model(self):
-        self._model = tf.keras.Model()
-        print()
+        # Opticalflow model definition
+        inputs = Input(shape=(self.img_width, self.img_height, self.no_channels), name='inputs')
+        x = self.pre_trained_model.trainable(inputs)
[…]

### 2019-11-22 src/main/learning/models/rgb_model.py

@@ -1,15 +1,33 @@
-import tensorflow as tf
+import logging
+
 from models.base_model import BaseModel
 
 
 class RGBModel(BaseModel):
-    def __init__(self, dataset_name, **kwargs):
-        pass
+    SAVED_MODEL_FOLDER_NAME = 'saved-models/rgb/'
 
-    def get_dataset(self):
-        ds = tf.data.Dataset
+    def __init__(self, working_folder, **kwargs):
+        super().__init__(working_folder, **kwargs)
 
+        self.logger = logging.getLogger(__name__)
+
+    def load_dataset(self):
         pass
 
     def get_model(self):
         pass
+
+    def train(self):
+        self.logger.debug('Opticalflow model train started with the following parameters')
+        self.model_history = self.model.fit()
+
+        # save trained model and weights to the file system for future use
+        self.__save_model()
+
+    def evaluate(self):
+        self.logger.debug('Opticalflow model evaluation started')
+        self.model.evaluate()
+
+    def predict(self):
+        self.logger.debug('Opticalflow model predict started')
+        self.model.predict()

### 2019-11-23 src/main/learning/main.py
added better support for tf.dataset

@@ -7,19 +7,22 @@ from argparse import ArgumentParser
 
 from logger_config import setup_logging
 
+DEFAULT_NO_CLASSES = 10
+DEFAULT_LEARNING_RATE = 0.001
+DEFAULT_NO_EPOCHS = 50
+DEFAULT_BATCH_SIZE = 10
+
 optical_flow_network_name = 'opticalflow'
 rgb_network_name = 'rgb'
 
 
 def get_opticalflow_model():
-    dataset_path = os.path.join(base_dataset_path, 'opticalflow')
-    nn = OpticalFlowModel(dataset_path, **kwargs)
+    nn = OpticalFlowModel(working_folder, dataset_path, **kwargs)
     return nn
 
 
 def get_rgb_model():
-    dataset_path = os.path.join(base_dataset_path, 'rgb')
-    nn = RGBModel(dataset_path, **kwargs)
+    nn = RGBModel(working_folder, dataset_path, **kwargs)
     return nn
 
 
@@ -28,10 +31,10 @@ def get_cmd_args():
 
     parser.add_argument('-n', '--model', help='Model Name', required=True)
     parser.add_argument('-o', '--operation', help='Operation', default='train')
-
-    parser.add_argument('-bs', '--batch_size', help='Batch Size', default=10)
-    parser.add_argument('-nc', '--no_classes', help='Number of classes', default=10)
-    parser.add_argument('-ne', '--no_epochs', help='Number of epochs', default=20)
+    parser.add_argument('-ds', '--dataset_path', help='Dataset Path', required=True)
+    parser.add_argument('-bs', '--batch_size', help='Batch Size', default=DEFAULT_BATCH_SIZE)
+    parser.add_argument('-ne', '--no_epochs', help='Number of epochs', default=DEFAULT_NO_EPOCHS)
+    parser.add_argument('-lr', '--learning_rate', help='Leaning Rate', default=DEFAULT_LEARNING_RATE)
 
     return parser.parse_args()
 
@@ -43,13 +46,13 @@ if __name__ == '__main__':
     setup_logging(working_folder, 'learning-logging.yaml')
     logger = logging.getLogger(__name__)
 
-    base_dataset_path = os.path.join(working_folder, 'dataset')
     args = get_cmd_args()
 
+    dataset_path = args.dataset_path
     kwargs = {
         'BatchSize': args.batch_size,
-        'NoClasses': args.no_classes,
         'NoEpochs': args.no_epochs,
+        'LearningRate': args.learning_rate
     }
 
     models = {

### 2019-11-23 src/main/learning/models/opticalflow_model.py

@@ -3,6 +3,7 @@ import tensorflow as tf
 import os
 
 from models.base_model import BaseModel
+
 from tensorflow.keras.applications.resnet_v2 import ResNet152V2
 from tensorflow.keras.models import Model
 from tensorflow.keras.layers import Dense, GlobalAveragePooling2D
@@ -14,49 +15,54 @@ from tensorflow.keras.metrics import Recall, AUC, Precision
 
 
 class OpticalFlowModel(BaseModel):
-    DEFAULT_NO_CHANNELS = 3  # Color images
-    DEFAULT_IMG_WIDTH = 224  # Imagenet default image width
-    DEFAULT_IMG_HEIGHT = 224  # Imagenet default image height
-
     SAVED_MODEL_FOLDER_NAME = 'saved-models/opticalflow/'
 
-    def __init__(self, working_folder, **kwargs):
-        super(OpticalFlowModel, self).__init__(working_folder, **kwargs)
-
-        self.no_channels = kwargs['NoChannels'] or self.DEFAULT_NO_CHANNELS
-        self.img_width = kwargs['ImageWidth'] or self.DEFAULT_IMG_WIDTH
-        self.img_height = kwargs['ImageHeight'] or self.DEFAULT_IMG_HEIGHT
-
-        # use a ResNet152 pre trained model as a base model, we are not including the top layer for maximizing the
-        # learned features
-        self.pre_trained_model: Model = ResNet152V2(include_top=False, weights='imagenet')
+    def __init__(self, working_folder, dataset_path, **kwargs):
+        super(OpticalFlowModel, self).__init__(working_folder, dataset_path, **kwargs)
 
         self.logger = logging.getLogger(__name__)
 
-    def load_dataset(self):
-        print()
+        self.dataset: tf.data.Dataset = self.get_dataset()
+        self.model: Model = self.get_model()
+        self.model_history = None
 
-    def get_model(self):
+    def get_model(self) -> Model:
+        # use a ResNet152 pre trained model as a base model, we are not including the top layer for maximizing the
+        # learned features
+        pre_trained_model: Model = ResNet152V2(include_top=False, weights='imagenet')
         # Freeze pre-trained base model weights
-        self.pre_trained_model.trainable = False
+        pre_trained_model.trainable = False
+
+        no_classes = len(self.classes)
 
         # Opticalflow model definition
         inputs = Input(shape=(self.img_width, self.img_height, self.no_channels), name='inputs')
[…]

### 2019-11-26 src/main/learning/models/opticalflow_model.py
updated the article to include model architecture diagrams

@@ -59,7 +59,7 @@ class OpticalFlowModel(BaseModel):
 
         # Opticalflow model assembling
         model = Model(inputs=inputs, outputs=outputs, name='OpticalflowModel')
-        self.logger.debug('Opticalflow model summary: \n %s', model.summary())
+        # self.logger.debug('Opticalflow model summary: \n %s', model.summary())
 
         # Opticalflow model compilation
         optimizer = Adam(learning_rate)

### 2019-11-26 src/main/learning/models/rgb_model.py

@@ -1,26 +1,33 @@
 import logging
+import tensorflow as tf
 
 from models.base_model import BaseModel
+from tensorflow.keras.models import Model
+from tensorflow.keras.optimizers import Adam
+from tensorflow.keras.losses import CategoricalCrossentropy
+from tensorflow.keras.metrics import Recall, AUC, Precision
 
 
 class RGBModel(BaseModel):
     SAVED_MODEL_FOLDER_NAME = 'saved-models/rgb/'
 
-    def __init__(self, working_folder, dataset_path, **kwargs):
-        super().__init__(working_folder, dataset_path, **kwargs)
+    def __init__(self, working_folder, dataset_root_path):
+        super(RGBModel, self).__init__(working_folder, dataset_root_path)
         self.logger = logging.getLogger(__name__)
 
-    def train(self):
-        self.logger.debug('Opticalflow model train started with the following parameters')
-        self.model_history = self.model.fit()
+    def get_dataset(self, dataset_path, **kwargs) -> tf.data.Dataset:
+        pass
 
-        # save trained model and weights to the file system for future use
-        self.__save_model()
+    def get_model(self, **kwargs) -> Model:
+        img_width = kwargs['ImageWidth']
+        img_height = kwargs['ImageHeight']
+        no_channels = kwargs['NoChannels']
+        learning_rate = kwargs['LearningRate']
 
-    def evaluate(self):
-        self.logger.debug('Opticalflow model evaluation started')
-        self.model.evaluate()
+        model = Model()
 
-    def predict(self):
-        self.logger.debug('Opticalflow model predict started')
-        self.model.predict()
+        optimizer = Adam(learning_rate)
+        model.compile(optimizer=optimizer, loss=CategoricalCrossentropy(),
+                      metrics=[Recall(), AUC(curve='PR'), Precision()])
+
+        return model

### 2019-12-01 article/article-tesis.tex
updated the article to include model architecture diagrams
input pipeline for nsdm

@@
 	\textbf{Talking with signs} \\	
-	\Large \textbf{A method for progressively detect signs in a non annotated signs language corpus}
+	\Large \textbf{A simple method to detect nouns and numbers in a non annotated signs language corpus}
 }
@@
 
-The Peruvian Institute of Informatics and Statistics (INEI) conducted a national disabilities survey with the objective of segmenting and acquire a better understanding about disabilities that affect the Peruvian population \cite{disabilities_survey_2012}, the results showed that 1.8\% of the Peruvian population suffer of at least partial when not permanent deafness or hearing limitations.
+The Peruvian Institute of Informatics and Statistics (INEI) conducted a national disabilities survey with the objective of segmenting and acquire a better understanding about disabilities that affect the Peruvian population \cite{disabilities_survey_2012}, the results showed that 1.8\% of the Peruvian population suffer at least partial when not permanent deafness or hearing limitations.
 
-With today's Deep Learning techniques for Computer Vision and Natural Language Processing it is completely feasible to propose and design systems that are able to detect and transcribe sign languages elements improving in systems accessibility for people with physical limitations.
-Sings detection is usually a supervised learning task that requires a large and fully labeled dataset in order to achieve satisfactory results during model training and inference tasks.
-At the time  of writing this work there were no labeled data sets available for the PSL \cite{lsp_2015}, previous work was done by the \textit{Grammar and Signs research group of the Pontifical Catholic University of Peru (PUCP)} who created the first PSL corpus \cite{lsp_dataset}, this corpus is not annotated because there is no available metadata that describes signs being emitted and
+Peruvians with deafness or hearing limitations use the Peruvian Signs Language (PSL) as their main communication medium.
+PSL is of mandatory usage at universities and certain public institutions, here the importance of designing systems that are capable to support PSL inputs and outputs.
+Furthermore, in the same way as spoken languages, signs languages also present local variations e.g.
+people who live in Lima metropolitan area are not expected to use the same set of signs as people in other parts of the territory.
+This work uses the PSL variation used in Lima due to the difficulty or inability to find datasets for other PSL variations.
 
-  on annotating a Peruvian Sign Language (PSL) \cite{lsp_2015} a dataset, configuring an expensive and difficult task to achieve.
-The Pontifical Catholic University of Peru (PUCP) Grammar and  Signs research group created a non annotated PSL dataset \cite{lsp_dataset} because it does not show exactly the relation between the instant when a sign that is emitted and its corresponding translation.
+\textit{The Grammar and Signs research group of the Pontifical Catholic University of Peru (PUCP)} built the first PSL corpus \cite{lsp_dataset} which is publicly available at the university digital archives, it is important to highlight that the corpus is neither labeled or annotated and cannot be used as it is for training or testing a model.
 
-Furthermore, in the same way as spoken languages, signs languages also present local variations e.g.
-people who live in Lima metropolitan area are not expected to use the same set of signs as people leaving in other parts of the country.
-A need exists for approaching this task with self-supervised learning or by means of weakly or nonexistent labeled data.
+In this work we are approaching signs detection as a supervised learning task.
+Supervised learning requires labeled dataset to achieve satisfactory results during training and inference tasks.
+At the time of writing this work there were no labeled datasets available for PSL \cite{lsp_2015}, it configures a gap that could prevent or hinder research work on Human-Computer-Interaction at the Peruvian or Latin American space.
 
-This work reports research conducted with the goal of proposing a method for detecting PSL elements using a not annotated dataset, and answering the following research questions:
+Current advances in Computer Vision (CV) and Natural Language Processing (NLP) make it possible to conceive systems that are capable of detecting and transcribing elements of sign languages thereby improving systems accessibility for people with physical limitations.
+This work reports results of a research conducted with the goal of producing a labeled PSL dataset for a set of signs limited to nouns and numbers as well as a novel method for detecting PSL signs by answering the following research questions:
 
 \begin{enumerate}[(i)]
-\item What are most relevant and currently available techniques for training an AI model from not annotated videos and audio transcription?\label{q1}
-\item How precise and exhaustive is the model described in the above question on the recognition of a reduced set of signs in a not annotated dataset?\label{q2}
-\item What is the relation between the number of samples and the detection of new LSP elements in terms of nSDm precision and recall?\label{q3}
+\item What are the currently available techniques for producing a labeled dataset for a set of signs limited to nouns and number from the non annotated PSL corpus?\label{q1}
+\item What are most relevant and currently available techniques for training a model with the labeled dataset described in the question above for detecting PSL nouns and numbers?\label{q2}
+\item How precise and exhaustive is the model described in the above question on the detection of PSL nouns and numbers?\label{q3}
 \end{enumerate}
 
-nSDm will take advantage of the PSL dataset \cite{lsp_dataset} and will contribute with the generalization and progressive signs detection that could be used as a starting point for future studies including software accessibility improvements and human computer interaction for people with deafness and hearing limitations.
+This work has the main objective of producing a simple method that can be used as a baseline for other researchers interested on studying signs language and their different applications on the Human-Computer-Interaction field, we believe this work will produce a positive impact on the artificial intelligence community towards to an increase in the number of research works using PSL which can contribute to increasing the number of people with deafness or hearing disabilities that can use computer based systems.
+We have divided this work main objective into following specific objectives for better traceability:
+
+\begin{enumerate}[(i)]
+\item Produce a labeled PSL dataset limited to nouns and number\label{o1}
+\item Design a train a novel signs detection model (nSDm) for detecting signs at the labeled PSL dataset\label{o2}
+\item Determine what is the performance of nSDm in terms of precision and recall\label{o3}
+\end{enumerate}
 
 The rest of the article is organized as follows.
-In sec \ref{relatedwork} we review the related work on video classification and body actions recognition using both Optical Flow+Two Stream CNN and 3D CNN architectures.
-In sec \ref{method} we introduce nSDm and specify its architecture and design.
-In sec \ref{experimentation} we evaluate nSDm precision and recall and provide answers for the research questions \ref{q1},\ref{q2},\ref{q3}.
-In sec \ref{datasetdesc} we describe and provide the dataset details.
-In sec \ref{videoannot} we describe the video annotation and data pre-processing techniques applied to the dataset and finally in sec \ref{conclusion} we present our conclusion and describe future work.
+In section \ref{relatedwork} we review the related work on video classification for human actions recognition using network architectures that combine CNNs, 3D CNNs and movement patterns  for better features learning, we also review state of art pose estimation techniques.
+In section \ref{method} we introduce nSDm describing its design and architecture.
+In section \ref{experimentation} we evaluate nSDm precision and recall and answer research questions \ref{q1},\ref{q2},\ref{q3}.
+In section \ref{datasetdesc} we describe the PSL dataset produced at PUCP.
+In section \ref{videoannot} we describe the video annotation and data pre-processing techniques applied to produce the labeled PSL dataset and finally in section \ref{conclusion} we present our conclusions and future work.
 \section{Related Work} \label{relatedwork}
@@
 \subsection{Video Annotation}\label{videoannot}
-The PSL dataset is weakly annotated because there is not a direct relation between the instant when a sign is emitted and when its translation to Spanish is delivered, nSDm proposes a semi automatic video annotation pipeline described in Figure \ref{fig:video-annotation-pipeline} for cleaning, pre-processing and analyzing PSL videos in order to produce an annotated PSL dataset that can be used for training the nSDm model using supervised learning.
-The pipeline is described in detail in \ref{manual-video-cleanup}, \ref{video-pre-processing}, \ref{audio-transcription-analysis} and \ref{samples-generation}
+The PSL dataset is non annotated because there is not a direct relation between the instant when a sign is emitted and when its translation to Spanish is delivered, we propose a semi automatic video annotation pipeline described in Figure \ref{fig:video-annotation-pipeline} for cleaning, pre-processing and analyzing PSL videos in order to produce an labeled PSL dataset that can be used for training nSDm using supervised learning.
+The pipeline is described in detail in sections  \ref{manual-video-cleanup}, \ref{video-pre-processing}, \ref{audio-transcription-analysis} and \ref{samples-generation}
 
@@
 \subsubsection{Manual Automatic Video Clean Up}\label{manual-video-cleanup}
-The PSL recordings described on \ref{datasetdesc} contain a considerable amount of noise introduced during the recording sessions which makes difficult to easily find video intervals that clearly show a relation between signs emitted by the informant and the translation delivered by the translator, the noise factors are listed in the following list:
+The PSL recordings described on \ref{datasetdesc} contain a considerable amount of noise introduced during recording sessions.
+It makes difficult to easily find video intervals that clearly show a relation between signs emitted by the informant and the translation delivered by the translator, the noise factors are listed in the following list:
 \begin{itemize}
@@
 	\item Conversations between participants that are not relevant to emitted sings.
-	\item High frequency of silent periods.
+	\item High frequency of large silent periods.
 \end{itemize}
-A manual video cleanup process is required to find noise free video intervals, it is a manual process that requires watching all the videos available in the PSL dataset and manually annotating the instant when an informant started emitting sings and the instant when the translator completed delivering a translation, table \ref{table:noise-free-video-segments} shows an annotation example.
+A manual video cleanup process is required to find noise free video intervals, it is a process that requires watching all the videos available at the PSL corpus for manually annotating the instant when an informant started emitting sings along with the instant when the translator delivered a translation, table \ref{table:noise-free-video-segments} shows a manual annotation example.
 
-The recordings show the informant in two alignments (centered and left), the manual video clean up process also stores the informant alignment, table \ref{table:informant-alignment} shows the two available alignments, we use the alignment annotation later in the process during the video frames extraction to create the PSL annotated dataset.
+The recordings show the informant in two alignments (centered and left), the manual video clean up process also stores the informant alignment, table \ref{table:informant-alignment} shows the two available alignments, we use the alignment annotation later in the process during the video frames extraction to create the labeled PSL dataset.
 
@@
 \subsubsection{Video Pre-Processing}\label{video-pre-processing}
-Weakly PSL annotated videos require processing before any metadata can be extracted, we propose a sequence of pre-processing tasks that take advantage of the annotation generated on \ref{manual-video-cleanup}.
-A \textbf{video splitting processor} generates a set of video chunks using the \textit{ffmpeg} multimedia framework and stores the produced chunks in Amazon S3 for later usage.
-The audio within the video chunks is then transcribed by an \textbf{audio transcription processor}, using the Amazon Transcription service, we selected the Amazon Transcription service because it provides an accurate mapping between audio participants and transcribed words along with useful metadata that describes the start and end time when words are pronounced.
+Non annotated PSL videos require processing before any metadata can be extracted, we propose a sequence of pre-processing tasks that take advantage of the annotation generated on \ref{manual-video-cleanup}.
+A \textbf{video splitting processor} generates a set of video chunks using the \textit{ffmpeg} multimedia framework and stores produced video chunks in Amazon S3 for later usage.
+Audio within video chunks is then transcribed by an \textbf{audio transcription processor}, using the Amazon Transcription service, we selected the Amazon Transcription service because it provides an accurate mapping between audio participants and transcribed words along with useful metadata that describes the start and end time when words are pronounced by the translator.
 
-At the moment of writing this work Amazon Transcription service only supported Spanish for Spain and the US, it caused certain words that are specific for Peruvian Spanish not being fully recognized, in order to improve transcription accuracy we provided a custom vocabulary containing Peruvian expressions which improved  Peruvian words recognition, for the matters of this work unrecognized Peruvian words were omitted and not processed.
+At the moment of writing this work Amazon Transcription service only supported Spain and US Spanish, it caused certain words that are specific for Peruvian Spanish not being fully recognized, in order to improve transcription accuracy we built a custom vocabulary containing Peruvian expressions which improved  Peruvian words recognition, for the matters of this work Peruvian words that remained unrecognized were omitted and not processed.
 
@@
 Bag of Embedding Words (BoEW) is a widely used technique on Natural Language Processing tasks providing a easy and flexible way to list the most relevant words based on frequency.
-This work is focused on detecting nouns and numbers (nSDm is a ongoing method that will be progressively improved to handle a wider set of PSL elements) assuming that nouns (numbers are a subset of nouns) suffer less variations in spoken Spanish than verbs, pronouns, adverbs and adjectives, and provide more semantic value than conjunctions, prepositions and interjections.
+This work is focused on detecting nouns and numbers (our method is designed to be progressively improved to handle a wider set of PSL elements) assuming that nouns (numbers are a subset of nouns) suffer less variations in spoken Spanish than verbs, pronouns, adverbs and adjectives, and provide more semantic value than conjunctions, prepositions and interjections.
 
@@
 \subsubsection{Samples Generation}\label{samples-generation} 
-nSDm requires PSL elements to be represented as a set of RGB frames and a calculated Optical Flow using the Lucas-Kanade method, both representations are inputs of two different models as presented on \ref{nsdm}.
+Our method requires PSL elements to be represented as a set of RGB frames and a calculated Optical Flow using the Lucas-Kanade method, both representations are inputs of two different models as presented on \ref{nsdm}.
 
-\textbf{Translation Delay Factor: } The difference in time between the instant when a sign is emitted and when a translation for that given sign is delivered is uncertain, we are calling that uncertainty as the \textit{translation delay factor} which we are trying to approximate using a constant value, we have chosen a 3 seconds translation delay factor assuming that most of the translations will occur between three seconds after a sign is emitted.
+\textbf{Translation Delay Factor: } The difference in time between the instant when a sign is emitted and when a translation for that given sign is delivered is uncertain, we are calling that uncertainty \textit{the translation delay factor}, we are trying to approximate it using a constant value, we chose a three seconds translation delay factor assuming that most of the translations will occur between three seconds after a sign is emitted.
 
 A \textbf{RGB Samples generation processor} uses samples metadata in combination with the translation delay factor to determine frames that represent a given PSL element.
-We use \textit{OpenCV} to extract frames and store them following a hierarchical folder structure (listing \ref{list:rgb-samples-folders}) that nSDm data loaders will use to feed data into the RGM branch on the nSDm model architecture during training and testing.
+We use \textit{OpenCV} to extract frames and store them following a hierarchical folder structure (listing \ref{list:rgb-samples-folders}) that nSDm data loaders will use to feed data into the RGB branch in the nSDm model architecture \ref{nsdm-architecture} during training and testing.
 
-A \textbf{Optical Flow Samples generation processor} uses video frames and the hierarchical folder structure generated by the RGM samples generation processor to calculate an Optical Flow representation for PSL elements and store them in a hierarchical folder structure that will also used by the nSDm data loaders to feed the optical flow branch on the nSDm model architecture during training and testing.
-We selected optical flow as a samples generation strategy because its ability to represent movement traces from previous frames, it is particular useful for representing PSL elements, a PSL element is composed on different body movements including: elbow, arms, neck, eyes, shoulders and hands, which are performed quickly, a way to detect movement traces between frames allows to generate a single image representation of all movement involved on a sign emitting (Figure \ref{fig:opticalflow-two}).
+A \textbf{Optical Flow Samples generation processor} uses video frames and the hierarchical folder structure generated by the RGB samples generation processor to calculate an Optical Flow representation for PSL elements and store them in a hierarchical folder structure that will also used by the nSDm data loaders to feed the optical flow branch on the nSDm model architecture \ref{nsdm-architecture} during training and testing.
+We selected optical flow as a samples generation strategy because its ability to represent movement traces from previous frames, it is particular useful for representing body movement patterns executed by informant while emitting a PSL sign, a PSL sign is made up of different body movements including: elbow, arms, neck, eyes, shoulders and hands, which are performed quickly, a way to detect movement traces between frames allows to generate a single image representation of all movement involved on a sign emitting (Figure \ref{fig:opticalflow-two}).
 
@@
 
-\subsection{Novel Signs Detection Method (nSDm)}\label{nsdm}
+\subsection{Novel Signs Detection Model (nSDm)}\label{nsdm}
+We propose a novel model for signs detection that combine two neural networks architectures, a CNN and a RNN with the objective to learn visual features like edges, corners and ridges (CNN) and at the same time patterns learned from a time based series of inputs (RNN).
+CNN network receives optical flow inputs and the RNN branch receives RGB frames extracted from the labeled PSL dataset described in \ref{videoannot}.
+
+For this work we selected the Tensorflow/Keras functional API for its ability to define combined models along with a versatile data extraction and transformation layer.
+More details and usage are provided in sections \ref{nsdm-architecture} and \ref{new-videos-testing}
+
+Implementation details can be found at https://github.com/erichuizapucp/signs-recognition
 \subsubsection{Architecture}\label{nsdm-architecture}
+nSDm is composed of several layers \ref{video-annotation-results}\\
+\textbf{Samples feeding layer:} A custom layer with the responsibility direct PSL samples \\
+\textbf{Optical flow model layer:} \\
+\textbf{Recurrent RGB model layer:} \\
+\textbf{Concatenation layer:} \\
+\textbf{Fully connected layer:} \\
+
 \begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/network-architecture.png}
-\caption{nSDm network architecture}
+\includegraphics[width=\linewidth]{images/nsdm-model-architecture.png}
+\caption{nSDm model architecture}
 \label{fig:opticalflow-two}
 \end{figure}
-\subsection{New Videos Testing}
+
+\begin{table}[!htb]
+\captionsetup{size=footnotesize}
+\begin{tabular}{ p{16em} c}
+\toprule
+\multicolumn{1}{c}{\textbf{Hyper parameter}} & 
+	\multicolumn{1}{c}{\textbf{Value}}\\
+\midrule
+\textbf{Learning Rate}&	0.001\\
+\textbf{No Epochs}&	5\\
+\textbf{No Steps per Epoch}&	5\\
+\textbf{Batch Size}&	10\\
+\textbf{Shuffle Buffer Size}&	50\\
+\bottomrule
+\end{tabular}
+\caption{nSDm training hyper parameters} \label{tab:nSDm-hyperparams}
+\end{table}
+
+\begin{figure}[hbt!]
+\includegraphics[width=\linewidth]{images/opticalflow-model-architecture.png}
+\caption{Opticalflow model architecture}
+\label{fig:opticalflow-two}
+\end{figure}
+
+\begin{figure}[hbt!]
+\includegraphics[width=\linewidth]{images/recurrent-rgb-model-architecture.png}
+\caption{Recurrent RGB model architecture}
+\label{fig:opticalflow-two}
+\end{figure}
+
+\subsection{Sign detection testing}\label{new-videos-testing}
+
+\begin{figure}[hbt!]
+\includegraphics[width=\linewidth]{images/new-video-test-architecture.png}
+\caption{Signs detection testing process architecture}
+\label{fig:opticalflow-two}
+\end{figure}
 
@@
 
-\subsection{Video Annotation Results}
+\subsection{Video Annotation Results}\label{video-annotation-results}
 The video annotation pipeline described on \ref{videoannot} produced an annotated PSL dataset suitable for using it in a supervised learning experiment.
@@
 
-\subsection{nSDm results}
+\subsection{Sign detection results}
 

### 2019-12-01 src/main/learning/main.py

@@ -1,8 +1,10 @@
 import logging
 import os
+import tensorflow as tf
 
 from learning.models.opticalflow_model import OpticalFlowModel
-from learning.models.rgb_model import RGBModel
+from learning.models.rgb_recurrent_model import RGBRecurrentModel
+from learning.models.novel_signs_detection_model import NovelSignsDetectionModel
 from argparse import ArgumentParser
 
 from logger_config import setup_logging
@@ -52,7 +54,8 @@ def get_train_params(args):
 def get_model(working_folder, dataset_root_path, model_name):
     models = {
         'opticalflow': lambda: OpticalFlowModel(working_folder, dataset_root_path),
-        'rgb': lambda: RGBModel(working_folder, dataset_root_path),
+        'rgb': lambda: RGBRecurrentModel(working_folder, dataset_root_path),
+        'nsdm': lambda: NovelSignsDetectionModel(working_folder, dataset_root_path)
     }
     model = models[model_name]()
     return model

### 2019-12-01 src/main/learning/models/base_model.py

@@ -69,7 +69,7 @@ class BaseModel:
 
     def get_labels(self, train=True) -> list:
         dataset_folder_name = self.TRAIN_DATASET_FOLDER_NAME if train else self.TEST_DATASET_FOLDER_NAME
-        path = self.get_dataset_path(dataset_folder_name)
+        path = os.path.join(self.get_dataset_path(dataset_folder_name), 'opticalflow')
         return [label for label in os.listdir(path) if os.path.isdir(os.path.join(path, label))]
 
     def get_saved_model(self) -> Model:
@@ -83,7 +83,7 @@ class BaseModel:
         self.logger.debug('model saved to %s', saved_model_path)
 
     def get_dataset_path(self, operation_folder_name):
-        dataset_path = os.path.join(self.dataset_root_path, operation_folder_name, self.MODEL_NAME)
+        dataset_path = os.path.join(self.dataset_root_path, operation_folder_name)
 
         if not os.path.exists(dataset_path):
             self.logger.error('The provided dataset path %s does not exist.', dataset_path)

### 2019-12-01 src/main/learning/models/combined_model.py

@@ -1,45 +1,0 @@
-import logging
-import os
-import tensorflow as tf
-
-from base_model import BaseModel
-from opticalflow_model import OpticalFlowModel
-from rgb_model import RGBModel
-
-
-class CombinedModel(BaseModel):
-    OPTICALFLOW_MODEL_FOLDER_NAME = 'saved-models/opticalflow/'
-    RGB_MODEL_FOLDER_NAME = 'saved-models/rgb/'
-
-    def __init__(self, working_folder, dataset_path, **kwargs):
-        super(CombinedModel, self).__init__(working_folder, dataset_path, **kwargs)
-
-        self.opticalflow_model_path = os.path.join(self.working_folder, OpticalFlowModel.SAVED_MODEL_FOLDER_NAME)
-        self.rgb_model_path = os.path.join(self.working_folder, RGBModel.SAVED_MODEL_FOLDER_NAME)
-
-        self.logger = logging.getLogger(__name__)
-
-    def __get_model(self):
-        if not os.path.exists(self.opticalflow_model_path):
-            self.logger.error('the pre-trained opticalflow model does not exist at %s', self.opticalflow_model_path)
-
-        # load a pre-trained opticalflow model
-        opticalflow_model = tf.saved_model.load(self.opticalflow_model_path)
-
-        if not os.path.exists(self.rgb_model_path):
-            self.logger.error('the pre-trained rgb recurrent model does not exist at %s', self.rgb_model_path)
-
-        # load a pre-trained rgb recurrent model
-        rgb_model = tf.saved_model.load(self.rgb_model_path)
-
-    def train(self):
-        self.logger.debug('Opticalflow model train started with the following parameters')
-        self.model_history = self.model.fit()
-
-    def evaluate(self):
-        self.logger.debug('Opticalflow model evaluation started')
-        self.model.evaluate()
-
-    def predict(self):
-        self.logger.debug('Opticalflow model predict started')
-        self.model.predict()

### 2019-12-01 src/main/learning/models/novel_signs_detection_model.py

@@ -1,0 +1,137 @@
+import logging
+import os
+
+import tensorflow as tf
+import matplotlib.pyplot as plt
+
+from tensorflow.keras.models import Model
+from tensorflow import feature_column
+
+from base_model import BaseModel
+from opticalflow_model import OpticalFlowModel
+from rgb_recurrent_model import RGBRecurrentModel
+from wrapper_layer import WrapperLayer
+from tensorflow.keras import Input
+from tensorflow.keras.layers import Dense
+
+from tensorflow.keras.optimizers import Adam
+from tensorflow.keras.losses import CategoricalCrossentropy
+from tensorflow.keras.metrics import Recall, AUC, Precision
+
+
+class NovelSignsDetectionModel(BaseModel):
+    MODEL_NAME = 'nsdm'
+
+    def __init__(self, working_folder, dataset_root_path):
+        super(NovelSignsDetectionModel, self).__init__(working_folder, dataset_root_path)
+        self.logger = logging.getLogger(__name__)
+
+        # self.opticalflow_model_path = os.path.join(self.working_folder, OpticalFlowModel.SAVED_MODEL_FOLDER_NAME)
+        # self.rgb_model_path = os.path.join(self.working_folder, RGBModel.SAVED_MODEL_FOLDER_NAME)
+
+    def get_dataset(self, dataset_path, training=True, **kwargs) -> tf.data.Dataset:
+        # Obtain opticalflow images and class
+        # from the class from optical flow look for rgb frames using the class and
+
+        batch_size = kwargs['BatchSize']
+        shuffle_buffer_size = kwargs['ShuffleBufferSize']
+        img_width = kwargs['ImageWidth']
+        img_height = kwargs['ImageHeight']
+        no_channels = kwargs['NoChannels']
+
+        opticalflow_dataset_path = os.path.join(dataset_path, OpticalFlowModel.MODEL_NAME)
+        rgb_dataset_path = os.path.join(dataset_path, RGBRecurrentModel.MODEL_NAME)
+
+        # First we will need only opticalflow samples file paths
+        dataset = tf.data.Dataset.list_files(opticalflow_dataset_path + '/*/*'). \
+            map(lambda x: self.samples_processing(x, rgb_dataset_path, no_channels, img_width, img_height))
+
+        # for X, label in dataset.take(1):
+        #     print(X.numpy().shape)
+        #     print(X.numpy())
+        #     print(label.numpy())
+
+        return dataset
+
[…]

### 2019-12-01 src/main/learning/models/rgb_recurrent_model.py

@@ -8,11 +8,11 @@ from tensorflow.keras.losses import CategoricalCrossentropy
 from tensorflow.keras.metrics import Recall, AUC, Precision
 
 
-class RGBModel(BaseModel):
-    SAVED_MODEL_FOLDER_NAME = 'saved-models/rgb/'
+class RGBRecurrentModel(BaseModel):
+    MODEL_NAME = 'rgb'
 
     def __init__(self, working_folder, dataset_root_path):
-        super(RGBModel, self).__init__(working_folder, dataset_root_path)
+        super(RGBRecurrentModel, self).__init__(working_folder, dataset_root_path)
         self.logger = logging.getLogger(__name__)
 
     def get_dataset(self, dataset_path, **kwargs) -> tf.data.Dataset:

### 2019-12-01 src/main/learning/models/wrapper_layer.py

@@ -1,0 +1,42 @@
+import tensorflow as tf
+
+from tensorflow.keras.layers import Layer
+from models.opticalflow_model import OpticalFlowModel
+from models.rgb_recurrent_model import RGBRecurrentModel
+
+
+class WrapperLayer(Layer):
+    def __init__(self, working_folder, dataset_root_path, **kwargs):
+        super(WrapperLayer, self).__init__()
+
+        self.working_folder = working_folder
+        self.dataset_root_path = dataset_root_path
+
+        self.img_width = kwargs['ImageWidth']
+        self.img_height = kwargs['ImageHeight']
+        self.no_channels = kwargs['NoChannels']
+        self.learning_rate = kwargs['LearningRate']
+
+        self.opticalflow = OpticalFlowModel(working_folder, dataset_root_path)
+        self.recurrent_rgb = RGBRecurrentModel(working_folder, dataset_root_path)
+
+        self.opticalflow_model = None
+        self.rgb_recurrent_model = None
+
+    def build(self, input_shape):
+        self.opticalflow_model = self.opticalflow.get_model(
+            ImageWidth=self.img_width,
+            ImageHeight=self.img_height,
+            NoChannels=self.no_channels,
+            LearningRate=self.learning_rate
+        )
+
+        self.rgb_recurrent_model = self.recurrent_rgb.get_model(
+            ImageWidth=self.img_width,
+            ImageHeight=self.img_height,
+            NoChannels=self.no_channels,
+            LearningRate=self.learning_rate
+        )
+
+    def call(self, inputs, **kwargs):
+        return tf.matmul(input, 0)

### 2019-12-01 src/main/pre_processing/main.py

@@ -43,8 +43,8 @@ if __name__ == '__main__':
     working_folder = os.getenv('WORK_DIR', './')
 
     # configure logging
-    setup_logging(working_folder, 'pre-processing-logging.yaml')
+    setup_logging(working_folder, 'pre_processing-logging.yaml')
     logger = logging.getLogger(__name__)
 
-    # execute video pre-processing logic
+    # execute video pre_processing logic
     main()

### 2019-12-02 article/article-tesis.tex
updating the article

@@
 We produced a set of video annotations indicating signs appearances for a small set of nouns and numbers along with a \textit{labeled PSL dataset} (PSL dataset).
-A composite model obtained from the combination of a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade optical flow, and a RNN with LSTM cells trained with raw rgb frames extracted from the PSL dataset achieved XX precision and XX exhaustivity on sings classification tasks, reporting state of art results over the PSL dataset.
+A composite model obtained from the combination of a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade optical flow, and a RNN with LSTM cells trained with raw rgb frames extracted from the PSL dataset achieved 0.5994 AUC for the Presission-Recall curve on sings classification tasks, reporting state of art results over the PSL dataset.
 \\
@@
 \subsubsection{Architecture}\label{nsdm-architecture}
-nSDm is composed of several layers \ref{video-annotation-results}\\
-\textbf{Samples feeding layer:} A custom layer with the responsibility direct PSL samples \\
-\textbf{Optical flow model layer:} \\
-\textbf{Recurrent RGB model layer:} \\
-\textbf{Concatenation layer:} \\
-\textbf{Fully connected layer:} \\
-
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/nsdm-model-architecture.png}
-\caption{nSDm model architecture}
-\label{fig:opticalflow-two}
-\end{figure}
-
+A series of models stacked in a composed architecture focused on learning visual and time series features using two network architecture flavors enforced with a flexible data input pipeline for data feeding as shown in Figure \ref{fig:ndsm-architecture}, transformation and normalization.
+The \textbf{Input pipeline} accesses the labeled PSL samples and applies transformations preparing the data for upper layers, transformations were applied to both optical flow and RGB frames, PSL samples were resized to be compliant with ImageNet pre-trained models using a 224 by 224 shape and three channels for color images, data augmentation transformation were not applied due to the nature of the experiment where samples were captured using similar light conditions and camera orientations, PSL samples were transformed to tensors and normalized to floats in the [0, 1] interval, finally the transformed and normalized versions of optical flow and RGB samples were tight together in a tensor structure along with the label for usage at upper layers.
 \begin{table}[!htb]
@@
 \end{table}
+The \textbf{Wrapper Layer} wraps a CNN network designed for learning visual features from optical flow and a RNN network designed for learning features from series of video frames, the layer receives data from the input pipeline then splits it into optical flow and RGB samples and passes them to the CNN and RNN networks respectively, the logits  produced are then concatenated and passed to the dense layer.
+The dense layer uses a softmax activation function to convert logits into probabilities used for a correct sign classification.
+Hyper parameters used during the training process are listed on Table \ref{tab:nSDm-hyperparams}
+
+\begin{figure}[hbt!]
+\includegraphics[width=\linewidth]{images/nsdm-model-architecture.png}
+\caption{nSDm model architecture}
+\label{fig:ndsm-architecture}
+\end{figure}
+
+The \textbf{Opticalflow model} described in Figure \ref{fig:opticalflow-architecture} receives decoded opticalflow samples, it uses a Resnet152 backbone pre-trained with ImageNet, we used a fine tuning transfer learning approach, the backbone produces a 7x7x2048 output that then is passed to a Global Average Pooling layer for obtaining a flattened output of 1x1x2048 which is then passed to a dense layer for only logits generation.
 
@@
 \caption{Opticalflow model architecture}
-\label{fig:opticalflow-two}
+\label{fig:opticalflow-architecture}
 \end{figure}
+
+The \textbf{Recurrent RGB model} described in Figure \ref{fig:rgb-architecture} receives a sequence of decoded video frames, it uses a \textit{many to one} architecture with LSTM cells that hold state of 2048 length, the output produced by the recurrent layers are then passed to a dense layer for only logits generation.
 
@@
 \caption{Recurrent RGB model architecture}
-\label{fig:opticalflow-two}
+\label{fig:rgb-architecture}
 \end{figure}
 
+Logits produced by the opticalflow and recurrent RGB models are concatenated and then passed to a dense layer with a softmax activation function producing PSL signs detection probabilities.
+
 \subsection{Sign detection testing}\label{new-videos-testing}
+We selected videos for consultant 24 \ref{datasetdesc} for testing nSDm (the model was not trained with consultant 24 videos), we proceed to extract RGB frames for video fragments of five seconds length moving ahead one second on each loop until the end of the video that way we can reduce the number of signs cut or incomplete between video fragments.
+Optical flow and RGB frames generators feed nSDm for detecting signs \ref{nsdm} present in the video fragment.
 
+A software component called Video Updater uses video metadata produced by the RGB Frames Generator like start and end time in seconds and detected signs for annotating the video adding a mask that highlights detected signs.
+We used OpenCV for incorporating masks to test videos, Figure \ref{fig:video-testing-architecture} shows the signs detection testing architecture.
 \begin{figure}[hbt!]
@@
 \caption{Signs detection testing process architecture}
-\label{fig:opticalflow-two}
+\label{fig:video-testing-architecture}
 \end{figure}
@@
 \subsection{Sign detection results}
+We trained nSDm with the train PSL dataset and validated it with the validation PSL dataset, preliminary we trained the model only for five epochs obtaining the results in Table \ref{tab:signs-detection-results}
+% Sign detection results
+\begin{table}[!htb]
+\captionsetup{font=footnotesize}
+\centering
+\begin{tabular}{ p{16em} c }
+\toprule
+\multicolumn{1}{c}{\textbf{Epoch}} & 
+	\multicolumn{1}{c}{\textbf{AUC}}\\
+\midrule
+\textbf{Epoch 1}&	0.0368\\
+\textbf{Epoch 2}&	0.5994\\
+\textbf{Epoch 3}&	0.0412\\
+\textbf{Epoch 4}&	0.0703\\
+\textbf{Epoch 5}&	0.0680\\
+\bottomrule
+\end{tabular}
+\caption{Shows results of training nSDm with the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{AUC} area under the precision-recall curve (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall.}
+\label{tab:signs-detection-results}
+\end{table}
 
 \section{Conclusion}\label{conclusion}
+\begin{enumerate}[i]
+\item The ability of detecting recording sessions participants and a mapping between transcripted words and video instants are necessary for a correct video annotation because it allows differentiating PSL translations from other recording session transcriptions.
+\item Video annotation requires a considerable amount of manual work, even though several steps in the proposed video annotation pipeline can be automated e.g.
+video splitting, audio transcription, syntax detection and metadata generation, pre video annotation and post verification steps are executed manually.
+\item Post samples verification steps are required due to the \textit{Delay Factor}, frames not related to a sign are deleted to improve the labeled PSL dataset quality.
+\item Using optical flow for movement pattern representation allow capturing all actions and movements executed by a PSL consultant.
+\item A CNN can be trained with optical flow samples to recognize human actions making possible to learn PSL signs movement patterns.
+\item RGB video frames extracted from video fragments can be used to train a RNN for learning human actions, the network learns movement sequences between frames making possible to learn PSL signs movement sequences.
+\item Combining a CNN and a RNN in a composite model improves signs detection performance, while CNNs learns visual features in optical flow samples RNNs learns sequences of actions, networks outputs can be pondered for training a better classifier.
+\item The area under the precision-recall curve allow measuring how well is nSDm detecting because it summarizes the trade-off between the true positive signs rate and the predicted signs.
+\end{enumerate}
 

### 2019-12-02 src/main/learning/models/base_model.py

@@ -69,7 +69,7 @@ class BaseModel:
 
     def get_labels(self, train=True) -> list:
         dataset_folder_name = self.TRAIN_DATASET_FOLDER_NAME if train else self.TEST_DATASET_FOLDER_NAME
-        path = os.path.join(self.get_dataset_path(dataset_folder_name), 'opticalflow')
+        path = self.get_dataset_path(dataset_folder_name)
         return [label for label in os.listdir(path) if os.path.isdir(os.path.join(path, label))]
 
     def get_saved_model(self) -> Model:
@@ -83,7 +83,7 @@ class BaseModel:
         self.logger.debug('model saved to %s', saved_model_path)
 
     def get_dataset_path(self, operation_folder_name):
-        dataset_path = os.path.join(self.dataset_root_path, operation_folder_name)
+        dataset_path = os.path.join(self.dataset_root_path, operation_folder_name, self.MODEL_NAME)
 
         if not os.path.exists(dataset_path):
             self.logger.error('The provided dataset path %s does not exist.', dataset_path)

### 2019-12-03 article/article-tesis.tex

@@
 	\textbf{Talking with signs} \\	
-	\Large \textbf{A simple method to detect nouns and numbers in a non annotated signs language corpus}
+	\Large \textbf{A simple method to detect nouns and numbers in a non-annotated signs language corpus}
 }
@@
 \begin{abstract}
-People with deafness or hearing disabilities who aim to use computer based systems relies on state of art video classification and human action recognition techniques that combine traditional movement pattern recognition and deep learning techniques.
-In this work we present a pipeline for semi automatic video annotation applied to a non annotated \textit{Peruvian Signs Language} (PSL) corpus along with a novel method for a progressive detection of PSL elements (nSDm).
+People with deafness or hearing disabilities who aim to use computer based systems rely on state-of-art video classification and human action recognition techniques that combine traditional movement pattern recognition and deep learning techniques.
+In this work we present a pipeline for semi-automatic video annotation applied to a non-annotated \textit{Peruvian Signs Language} (PSL) corpus along with a novel method for a progressive detection of PSL elements (nSDm).
 We produced a set of video annotations indicating signs appearances for a small set of nouns and numbers along with a \textit{labeled PSL dataset} (PSL dataset).
-A composite model obtained from the combination of a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade optical flow, and a RNN with LSTM cells trained with raw rgb frames extracted from the PSL dataset achieved 0.5994 AUC for the Presission-Recall curve on sings classification tasks, reporting state of art results over the PSL dataset.
+A composite model obtained from the combination of a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade optical flow, and a RNN with LSTM cells trained with raw rgb frames extracted from the PSL dataset achieved 0.5994 AUC for the Precision-Recall curve on signs classification tasks, reporting state-of-art results over the PSL dataset.
 \\
@@
 
-The Peruvian Institute of Informatics and Statistics (INEI) conducted a national disabilities survey with the objective of segmenting and acquire a better understanding about disabilities that affect the Peruvian population \cite{disabilities_survey_2012}, the results showed that 1.8\% of the Peruvian population suffer at least partial when not permanent deafness or hearing limitations.
+The Peruvian Institute of Informatics and Statistics (INEI) conducted a national disabilities survey with the objective of segmenting and acquiring a better understanding about disabilities that affect the Peruvian population \cite{disabilities_survey_2012}.
+Results showed that 1.8\% of the Peruvian population suffer at least partial when not permanent deafness or hearing limitations.
 
 Peruvians with deafness or hearing limitations use the Peruvian Signs Language (PSL) as their main communication medium.
-PSL is of mandatory usage at universities and certain public institutions, here the importance of designing systems that are capable to support PSL inputs and outputs.
+PSL is of mandatory usage at universities and certain public institutions, henceforth the importance of designing systems that are capable to support PSL inputs and outputs.
 Furthermore, in the same way as spoken languages, signs languages also present local variations e.g.
@@
 
-\textit{The Grammar and Signs research group of the Pontifical Catholic University of Peru (PUCP)} built the first PSL corpus \cite{lsp_dataset} which is publicly available at the university digital archives, it is important to highlight that the corpus is neither labeled or annotated and cannot be used as it is for training or testing a model.
+\textit{The Grammar and Signs research group of the Pontifical Catholic University of Peru (PUCP)} built the first PSL corpus \cite{lsp_dataset} which is publicly available at the university digital archives.
+It is important to highlight that the corpus is neither labeled or annotated and cannot be used as it is for training or testing a model.
 
 In this work we are approaching signs detection as a supervised learning task.
-Supervised learning requires labeled dataset to achieve satisfactory results during training and inference tasks.
-At the time of writing this work there were no labeled datasets available for PSL \cite{lsp_2015}, it configures a gap that could prevent or hinder research work on Human-Computer-Interaction at the Peruvian or Latin American space.
+Supervised learning requires labeled datasets to achieve satisfactory results during training and inference tasks.
+At the time of writing this work there were no labeled datasets available for PSL \cite{lsp_2015}.
+It configures a gap that could prevent or hinder research work on Human-Computer-Interaction at the Peruvian or Latin American space.
 
@@
 \begin{enumerate}[(i)]
-\item What are the currently available techniques for producing a labeled dataset for a set of signs limited to nouns and number from the non annotated PSL corpus?\label{q1}
+\item What are the currently available techniques for producing a labeled dataset for a set of signs limited to nouns and number from the non-annotated PSL corpus?\label{q1}
 \item What are most relevant and currently available techniques for training a model with the labeled dataset described in the question above for detecting PSL nouns and numbers?\label{q2}
@@
 
-This work has the main objective of producing a simple method that can be used as a baseline for other researchers interested on studying signs language and their different applications on the Human-Computer-Interaction field, we believe this work will produce a positive impact on the artificial intelligence community towards to an increase in the number of research works using PSL which can contribute to increasing the number of people with deafness or hearing disabilities that can use computer based systems.
+This work has the main objective of producing a simple method that can be used as a baseline for other researchers interested on studying signs language and their different applications on the Human-Computer-Interaction field, we believe this work will produce a positive impact on the artificial intelligence community towards an increase in the number of research works using PSL which can contribute to increasing the number of people with deafness or hearing disabilities that can use computer based systems.
 We have divided this work main objective into following specific objectives for better traceability:
@@
 \begin{enumerate}[(i)]
-\item Produce a labeled PSL dataset limited to nouns and number\label{o1}
-\item Design a train a novel signs detection model (nSDm) for detecting signs at the labeled PSL dataset\label{o2}
+\item Produce a labeled PSL dataset limited to nouns and numbers\label{o1}
+\item Design and train a novel signs detection model (nSDm) for detecting signs at the labeled PSL dataset\label{o2}
 \item Determine what is the performance of nSDm in terms of precision and recall\label{o3}
@@
 The rest of the article is organized as follows.
-In section \ref{relatedwork} we review the related work on video classification for human actions recognition using network architectures that combine CNNs, 3D CNNs and movement patterns  for better features learning, we also review state of art pose estimation techniques.
+In section \ref{relatedwork} we review the related work on video classification for human actions recognition using network architectures that combine CNNs, 3D CNNs and movement patterns  for better features learning, we also review state-of-art pose estimation techniques.
 In section \ref{method} we introduce nSDm describing its design and architecture.
@@
 \subsection{Action Recognition}
-Human action recognition is a extensively studied field, action recognition dataset like UCF101, HMDB51, THUMOS14  are available, researches tried to solve the human action recognition problem using different approaches including Optical Flow and 3D CNN \cite{qiu2017learning}.
-\textbf{Optical Flow,} is defined as the pattern obtained from the motion of objects, surfaces and edges in a visual scene caused by the relative motion between the observer and a scene, it is achieved by distributing movement velocities and brightness across frames.
+Human action recognition is an extensively studied field.
+Action recognition dataset like UCF101, HMDB51, THUMOS14  are available, researches tried to solve the human action recognition problem using different approaches including Optical Flow and 3D CNN \cite{qiu2017learning}.
+
+\textbf{Optical Flow,} is defined as the pattern obtained from the motion of objects, surfaces and edges in a visual scene caused by the relative motion between the observer and a scene.
+It is computed by distributing movement velocities and brightness across frames.
 It is a  key concept in action recognition from videos \cite{wang2019hallucinating}.
@@
 Various types of camera motion can be observed in realistic videos, e.g., zooming, tilting, rotation, etc.
-\textbf{Motion Boundary Histogram (MBH),} It is a simple an efficient way to achieve robustness during human action detection when camera movements are mixed within the recorded actions by computing derivatives separately for the horizontal and vertical components of the optical flow.
+
+\textbf{Motion Boundary Histogram (MBH)} is a simple an efficient way to achieve robustness during human action detection when camera movements are mixed within the recorded actions by computing derivatives separately for the horizontal and vertical components of the optical flow.
 Since MBH represents the gradient of optical flow, locally constant camera motion is removed and information about changes in the flow field is kept.
 MBH is more robust to camera motion than optical flow, thus more discriminative for action recognition.\cite{wang2013dense}.
-3D CNN are not as effective as optical flow to detect human actions on its own, 3D CNN can be trained to learn optical flow so we can avoid costly computation and storage and obtain task-specific motion representation  \cite{zhu2018hidden} and increase models performance, precision and exhaustiveness on human action recognition.
+3D CNN are not as effective as optical flow to detect human actions on its own, 3D CNN can be trained to learn optical flow so we can avoid costly computation and storage and obtain task-specific motion representation  \cite{zhu2018hidden} and increase models performance, precision and recall on human action recognition.
 \subsection{Pose Estimation}
 Pose estimation is also an extensively studied field.
-Techniques based on key points have shown state of art results on human pose estimation.
+Techniques based on key points have shown state-of-art results on human pose estimation.
 An approach on key points estimation \cite{tulsianiMalik} uses \textit{Point of View Determination} and \textit{Key Points Prediction} components.
-Point of View Determination is formulated by the prediction of three Euler angles (azimut, elevation and cyclotation) generating a global position estimate, then a local appearance is modeled by obtaining a heat map that corresponds to the verisimilitude spatial distribution for each key point, finally key points predictions are obtained by combining heat maps obtained in a previous stage with a conditioned verisimilitude at the point of view predicted in the previous stage.
+Point of View Determination is formulated by the prediction of three Euler angles (azimut, elevation and cyclotation) generating a global position estimate, then a local appearance is modeled by obtaining a heat map that corresponds to the spatial distribution likelihood for each key point, finally key points predictions are obtained by combining heat maps obtained in a previous stage with a conditioned likelihood at the point of view predicted in the previous stage.
 
-Key points detection methods based CNNs have received an special attention in Human Pose Detection problems, CNNs methods are divided in bottom-up and top-down.
-Bottom-up methods process images from low resolution  to high resolution, focusing first on detecting joints and then before associating them to human actions.
-Top-down methods process focused first on detecting humans and then estimating the human pose to predict key points.
+Key points detection methods based CNNs have received an special attention in Human Pose Detection problems.
+CNNs methods are divided in bottom-up and top-down.
+Bottom-up methods process images from low resolution  to high resolution, focusing first on detecting joints before associating them to human actions.
+Top-down methods focus first on detecting human subjects and then estimating the human pose to predict key points.
 
-The datasets MPII and COCO have been used in state of art methods obtaining good results\cite{XiaoWuWeiSimpleBaseline} and establishing a framework for future work in combination with classic approaches like optical flow for recognizing patterns movement between frames by increasing accuracy on key points detection.
+The datasets MPII and COCO have been used in state-of-art methods obtaining good results\cite{XiaoWuWeiSimpleBaseline} and establishing a framework for future work in combination with classic approaches like optical flow for recognizing patterns movement between frames by increasing accuracy on key points detection.
 \subsection{Video Classification}
@@
 In its simplest form, it consists of (i) clustering with k-means a collection of descriptor vectors from the training set to build so-called visual vocabulary, (ii) as signing each descriptor to its nearest cluster center from the visual dictionary, and (iii) aggregating the one-hot assignment vectors via average pooling \cite{wang2019hallucinating}, when applied to Computer Vision is a technique used to create images representations or features vectors used that can be learned by CNNs, resulting on improved images classification and video classification.
-Feature trajectory detection are much improved using statistical methods like \textbf{Fisher Vectors} obtaining better results over traditional BoW \textbf{Fussing parallel CNN.}, the Bag of Visual Words representation suffers from sparsity and high dimensionality, in the other hand representations obtained using the Fisher Vectors kernel are more compact and dense which results on better results on image and video classification problems.
+Feature trajectory detection are much improved using statistical methods like \textbf{Fisher Vectors} obtaining better results over traditional BoW \textbf{Fussing parallel CNN.}.
+The Bag of Visual Words representation suffers from sparsity and high dimensionality, in the other hand representations obtained using the Fisher Vectors kernel are more compact and dense which results on better results for image and video classification problems.
 \section{Method}\label{method}
 \subsection{Video Annotation}\label{videoannot}
-The PSL dataset is non annotated because there is not a direct relation between the instant when a sign is emitted and when its translation to Spanish is delivered, we propose a semi automatic video annotation pipeline described in Figure \ref{fig:video-annotation-pipeline} for cleaning, pre-processing and analyzing PSL videos in order to produce an labeled PSL dataset that can be used for training nSDm using supervised learning.
+The PSL dataset is non-annotated because there is not a direct relation between the instant when a sign is emitted and when its translation to Spanish is delivered.
+We propose a semi-automatic video annotation pipeline described in Figure \ref{fig:video-annotation-pipeline} for cleaning, pre-processing and analyzing PSL videos in order to produce an labeled PSL dataset that can be used for training nSDm using supervised learning.
 The pipeline is described in detail in sections  \ref{manual-video-cleanup}, \ref{video-pre-processing}, \ref{audio-transcription-analysis} and \ref{samples-generation}
@@
 The PSL recordings described on \ref{datasetdesc} contain a considerable amount of noise introduced during recording sessions.
-It makes difficult to easily find video intervals that clearly show a relation between signs emitted by the informant and the translation delivered by the translator, the noise factors are listed in the following list:
+It makes difficult to easily find video intervals that clearly show a relation between signs emitted by the informant and the translation delivered by the translator.
+Noise factors are the following:
 \begin{itemize}
@@
 \end{itemize}
-A manual video cleanup process is required to find noise free video intervals, it is a process that requires watching all the videos available at the PSL corpus for manually annotating the instant when an informant started emitting sings along with the instant when the translator delivered a translation, table \ref{table:noise-free-video-segments} shows a manual annotation example.
+A manual video cleanup process is required to find noise free video intervals.
+This process requires watching all the videos available at the PSL corpus for manually annotating the instant when an informant started emitting sings along with the instant when the translator delivered a translation, Table \ref{table:noise-free-video-segments} shows a manual annotation example.
 
@@
 \subsubsection{Video Pre-Processing}\label{video-pre-processing}
-Non annotated PSL videos require processing before any metadata can be extracted, we propose a sequence of pre-processing tasks that take advantage of the annotation generated on \ref{manual-video-cleanup}.
+Non-annotated PSL videos require processing before any metadata can be extracted, we propose a sequence of pre-processing tasks that take advantage of the annotation generated on \ref{manual-video-cleanup}.
 A \textbf{video splitting processor} generates a set of video chunks using the \textit{ffmpeg} multimedia framework and stores produced video chunks in Amazon S3 for later usage.
@@
 
-At the moment of writing this work Amazon Transcription service only supported Spain and US Spanish, it caused certain words that are specific for Peruvian Spanish not being fully recognized, in order to improve transcription accuracy we built a custom vocabulary containing Peruvian expressions which improved  Peruvian words recognition, for the matters of this work Peruvian words that remained unrecognized were omitted and not processed.
+At the moment of writing this work Amazon Transcription service only supported Spain and US Spanish.
+This caused certain words that are specific for Peruvian Spanish not being fully recognized, in order to improve transcription accuracy we built a custom vocabulary containing Peruvian expressions which improved  Peruvian words recognition, for the matters of this work Peruvian words that remained unrecognized were omitted and not processed.
 
@@
 
-A \textbf{Optical Flow Samples generation processor} uses video frames and the hierarchical folder structure generated by the RGB samples generation processor to calculate an Optical Flow representation for PSL elements and store them in a hierarchical folder structure that will also used by the nSDm data loaders to feed the optical flow branch on the nSDm model architecture \ref{nsdm-architecture} during training and testing.
-We selected optical flow as a samples generation strategy because its ability to represent movement traces from previous frames, it is particular useful for representing body movement patterns executed by informant while emitting a PSL sign, a PSL sign is made up of different body movements including: elbow, arms, neck, eyes, shoulders and hands, which are performed quickly, a way to detect movement traces between frames allows to generate a single image representation of all movement involved on a sign emitting (Figure \ref{fig:opticalflow-two}).
+An \textbf{Optical Flow Samples generation processor} uses video frames and the hierarchical folder structure generated by the RGB samples generation processor to calculate an Optical Flow representation for PSL elements and store them in a hierarchical folder structure that will also be used by the nSDm data loaders to feed the optical flow branch on the nSDm model architecture \ref{nsdm-architecture} during training and testing.
+We selected optical flow as a samples generation strategy due to its ability to represent movement traces from previous frames.
+It is particular useful for representing body movement patterns executed by informant while emitting a PSL sign.
+A PSL sign is made up of different body movements including: elbow, arms, neck, eyes, shoulders and hands, which are performed quickly, a way to detect movement traces between frames allows to generate a single image representation of all movement involved on a sign emission(Figure \ref{fig:opticalflow-two}).
 
@@
 A series of models stacked in a composed architecture focused on learning visual and time series features using two network architecture flavors enforced with a flexible data input pipeline for data feeding as shown in Figure \ref{fig:ndsm-architecture}, transformation and normalization.
+
 The \textbf{Input pipeline} accesses the labeled PSL samples and applies transformations preparing the data for upper layers, transformations were applied to both optical flow and RGB frames, PSL samples were resized to be compliant with ImageNet pre-trained models using a 224 by 224 shape and three channels for color images, data augmentation transformation were not applied due to the nature of the experiment where samples were captured using similar light conditions and camera orientations, PSL samples were transformed to tensors and normalized to floats in the [0, 1] interval, finally the transformed and normalized versions of optical flow and RGB samples were tight together in a tensor structure along with the label for usage at upper layers.
@@
 \end{table}
-The \textbf{Wrapper Layer} wraps a CNN network designed for learning visual features from optical flow and a RNN network designed for learning features from series of video frames, the layer receives data from the input pipeline then splits it into optical flow and RGB samples and passes them to the CNN and RNN networks respectively, the logits  produced are then concatenated and passed to the dense layer.
+
+The \textbf{Wrapper Layer} wraps a CNN network designed for learning visual features from optical flow and a RNN network designed for learning features from series of video frames, the layer receives data from the input pipeline then splits it into optical flow and RGB samples and passes them to the CNN and RNN networks respectively.
+The logits produced are then concatenated and passed to the dense layer.
+
 The dense layer uses a softmax activation function to convert logits into probabilities used for a correct sign classification.
@@
 
-The \textbf{Opticalflow model} described in Figure \ref{fig:opticalflow-architecture} receives decoded opticalflow samples, it uses a Resnet152 backbone pre-trained with ImageNet, we used a fine tuning transfer learning approach, the backbone produces a 7x7x2048 output that then is passed to a Global Average Pooling layer for obtaining a flattened output of 1x1x2048 which is then passed to a dense layer for only logits generation.
+The \textbf{Opticalflow model} described in Figure \ref{fig:opticalflow-architecture} receives decoded opticalflow samples, it uses a Resnet152 backbone pre-trained with ImageNet, we used a fine tuning transfer learning approach, the backbone produces a 7x7x2048 output that then is passed to a Global Average Pooling layer for obtaining a flattened output of 1x1x2048 which is then passed to a dense layer for logits computation.
 
@@
 
-The \textbf{Recurrent RGB model} described in Figure \ref{fig:rgb-architecture} receives a sequence of decoded video frames, it uses a \textit{many to one} architecture with LSTM cells that hold state of 2048 length, the output produced by the recurrent layers are then passed to a dense layer for only logits generation.
+The \textbf{Recurrent RGB model} described in Figure \ref{fig:rgb-architecture} receives a sequence of decoded video frames, it uses a \textit{many-to-one} architecture with LSTM cells that hold state of 2048 units length, the output produced by the recurrent layers is then passed to a dense layer for logits computation.
 
@@
 \subsection{Sign detection testing}\label{new-videos-testing}
-We selected videos for consultant 24 \ref{datasetdesc} for testing nSDm (the model was not trained with consultant 24 videos), we proceed to extract RGB frames for video fragments of five seconds length moving ahead one second on each loop until the end of the video that way we can reduce the number of signs cut or incomplete between video fragments.
+We selected videos for consultant 24 which will be described in \ref{datasetdesc} for testing nSDm (the model was not trained with consultant 24 videos), we proceed to extract RGB frames for video fragments of five seconds length moving ahead one second on each loop until the end of the video that way we can reduce the number of signs cut or incomplete between video fragments.
 Optical flow and RGB frames generators feed nSDm for detecting signs \ref{nsdm} present in the video fragment.
@@
 \subsection{Dataset Description \cite{lsp_dataset}}\label{datasetdesc}
-The PSL dataset was developed by the PUCP Grammar and Signs research group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female informants, all of them are Lima Peru residents and reported to be borne with a permanent deafness condition or acquired the condition before the acquisition of Spanish.
+The PSL dataset was developed by the PUCP Grammar and Signs research group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female informants, all of them are Lima Peru residents and reported to be born with a permanent deafness condition or acquired the condition before the acquisition of Spanish.
 

### 2019-12-03 article/article-tesis.tex
adding additional signs detection results to the article

@@
 \subsection{Sign detection results}
-We trained nSDm with the train PSL dataset and validated it with the validation PSL dataset, preliminary we trained the model only for five epochs obtaining the results in Table \ref{tab:signs-detection-results}
-% Sign detection results
+We trained nSDm with the train PSL dataset and validated it with the validation PSL dataset, we trained the model  for five epochs using five and ten percent of the dataset obtaining the results in Tables \ref{tab:signs-detection-results-5-percent} and \ref{tab:signs-detection-results-10-percent}
+
+Train results shows a trend indicating that the more PSL samples we use the nSDm performance improves highlighting the need for processing additional videos at the PSL corpus.
+% Sign detection results 5% of the dataset
 \begin{table}[!htb]
@@
 \midrule
-\textbf{Epoch 1}&	0.0368\\
-\textbf{Epoch 2}&	0.5994\\
-\textbf{Epoch 3}&	0.0412\\
-\textbf{Epoch 4}&	0.0703\\
-\textbf{Epoch 5}&	0.0680\\
+\textbf{Epoch 1}&	0.0708\\
+\textbf{Epoch 2}&	0.0539\\
+\textbf{Epoch 3}&	0.0558\\
+\textbf{Epoch 4}&	0.1175\\
+\textbf{Epoch 5}&	0.0368\\
 \bottomrule
 \end{tabular}
-\caption{Shows results of training nSDm with the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{AUC} area under the precision-recall curve (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall.}
-\label{tab:signs-detection-results}
+\caption{Shows results of training nSDm with the 5\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{AUC} area under the precision-recall curve (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall.}
+\label{tab:signs-detection-results-5-percent}
 \end{table}
+
+% Sign detection results 10% of the dataset
+\begin{table}[!htb]
+\captionsetup{font=footnotesize}
+\centering
+\begin{tabular}{ p{16em} c }
+\toprule
+\multicolumn{1}{c}{\textbf{Epoch}} & 
+	\multicolumn{1}{c}{\textbf{AUC}}\\
+\midrule
+\textbf{Epoch 1}&	0.3686\\
+\textbf{Epoch 2}&	0.5994\\
+\textbf{Epoch 3}&	0.1412\\
+\textbf{Epoch 4}&	0.1703\\
+\textbf{Epoch 5}&	0.1680\\
+\bottomrule
+\end{tabular}
+\caption{Shows results of training nSDm with the 10\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{AUC} area under the precision-recall curve (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall.}
+\label{tab:signs-detection-results-10-percent}
+\end{table}
+
+Results of infer PSL samples are presented on Figure \ref{fig:rgb-pareja-sample} we observed the noun \textit{Pareja} being correctly classified, in the other hand we observe that silent periods are being incorrectly classified as shown on Figure \ref{fig:rgb-pareja-wrong-sample}, adding one additional class for \textit{silent} could improve our classifier allowing it to infer when a consultant is to performing any action.
+
+\begin{figure}[hbt!]
+\includegraphics[width=\linewidth]{images/pareja-rgb-frames.png}
+\caption{PSL noun "Pareja" RGB Frames correctly classified}
+\label{fig:rgb-pareja-sample}
+\end{figure}
+
+\begin{figure}[hbt!]
+\includegraphics[width=\linewidth]{images/pareja-rgb-frame-wrong.png}
+\caption{PSL noun "Pareja" RGB Frames incorrectly recognized. The classifier is incorrectly classifying silent periods where the consultant is not performing any action}
+\label{fig:rgb-pareja-wrong-sample}
+\end{figure}
+
+\section{Discussion and Future Work}\label{discussion-future-work}
+\begin{enumerate}[i]
+\item Manual video annotation to be replaced for a fully automated process
+\item Adding an additional class called \textit{silent} which we can use to label PSL samples where the consultant is not performing any action.
+
+\end{enumerate}
 

### 2019-12-03 src/main/learning/main.py

@@ -11,7 +11,7 @@ from logger_config import setup_logging
 
 DEFAULT_LEARNING_RATE = 0.001
 DEFAULT_NO_EPOCHS = 5
-DEFAULT_NO_STEPS_EPOCHS = 2
+DEFAULT_NO_STEPS_EPOCHS = 4
 DEFAULT_BATCH_SIZE = 1
 DEFAULT_NO_CHANNELS = 3  # Color images
 DEFAULT_IMG_WIDTH = 224  # Imagenet default image width

### 2020-02-03 article/article-tesis.tex
fix to the first version Signs Language Scraper

@@
 \section{Discussion and Future Work}\label{discussion-future-work}
-\begin{enumerate}[i]
-\item Manual video annotation to be replaced for a fully automated process
-\item Adding an additional class called \textit{silent} which we can use to label PSL samples where the consultant is not performing any action.
+\begin{enumerate}[•]
+\item The ability to detect all participants in recording sessions individually is mandatory for a successful video annotation pipelines based on video transcriptions as the one we presented at \ref{videoannot}.
+Future work about annotating videos based on patterns recognized in video content itself using non supervised learning approaches may improve the generated dataset driving to better results.
+\item Existing video annotation pipelines require large amounts of manual annotation work, generally for cleaning and validation tasks.
+Even though several steps within the overall annotation process can be fully automated human intervention for ensuring the quality of the generated dataset.
+Future work that use Reinforcement Learning for training an agent trained to clean and validate non-annotated videos before and after they are passed through the annotation pipeline may minimize or remove the need of human intervention during video annotation tasks.
+\item Not all the corporal movements executed by a PSL consultant are valid PSL signs or elements.
+Not handling these exceptions correctly could affect greatly the nSDm performance.
+Future work that add support for an additional class called \textit{invalid} and labeling invalid signs or corporal movements could help nSDm to learn invalid PSL actions and obtain better results.
+\item Optical flow represents movement patterns between video frames allowing models to learn these given movement patterns.
+However, Optical flow could consider camera unintentional movements and camera focus switches incorrectly as valid movement patterns between frames.
+Future work that adds Motion Boundary Histogram (MBH) \cite{wang2013dense} support to the optical flow gradient could add robustness to the ability to detect camera movements driving to a better optical flow samples generation.
+\item nSDm internal models (CNN and RNN models) as described on \ref{nsdm-architecture} can be trained independently using different hyper parameters, optimizers and loss functions.
+Pre-trained models will boost the nSDm training process enabling future works to use PSL to train nSDm using more epochs, data and being more flexible on using single models (CNN or RNN) or combining them as in nSDm.
+\item Improving they way people with deafness or hearing disabilities interact with computers requires delivering an output after PSL signs are detected.
+Future work that use Auto Encoder or Adversarial Generative Networks can be used to generate output in the form of signs that are recognizable completing the cycle of a bidirectional communication.
+\item Training nSDm with more PSL samples greatly improves its performance. Future work on annotating more videos including PSL content will increase nSDm performance.
 
@@
 \section{Conclusion}\label{conclusion}
-\begin{enumerate}[i]
-\item The ability of detecting recording sessions participants and a mapping between transcripted words and video instants are necessary for a correct video annotation because it allows differentiating PSL translations from other recording session transcriptions.
-\item Video annotation requires a considerable amount of manual work, even though several steps in the proposed video annotation pipeline can be automated e.g.
-video splitting, audio transcription, syntax detection and metadata generation, pre video annotation and post verification steps are executed manually.
-\item Post samples verification steps are required due to the \textit{Delay Factor}, frames not related to a sign are deleted to improve the labeled PSL dataset quality.
+\begin{enumerate}[•]
+\item Video annotation pipelines based on video transcriptions require recognizing all video participants for a correct video annotation.
+\item The \textit{Delay Factor} introduced noise to the generated labeled PSL dataset. Post annotation human intervention is needed to reduce the delay factor effect.
+\item Annotating more PSL videos and training nSDm with these new generated samples increases the model performance.
+\item Optical flow successfully represents PSL consultant corporal movements over time.
+\item Transfering weights learned a ResNet152 model trained with ImageNet influence positively Optical flow patterns learning.
 \item Using optical flow for movement pattern representation allow capturing all actions and movements executed by a PSL consultant.
-\item A CNN can be trained with optical flow samples to recognize human actions making possible to learn PSL signs movement patterns.
-\item RGB video frames extracted from video fragments can be used to train a RNN for learning human actions, the network learns movement sequences between frames making possible to learn PSL signs movement sequences.
+\item RGB video frames extracted from video fragments allow networks to learn sequences of movement over time.
+\item Using LSTM cells allow successfully usage of larger PSL sequences (a larger number of frames) due to its better handling of the vanishing gradients.
 \item Combining a CNN and a RNN in a composite model improves signs detection performance, while CNNs learns visual features in optical flow samples RNNs learns sequences of actions, networks outputs can be pondered for training a better classifier.

### 2020-02-03 src/main/scraping/signs_language_scraper.py

@@ -1,0 +1,31 @@
+import requests
+from bs4 import BeautifulSoup
+
+
+class SignsLanguageScraper:
+    def __init__(self, base_url):
+        self.base_url = base_url
+        page = requests.get(self.base_url)
+        self.soup = BeautifulSoup(page.content, 'html.parser')
+
+    def scrap(self):
+        recording_links = self.get_recording_links()
+        for recording_link in recording_links:
+            recording_name = recording_link.text
+            recording_url = recording_link['href']
+
+            video_links = self.get_recording_video_links(recording_url)
+            for video_link in video_links:
+                print()
+
+    def get_recording_links(self):
+        links = self.soup.find_all('a', lambda x: x.next_sibling['class'] == 'Z3988')
+        return links
+
+    @staticmethod
+    def get_recording_video_links(link_url):
+        recording_page = requests.get(link_url)
+        recording_soup = BeautifulSoup(recording_page.content, 'html.parser')
+
+        video_links = recording_soup.find_all('a', href=lambda url: '.mp4' or '.flv' in url.lower())
+        return video_links

### 2020-07-06 article/article-tesis.tex
casting no_epochs argument to integer

@@
 
-\begin{figure}[hbt!]
+We used the PSL dataset to train and test a set of neural networks described in detail in sections \ref{opticalflowmodel}, \ref{rgbrecurrentmodel} and \ref{nsdm}
+
+Implementation details can be found at https://github.com/erichuizapucp/signs-recognition
+
+\begin{figure*}[hbt!]
 \includegraphics[width=\linewidth]{images/video-annotation-pipeline.png}
@@
 \label{fig:video-annotation-pipeline}
-\end{figure}
+\end{figure*}
 
@@
 A manual video cleanup process is required to find noise free video intervals.
-This process requires watching all the videos available at the PSL corpus for manually annotating the instant when an informant started emitting sings along with the instant when the translator delivered a translation, Table \ref{table:noise-free-video-segments} shows a manual annotation example.
+This process requires watching all videos available at the PSL corpus for manually annotate the instant when an informant started emitting sings along with the instant when the translator delivered a translation, Table \ref{table:noise-free-video-segments} shows a manual annotation example.
 
@@
 
-\subsection{Novel Signs Detection Model (nSDm)}\label{nsdm}
-We propose a novel model for signs detection that combine two neural networks architectures, a CNN and a RNN with the objective to learn visual features like edges, corners and ridges (CNN) and at the same time patterns learned from a time based series of inputs (RNN).
-CNN network receives optical flow inputs and the RNN branch receives RGB frames extracted from the labeled PSL dataset described in \ref{videoannot}.
-
-For this work we selected the Tensorflow/Keras functional API for its ability to define combined models along with a versatile data extraction and transformation layer.
-More details and usage are provided in sections \ref{nsdm-architecture} and \ref{new-videos-testing}
-
-Implementation details can be found at https://github.com/erichuizapucp/signs-recognition
-\subsubsection{Architecture}\label{nsdm-architecture}
-A series of models stacked in a composed architecture focused on learning visual and time series features using two network architecture flavors enforced with a flexible data input pipeline for data feeding as shown in Figure \ref{fig:ndsm-architecture}, transformation and normalization.
-
-The \textbf{Input pipeline} accesses the labeled PSL samples and applies transformations preparing the data for upper layers, transformations were applied to both optical flow and RGB frames, PSL samples were resized to be compliant with ImageNet pre-trained models using a 224 by 224 shape and three channels for color images, data augmentation transformation were not applied due to the nature of the experiment where samples were captured using similar light conditions and camera orientations, PSL samples were transformed to tensors and normalized to floats in the [0, 1] interval, finally the transformed and normalized versions of optical flow and RGB samples were tight together in a tensor structure along with the label for usage at upper layers.
-\begin{table}[!htb]
-\captionsetup{size=footnotesize}
-\begin{tabular}{ p{16em} c}
-\toprule
-\multicolumn{1}{c}{\textbf{Hyper parameter}} & 
-	\multicolumn{1}{c}{\textbf{Value}}\\
-\midrule
-\textbf{Learning Rate}&	0.001\\
-\textbf{No Epochs}&	5\\
-\textbf{No Steps per Epoch}&	5\\
-\textbf{Batch Size}&	10\\
-\textbf{Shuffle Buffer Size}&	50\\
-\bottomrule
-\end{tabular}
-\caption{nSDm training hyper parameters} \label{tab:nSDm-hyperparams}
-\end{table}
-
-The \textbf{Wrapper Layer} wraps a CNN network designed for learning visual features from optical flow and a RNN network designed for learning features from series of video frames, the layer receives data from the input pipeline then splits it into optical flow and RGB samples and passes them to the CNN and RNN networks respectively.
-The logits produced are then concatenated and passed to the dense layer.
-
-The dense layer uses a softmax activation function to convert logits into probabilities used for a correct sign classification.
-Hyper parameters used during the training process are listed on Table \ref{tab:nSDm-hyperparams}
-
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/nsdm-model-architecture.png}
-\caption{nSDm model architecture}
-\label{fig:ndsm-architecture}
-\end{figure}
-
-The \textbf{Opticalflow model} described in Figure \ref{fig:opticalflow-architecture} receives decoded opticalflow samples, it uses a Resnet152 backbone pre-trained with ImageNet, we used a fine tuning transfer learning approach, the backbone produces a 7x7x2048 output that then is passed to a Global Average Pooling layer for obtaining a flattened output of 1x1x2048 which is then passed to a dense layer for logits computation.
-
-\begin{figure}[hbt!]
+\subsection{Opticalflow Model}\label{opticalflowmodel}
+The model uses a 2D CNN architecture to learn features from Opticalflow samples calculated from RGB frames using the Lucas Kanade method for features tracking.
+Opticalflow samples hold features tracked from an entire frames set sequentially that way all the features found across frame sets are condensed in a single image.
+\subsubsection{Model Architecture}\label{opticalflowmodel-architecture}
+The Opticalflow model architecture described in Figure \ref{fig:opticalflow-architecture} uses a Resnet152 backbone pre-trained with ImageNet.
+We used a fine tuning transfer learning approach, the backbone produces a 7x7x2048 output that then is passed to a Global Average Pooling layer for obtaining a flattened output of 1x1x2048 which is then passed to a dense layer for logits computation and finally to a softmax activation function for classes probability computation.
+\begin{figure*}[hbt!]
 \includegraphics[width=\linewidth]{images/opticalflow-model-architecture.png}
@@
 \label{fig:opticalflow-architecture}
-\end{figure}
+\end{figure*}
 
-The \textbf{Recurrent RGB model} described in Figure \ref{fig:rgb-architecture} receives a sequence of decoded video frames, it uses a \textit{many-to-one} architecture with LSTM cells that hold state of 2048 units length, the output produced by the recurrent layers is then passed to a dense layer for logits computation.
+\subsection{RGB Recurrent Model}\label{rgbrecurrentmodel}
+The model uses a RNN architecture to learn features in a sequential way from RGB frames set generated by the video annotation pipeline see Figure \ref{fig:video-annotation-pipeline}.
+RGB frame sets hold a sequence of images representing a PSL element.
+We selected a RNN architecture based on Natural Language Processing text based techniques that already shown good results.
+\subsubsection{Model Architecture}\label{rgbrecurrentmodel-architecture}
+The RGB recurrent model architecture described in Figure \ref{fig:rgb-architecture} receives a sequence of decoded video frames bidirectionally where each frame set represents a PSL sample, frames were resized to 128x128 for GPU memory optimization during training decreasing considerably the number of training parameters.
+Frame set samples length varies on each sample requiring a layer to mask entries ensuring same length samples.
+We decided on using a bidirectional approach because we found benefits on learning features from left to right and right to left in the same way as text based NLP.
+It uses a \textit{many-to-one} architecture with LSTM cells that hold state of 64 units length, the output produced by the recurrent layers is then passed to a dense layer for logits computation and subsequent softmax activation function for classes probability computation.
 
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/recurrent-rgb-model-architecture.png}
+\begin{figure*}[hbt!]
+\centering
+\includegraphics[scale=0.8]{images/recurrent-rgb-model-architecture.png}
 \caption{Recurrent RGB model architecture}
 \label{fig:rgb-architecture}
-\end{figure}
+\end{figure*}
 
-Logits produced by the opticalflow and recurrent RGB models are concatenated and then passed to a dense layer with a softmax activation function producing PSL signs detection probabilities.
+\subsection{Novel Signs Detection Model (nSDm)}\label{nsdm}
+We propose a novel model for signs detection that assemble the two neural networks architectures described in sections \ref{opticalflowmodel-architecture} and \ref{rgbrecurrentmodel-architecture} with the objective to learn visual features like edges, corners and ridges (CNN) and at the same time patterns learned from a time based series of inputs (RNN) to boost the performance on detecting PSL elements.
+CNN network receives optical flow inputs and the RNN branch receives RGB frames extracted from the labeled PSL dataset described in \ref{videoannot}.
+
+We designed two neural network architectures for nSDm, both architectures use pre trained Opticalflow and RGB models as base models and applies different model assemble techniques on top of them.
+This architectures are described in detail in sections \ref{nsdmv1-architecture} and \ref{nsdmv2-architecture}.
+
+For this work we selected the Tensorflow/Keras functional API for its ability to define combined models along with a versatile data extraction and transformation layer.
+
+\subsubsection{nSDmV1 Model Architecture}\label{nsdmv1-architecture}
+Pre-trained Opticalflow and RGB recurrent models are assembled using transfer learning with all layers freeze along with a flexible data input pipeline for data feeding, transformation and normalization.
+
+The \textbf{Input pipeline} accesses the labeled PSL samples and applies transformations preparing the data for upper layers, transformations were applied for both Opticalflow and RGB frames, PSL Opticalflow samples were resized to be compliant with ImageNet pre-trained models using a 224 by 224 shape and three channels for color images in the other hand PSL RGB samples were resized to a 128x128 shape for GPU memory optimization, data augmentation transformations were not applied due to the nature of the experiment where samples were captured using similar light conditions and camera orientations, PSL samples were transformed to tensors and normalized to floats in the [0, 1] interval, finally the transformed and normalized versions of optical flow and RGB samples were tight together in a tuple of tensors along with the label for usage at upper layers.
+nSDmV1 architecture is described in Figure \ref{fig:ndsm-v1-architecture}.
+
+We applied transfer learning from Opticalflow and RGB recurrent models where all their layers were frozen that way we save a considerable amount of computation resources, finally Opticalflow and RGB recurrent models outputs are averaged producing the nSDMV1 classifier output.
+
+\begin{figure*}[hbt!]
+\centering
+\includegraphics[scale=1.0]{images/nsdm-model-architecture.png}
+\caption{nSDmV1 model architecture}
+\label{fig:ndsm-v1-architecture}
+\end{figure*}
+
+\subsubsection{nSDm V2 Model Architecture}\label{nsdmv2-architecture}
+
+nSDmV2 architecture inherits many elements from nSDmV1 including the Data input pipeline, data transformation, normalization and Opticalflow and RGB recurrent base models.
+We removed the last dense layers (classifiers) from both base models with the objective to add a single classifier in an outer layer.
+We concatenated the outputs and finally added a Dense layer with a softmax activation function to convert logits into probabilities used for a correct sign classification.
+nSDmV2 architecture is described in Figure \ref{fig:ndsm-v2-architecture}.
+
+\begin{figure*}[hbt!]
+\centering
+\includegraphics[scale=1.0]{images/nsdm-model-architecture-v2.png}
+\caption{nSDmV2 model architecture}
+\label{fig:ndsm-v2-architecture}
+\end{figure*}
 
 \subsection{Sign detection testing}\label{new-videos-testing}
-We selected videos for consultant 24 which will be described in \ref{datasetdesc} for testing nSDm (the model was not trained with consultant 24 videos), we proceed to extract RGB frames for video fragments of five seconds length moving ahead one second on each loop until the end of the video that way we can reduce the number of signs cut or incomplete between video fragments.
-Optical flow and RGB frames generators feed nSDm for detecting signs \ref{nsdm} present in the video fragment.
+We selected videos in \ref{datasetdesc} that were not used for training nSDm models for testing nSDm models (models were not trained with the 24 consultant videos).
+We proceed the input video to extract RGB frames for video fragments of 0.5 seconds length moving ahead 0.1 seconds on each loop until the end of the video that way we can reduce the number of cut or incomplete signs between video fragments.
+Optical flow and RGB frames extractors feed nSDmV2 to detect signs \ref{nsdm} present in the video fragment.
 
-A software component called Video Updater uses video metadata produced by the RGB Frames Generator like start and end time in seconds and detected signs for annotating the video adding a mask that highlights detected signs.
-We used OpenCV for incorporating masks to test videos, Figure \ref{fig:video-testing-architecture} shows the signs detection testing architecture.
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/new-video-test-architecture.png}
+We selected nSDmV2 for signs detection on new videos because we obtained better results with nSDmV2 than with nSDmV1. Experimentation results are described on \ref{experimentation}
+
+A software component called Video Annotator uses video metadata produced by the RGB Frames Extractor including start and end time in seconds along with detected signs.
+Video annotation entries are added to an annotations document which can be used for further video processing like adding a mask that highlights detected signs using a graphics library like OpenCV for incorporating masks to test videos, Figure \ref{fig:video-testing-architecture} shows the signs detection testing architecture.
+\begin{figure*}[hbt!]
+\centering
+\includegraphics[scale=1.0]{images/new-video-test-architecture.png}
 \caption{Signs detection testing process architecture}
 \label{fig:video-testing-architecture}
-\end{figure}
+\end{figure*}
 
@@
 \subsection{Sign detection results}
-We trained nSDm with the train PSL dataset and validated it with the validation PSL dataset, we trained the model  for five epochs using five and ten percent of the dataset obtaining the results in Tables \ref{tab:signs-detection-results-5-percent} and \ref{tab:signs-detection-results-10-percent}
+We trained models described on sections \ref{opticalflowmodel}, \ref{rgbrecurrentmodel} and \ref{nsdm} with the train PSL dataset and validated it with the validation PSL dataset, models were trained during ten epochs using five and ten percent of the dataset obtaining the results in Tables \ref{tab:signs-detection-results-5-percent} and \ref{tab:signs-detection-results-10-percent}.
+
+We used the same hyper parameters while training all models. These are listed on Table \ref{tab:training-hyperparams}
+
+\begin{table}[!htb]
+\captionsetup{size=footnotesize}
+\begin{tabular}{ p{16em} c}
+\toprule
+\multicolumn{1}{c}{\textbf{Hyper parameter}} & 
+	\multicolumn{1}{c}{\textbf{Value}}\\
+\midrule
+\textbf{Learning Rate}&	0.001\\
+\textbf{No Epochs}&	10\\
+\textbf{Batch Size}&	64\\
+\textbf{Shuffle Buffer Size}&	5000\\
+\bottomrule
+\end{tabular}
+\caption{training hyper parameters} \label{tab:training-hyperparams}
+\end{table}
+
+We used the same loss function and optimizer for all models. These are listed on Table \ref{tab:training-loss-optimizer}
+
+\begin{table}[!htb]
+\captionsetup{size=footnotesize}
+\begin{tabular}{ p{10em} c}
+\toprule
+\multicolumn{1}{c}{\textbf{Name}} & 
+	\multicolumn{1}{c}{\textbf{Value}}\\
+\midrule
+\textbf{Loss Function}&	Categorical Cross Entropy\\
+\textbf{Optimizer}&	Adam\\
+\bottomrule
+\end{tabular}
+\caption{Loss and optimization functions} \label{tab:training-loss-optimizer}
+\end{table}
 

### 2020-07-06 src/main/learning/train_model.py

@@ -97,7 +97,7 @@ def main():
     logger = logging.getLogger(__name__)
 
     args = get_cmd_args()
-    no_epochs = args.no_epochs
+    no_epochs = int(args.no_epochs)
     no_steps_per_epoch = args.no_steps
     model_name = args.model
     executor_name = model_name

### 2020-07-06 article/article-tesis.tex
additional updates to the article where we added the new assemble model and updates to the other three existing models

@@
 \subsection{Sign detection results}
-We trained models described on sections \ref{opticalflowmodel}, \ref{rgbrecurrentmodel} and \ref{nsdm} with the train PSL dataset and validated it with the validation PSL dataset, models were trained during ten epochs using five and ten percent of the dataset obtaining the results in Tables \ref{tab:signs-detection-results-5-percent} and \ref{tab:signs-detection-results-10-percent}.
+We trained models described on sections \ref{opticalflowmodel}, \ref{rgbrecurrentmodel} and \ref{nsdm} with the 5\% of the PSL dataset and validated it with the 5\% of the validation PSL dataset, models were trained during ten epochs obtaining the results in Tables \ref{tab:opticalflow-detection-results-5-percent}, \ref{tab:rgb-detection-results-5-percent}, \ref{tab:nSDmV1-detection-results-5-percent} and \ref{tab:nSDmV2-detection-results-5-percent}.
 
@@
 \captionsetup{size=footnotesize}
-\begin{tabular}{ p{16em} c}
+\begin{tabular}{ p{17em} c}
 \toprule
@@
 \captionsetup{size=footnotesize}
-\begin{tabular}{ p{10em} c}
+\begin{tabular}{ p{9em} c}
 \toprule
@@
 
-Train results shows a trend indicating that the more PSL samples we use the nSDm performance improves highlighting the need for processing additional videos at the PSL corpus.
-% Sign detection results 5% of the dataset
-\begin{table}[!htb]
+% Opticalflow results 5% of the dataset
+\begin{table*}[!htb]
 \captionsetup{font=footnotesize}
 \centering
-\begin{tabular}{ p{16em} c }
+\begin{tabular}{ p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} }
 \toprule
-\multicolumn{1}{c}{\textbf{Epoch}} & 
-	\multicolumn{1}{c}{\textbf{AUC}}\\
+\multicolumn{5}{c}{\textbf{Opticalflow Model Training Results with the 5\% of the dataset}} \\
+\hline
+\hline
+\textbf{Epoch}&	\textbf{Loss}	&\textbf{Precision}	&\textbf{Recall}	&\textbf{AUC} \\
+\hline
 \midrule
-\textbf{Epoch 1}&	0.0708\\
-\textbf{Epoch 2}&	0.0539\\
-\textbf{Epoch 3}&	0.0558\\
-\textbf{Epoch 4}&	0.1175\\
-\textbf{Epoch 5}&	0.0368\\
+Epoch 1&	3.2775&	0.0000e+00&	0.0000e+00&	0.0900\\
+Epoch 2&	2.5090&	0.0000e+00&	0.0000e+00&	0.1327\\
+Epoch 3&	2.0908&	1.0000&	0.0455&	0.2616\\
+Epoch 4&	1.8167&	1.0000&	0.0455&	0.4361\\
+Epoch 5&	1.5668&	1.0000&	0.1364&	0.5651\\
+Epoch 6&	1.3303&	1.0000&	0.1364&	0.7557\\
+Epoch 7&	1.1146&	1.0000&	0.3182&	0.8657\\
+Epoch 8&	0.9199&	1.0000&	0.4545&	0.9205\\
+Epoch 9&	0.7464&	1.0000&	0.6364&	0.9657\\
+Epoch 10&	0.5981&	1.0000&	0.6818&	0.9877\\
 \bottomrule
 \end{tabular}
-\caption{Shows results of training nSDm with the 5\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{AUC} area under the precision-recall curve (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall.}
-\label{tab:signs-detection-results-5-percent}
-\end{table}
+\caption{Shows results of training the Opticalflow model with the 5\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{Loss} obtained loss (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall (5)\textit{AUC} area under the precision-recall curve.}
+\label{tab:opticalflow-detection-results-5-percent}
+\end{table*}
 
-% Sign detection results 10% of the dataset
-\begin{table}[!htb]
+% RGB Recurrent model results 5% of the dataset
+\begin{table*}[!htb]
 \captionsetup{font=footnotesize}
 \centering
-\begin{tabular}{ p{16em} c }
+\begin{tabular}{ p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} }
 \toprule
-\multicolumn{1}{c}{\textbf{Epoch}} & 
-	\multicolumn{1}{c}{\textbf{AUC}}\\
+\multicolumn{5}{c}{\textbf{RGB Recurrent Model Training Results with the 5\% of the dataset}} \\
+\hline
+\hline
+\textbf{Epoch}&	\textbf{Loss}	&\textbf{Precision}	&\textbf{Recall}	&\textbf{AUC} \\
+\hline
 \midrule
-\textbf{Epoch 1}&	0.3686\\
-\textbf{Epoch 2}&	0.5994\\
-\textbf{Epoch 3}&	0.1412\\
-\textbf{Epoch 4}&	0.1703\\
-\textbf{Epoch 5}&	0.1680\\
+Epoch 1&	2.7568&	0.0000e+00&	0.0000e+00&	0.0911\\
+Epoch 2&	2.6647&	0.0000e+00&	0.0000e+00&	0.1008\\
+Epoch 3&	2.6455&	0.0000e+00&	0.0000e+00&	0.1017\\
+Epoch 4&	2.6296&	0.0000e+00&	0.0000e+00&	0.1193\\
+Epoch 5&	2.6121&	0.0000e+00&	0.0000e+00&	0.1222\\
+Epoch 6&	2.6032&	0.0000e+00&	0.0000e+00&	0.1222\\
+Epoch 7&	2.5943&	0.0000e+00&	0.0000e+00&	0.1229\\
+Epoch 8&	2.5875&	0.0000e+00&	0.0000e+00&	0.1229\\
+Epoch 9&	2.5825&	0.0000e+00&	0.0000e+00&	0.1229\\
+Epoch 10&	2.5790&	0.0000e+00&	0.0000e+00&	0.1229\\
 \bottomrule
 \end{tabular}
-\caption{Shows results of training nSDm with the 10\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{AUC} area under the precision-recall curve (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall.}
-\label{tab:signs-detection-results-10-percent}
-\end{table}
+\caption{Shows results of training the RGB Recurrent model with the 5\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{Loss} obtained loss (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall (5)\textit{AUC} area under the precision-recall curve.}
+\label{tab:rgb-detection-results-5-percent}
+\end{table*}
 
-Results of infer PSL samples are presented on Figure \ref{fig:rgb-pareja-sample} we observed the noun \textit{Pareja} being correctly classified, in the other hand we observe that silent periods are being incorrectly classified as shown on Figure \ref{fig:rgb-pareja-wrong-sample}, adding one additional class for \textit{silent} could improve our classifier allowing it to infer when a consultant is to performing any action.
+% nSDmV1 model results 5% of the dataset
+\begin{table*}[!htb]
+\captionsetup{font=footnotesize}
+\centering
+\begin{tabular}{ p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} }
+\toprule
+\multicolumn{5}{c}{\textbf{nSDmV1 Training Results with the 5\% of the dataset}} \\
+\hline
+\hline
+\textbf{Epoch}&	\textbf{Loss}	&\textbf{Precision}	&\textbf{Recall}	&\textbf{AUC} \\
+\hline
+\midrule
+Epoch 1&	1.0210&	1.0000&	0.0455&	0.9929\\
+Epoch 2&	0.9115&	1.0000&	0.0455&	1.0000\\
+Epoch 3&	0.8364&	1.0000&	0.0909&	1.0000\\
+Epoch 4&	0.7748&	1.0000&	0.2273&	1.0000\\
+Epoch 5&	0.7277&	1.0000&	0.2727&	1.0000\\
+Epoch 6&	0.6929&	1.0000&	0.4545&	1.0000\\
+Epoch 7&	0.6673&	1.0000&	0.5000&	1.0000\\
+Epoch 8&	0.6479&	1.0000&	0.6364&	1.0000\\
+Epoch 9&	0.6327&	1.0000&	0.6364&	1.0000\\
+Epoch 10&	0.6207&	1.0000&	0.7273&	1.0000\\
+\bottomrule
+\end{tabular}
+\caption{Shows results of training nSDmV1 with the 5\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{Loss} obtained loss (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall (5)\textit{AUC} area under the precision-recall curve.}
+\label{tab:nSDmV1-detection-results-5-percent}
+\end{table*}
+
+% nSDmV2 model results 5% of the dataset
+\begin{table*}[!htb]
+\captionsetup{font=footnotesize}
+\centering
+\begin{tabular}{ p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} }
+\toprule
+\multicolumn{5}{c}{\textbf{nSDmV2 Training Results with the 5\% of the dataset}} \\
+\hline
+\hline
+\textbf{Epoch}&	\textbf{Loss}	&\textbf{Precision}	&\textbf{Recall}	&\textbf{AUC} \\
+\hline
+\midrule
+Epoch 1&	2.9711&	1.0000&	0.0000e+00&	0.2959\\
+Epoch 2&	2.2159&	1.0000&	0.0455&	0.6564\\
+Epoch 3&	1.7181&	1.0000&	0.0909&	0.8154\\
+Epoch 4&	1.3140&	1.0000&	0.1818&	0.9340\\
+Epoch 5&	0.9585&	1.0000&	0.3636&	0.9691\\
+Epoch 6&	0.6747&	1.0000&	0.5909&	1.0000\\
+Epoch 7&	0.4739&	1.0000&	0.8636&	1.0000\\
+Epoch 8&	0.3261&	1.0000&	0.8636&	1.0000\\
+Epoch 9&	0.2263&	1.0000&	0.9545&	1.0000\\
+Epoch 10&	0.1651&	1.0000&	1.0000&	1.0000\\
+\bottomrule
+\end{tabular}
+\caption{Shows results of training nSDmV2 with the 5\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{Loss} obtained loss (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall (5)\textit{AUC} area under the precision-recall curve.}
+\label{tab:nSDmV2-detection-results-5-percent}
+\end{table*}
+
+Even though models were trained with a small number of samples and are subject to over fitting, train results show patterns that indicates that performance will increase as we add more samples where metrics will become stronger as we add more samples to the input data pipeline, we are planning on processing more PSL samples as well as including PSL samples from external sources as described on section \ref{discussion-future-work}.
+
+Train results shows the RGB recurrent model having the lowest performance with a loss equals to 2.5790 and a AUC equals to 0.1229 and Precision and Recall equals to 0.0000 which indicates recurrent models are not learning enough features.
+In the other hand the Opticalflow model performs better with a loss equals to 0.5981 a AUC equals to 0.9877 a Precision equals to 1.000 and a Recall equals to 0.6818 which indicates features available in Lucas Kanade Opticalflow representations are learned more effectively with a 2D CNN architecture.
+2D CNN architectures show better performance than RNN architectures for detecting PSL elements.
+
+Assembled models (nSDmV1 and nSDmV2) show the highest performance where nSDmV1 reports a loss equals to 0.627 a AUC equals to 1.0000 a Precision equals to 1.0000 and a Recall equals to 0.7273 and where nSDmV2 reports a loss equals to 0.1651 a Precision equals to 1.0000 and Recall equals to 1.000 and a AUC equals to 1.0000.
+
+The results indicate assembled models perform better than individual models justifying the effort to design models that combine 2D CNN and RNN architectures.
+nSDmV1 combines the two individual models with a simple outputs average showing showing a better recall than the Opticalflow and RGB recurrent models.
+nSDmV2 shows the highest performance presumably related to the classifiers removal action applied to Opticalflow and RGB models and the subsequent concatenation which is then sent to a new classifier layer (dense layer with softmax activation) as described in section \ref{nsdmv2-architecture}.
+
+\textbf{PSL samples inference results} are presented on Figure \ref{fig:rgb-pareja-sample} we observed the noun \textit{Pareja} being correctly classified, in the other hand we observe that silent periods are being incorrectly classified as shown on Figure \ref{fig:rgb-pareja-wrong-sample}, adding one additional class for \textit{silent} could improve our classifier allowing it to infer when a consultant is performing any action.
 

### 2021-04-18 phd/thesis proposal/MainText/chapter2.tex
fixed references issues and added thesis proposal Latex files

--- 2020-02-03 article/article-tesis.tex
@@
-\documentclass[twocolumn,conference]{article}
+\chapter{Methodology} \label{chap-2}
 
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{xcolor}
-\usepackage{listings}
-\usepackage{inputenc}
-\usepackage{graphicx}
-\usepackage{authblk}
-\usepackage{enumerate}
-\usepackage{caption}
-
-\usepackage{array,booktabs,longtable,tabularx}
-\usepackage{ltablex}
-\usepackage{wrapfig,lipsum,booktabs}
-\usepackage{multirow}
-
-\providecommand{\keywords}[1]
-{
-  \small	
-  \textbf{\textit{Keywords---}} #1
-}
-
-\begin{document}
-\author[1]{Eric Raphael Huiza Pereyra}
-\affil[1]{Pontifical Catholic University of Peru}
-\affil[]{\textit{eric.huiza@pucp.edu.pe}}
-
-\author[2]{Cesar Augusto Olivares Poggi}
-\affil[2]{Pontifical Catholic University of Peru}
-\affil[]{\textit{cesar.olivares@pucp.edu.pe}}
-
-\title{%
-	\vspace{-2.0cm}
-	\textbf{Talking with signs} \\	
-	\Large \textbf{A simple method to detect nouns and numbers in a non-annotated signs language corpus}
-}
-
-\maketitle
-
-\begin{abstract}
-People with deafness or hearing disabilities who aim to use computer based systems rely on state-of-art video classification and human action recognition techniques that combine traditional movement pattern recognition and deep learning techniques.
-In this work we present a pipeline for semi-automatic video annotation applied to a non-annotated \textit{Peruvian Signs Language} (PSL) corpus along with a novel method for a progressive detection of PSL elements (nSDm).
-We produced a set of video annotations indicating signs appearances for a small set of nouns and numbers along with a \textit{labeled PSL dataset} (PSL dataset).
-A composite model obtained from the combination of a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade optical flow, and a RNN with LSTM cells trained with raw rgb frames extracted from the PSL dataset achieved 0.5994 AUC for the Precision-Recall curve on signs classification tasks, reporting state-of-art results over the PSL dataset.
-\\
-\keywords{Video Classification, Human Actions Detection, Peruvian Signs Language, Optical Flow, 2D CNN, LSTM.}
-\end{abstract}
-
-\section{Introduction}\label{intro}
-The World Health Organization (WHO) stated that 466 million people world wide have disabling hearing loss, estimating that by 2050 over 900 million people will have disabling hearing loss that will represent a global cost of 750 million dollars annually \cite{deafness_and_hearing_loss_2019}.
-
-The Peruvian Institute of Informatics and Statistics (INEI) conducted a national disabilities survey with the objective of segmenting and acquiring a better understanding about disabilities that affect the Peruvian population \cite{disabilities_survey_2012}.
-Results showed that 1.8\% of the Peruvian population suffer at least partial when not permanent deafness or hearing limitations.
-
-Peruvians with deafness or hearing limitations use the Peruvian Signs Language (PSL) as their main communication medium.
-PSL is of mandatory usage at universities and certain public institutions, henceforth the importance of designing systems that are capable to support PSL inputs and outputs.
-Furthermore, in the same way as spoken languages, signs languages also present local variations e.g.
-people who live in Lima metropolitan area are not expected to use the same set of signs as people in other parts of the territory.
-This work uses the PSL variation used in Lima due to the difficulty or inability to find datasets for other PSL variations.
-
-\textit{The Grammar and Signs research group of the Pontifical Catholic University of Peru (PUCP)} built the first PSL corpus \cite{lsp_dataset} which is publicly available at the university digital archives.
-It is important to highlight that the corpus is neither labeled or annotated and cannot be used as it is for training or testing a model.
-
-In this work we are approaching signs detection as a supervised learning task.
-Supervised learning requires labeled datasets to achieve satisfactory results during training and inference tasks.
-At the time of writing this work there were no labeled datasets available for PSL \cite{lsp_2015}.
-It configures a gap that could prevent or hinder research work on Human-Computer-Interaction at the Peruvian or Latin American space.
-
-Current advances in Computer Vision (CV) and Natural Language Processing (NLP) make it possible to conceive systems that are capable of detecting and transcribing elements of sign languages thereby improving systems accessibility for people with physical limitations.
-This work reports results of a research conducted with the goal of producing a labeled PSL dataset for a set of signs limited to nouns and numbers as well as a novel method for detecting PSL signs by answering the following research questions:
-
-\begin{enumerate}[(i)]
-\item What are the currently available techniques for producing a labeled dataset for a set of signs limited to nouns and number from the non-annotated PSL corpus?\label{q1}
-\item What are most relevant and currently available techniques for training a model with the labeled dataset described in the question above for detecting PSL nouns and numbers?\label{q2}
-\item How precise and exhaustive is the model described in the above question on the detection of PSL nouns and numbers?\label{q3}
-\end{enumerate}
-
-This work has the main objective of producing a simple method that can be used as a baseline for other researchers interested on studying signs language and their different applications on the Human-Computer-Interaction field, we believe this work will produce a positive impact on the artificial intelligence community towards an increase in the number of research works using PSL which can contribute to increasing the number of people with deafness or hearing disabilities that can use computer based systems.
-We have divided this work main objective into following specific objectives for better traceability:
-
-\begin{enumerate}[(i)]
-\item Produce a labeled PSL dataset limited to nouns and numbers\label{o1}
-\item Design and train a novel signs detection model (nSDm) for detecting signs at the labeled PSL dataset\label{o2}
-\item Determine what is the performance of nSDm in terms of precision and recall\label{o3}
-\end{enumerate}
-
-The rest of the article is organized as follows.
-In section \ref{relatedwork} we review the related work on video classification for human actions recognition using network architectures that combine CNNs, 3D CNNs and movement patterns  for better features learning, we also review state-of-art pose estimation techniques.
-In section \ref{method} we introduce nSDm describing its design and architecture.
-In section \ref{experimentation} we evaluate nSDm precision and recall and answer research questions \ref{q1},\ref{q2},\ref{q3}.
-In section \ref{datasetdesc} we describe the PSL dataset produced at PUCP.
-In section \ref{videoannot} we describe the video annotation and data pre-processing techniques applied to produce the labeled PSL dataset and finally in section \ref{conclusion} we present our conclusions and future work.
-\section{Related Work} \label{relatedwork}
-\subsection{Action Recognition}
-Human action recognition is an extensively studied field.
-Action recognition dataset like UCF101, HMDB51, THUMOS14  are available, researches tried to solve the human action recognition problem using different approaches including Optical Flow and 3D CNN \cite{qiu2017learning}.
-
-\textbf{Optical Flow,} is defined as the pattern obtained from the motion of objects, surfaces and edges in a visual scene caused by the relative motion between the observer and a scene.
-It is computed by distributing movement velocities and brightness across frames.
-It is a  key concept in action recognition from videos \cite{wang2019hallucinating}.
-Optical flow estimation is treated as an image reconstruction problem.
-Given a frame set, the optical flow is generated and allows to reconstruct one frame from the others \cite{zhu2018hidden}.
-Formally, taking the optical flow displacement field as input and training a CNN with it, then the network should have learned useful representations of the underlying motions.
-Even though Optical Flow represents the movement between a set of frames, if camera motion is considered as an action motion, it may corrupt the action classification \cite{wang2013dense}.
-Various types of camera motion can be observed in realistic videos, e.g., zooming, tilting, rotation, etc.
-
-\textbf{Motion Boundary Histogram (MBH)} is a simple an efficient way to achieve robustness during human action detection when camera movements are mixed within the recorded actions by computing derivatives separately for the horizontal and vertical components of the optical flow.
-Since MBH represents the gradient of optical flow, locally constant camera motion is removed and information about changes in the flow field is kept.
-MBH is more robust to camera motion than optical flow, thus more discriminative for action recognition.\cite{wang2013dense}.
-3D CNN are not as effective as optical flow to detect human actions on its own, 3D CNN can be trained to learn optical flow so we can avoid costly computation and storage and obtain task-specific motion representation  \cite{zhu2018hidden} and increase models performance, precision and recall on human action recognition.
-\subsection{Pose Estimation}
-Pose estimation is also an extensively studied field.
-Techniques based on key points have shown state-of-art results on human pose estimation.
-An approach on key points estimation \cite{tulsianiMalik} uses \textit{Point of View Determination} and \textit{Key Points Prediction} components.
-Point of View Determination is formulated by the prediction of three Euler angles (azimut, elevation and cyclotation) generating a global position estimate, then a local appearance is modeled by obtaining a heat map that corresponds to the spatial distribution likelihood for each key point, finally key points predictions are obtained by combining heat maps obtained in a previous stage with a conditioned likelihood at the point of view predicted in the previous stage.
-
-Key points detection methods based CNNs have received an special attention in Human Pose Detection problems.
-CNNs methods are divided in bottom-up and top-down.
-Bottom-up methods process images from low resolution  to high resolution, focusing first on detecting joints before associating them to human actions.
-Top-down methods focus first on detecting human subjects and then estimating the human pose to predict key points.
-
-The datasets MPII and COCO have been used in state-of-art methods obtaining good results\cite{XiaoWuWeiSimpleBaseline} and establishing a framework for future work in combination with classic approaches like optical flow for recognizing patterns movement between frames by increasing accuracy on key points detection.
-\subsection{Video Classification}
-\textbf{Bag of Words (BoW)} or \textbf{Bag of Visual Words (BoVW)} based on natural language processing techniques is one of the simplest and oldest local descriptor encoding strategies.
-In its simplest form, it consists of (i) clustering with k-means a collection of descriptor vectors from the training set to build so-called visual vocabulary, (ii) as signing each descriptor to its nearest cluster center from the visual dictionary, and (iii) aggregating the one-hot assignment vectors via average pooling \cite{wang2019hallucinating}, when applied to Computer Vision is a technique used to create images representations or features vectors used that can be learned by CNNs, resulting on improved images classification and video classification.
-Feature trajectory detection are much improved using statistical methods like \textbf{Fisher Vectors} obtaining better results over traditional BoW \textbf{Fussing parallel CNN.}.
-The Bag of Visual Words representation suffers from sparsity and high dimensionality, in the other hand representations obtained using the Fisher Vectors kernel are more compact and dense which results on better results for image and video classification problems.
-\section{Method}\label{method}
-\subsection{Video Annotation}\label{videoannot}
-The PSL dataset is non-annotated because there is not a direct relation between the instant when a sign is emitted and when its translation to Spanish is delivered.
-We propose a semi-automatic video annotation pipeline described in Figure \ref{fig:video-annotation-pipeline} for cleaning, pre-processing and analyzing PSL videos in order to produce an labeled PSL dataset that can be used for training nSDm using supervised learning.
-The pipeline is described in detail in sections  \ref{manual-video-cleanup}, \ref{video-pre-processing}, \ref{audio-transcription-analysis} and \ref{samples-generation}
-
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/video-annotation-pipeline.png}
-\caption{Video annotation process}
-\label{fig:video-annotation-pipeline}
-\end{figure}
-
-\subsubsection{Manual Automatic Video Clean Up}\label{manual-video-cleanup}
-The PSL recordings described on \ref{datasetdesc} contain a considerable amount of noise introduced during recording sessions.
-It makes difficult to easily find video intervals that clearly show a relation between signs emitted by the informant and the translation delivered by the translator.
-Noise factors are the following:
-\begin{itemize}
-	\item Multiple participants speaking during the session.
-	\item Conversations between participants that are not relevant to emitted sings.
-	\item High frequency of large silent periods.
-\end{itemize}
-A manual video cleanup process is required to find noise free video intervals.
-This process requires watching all the videos available at the PSL corpus for manually annotating the instant when an informant started emitting sings along with the instant when the translator delivered a translation, Table \ref{table:noise-free-video-segments} shows a manual annotation example.
-
-The recordings show the informant in two alignments (centered and left), the manual video clean up process also stores the informant alignment, table \ref{table:informant-alignment} shows the two available alignments, we use the alignment annotation later in the process during the video frames extraction to create the labeled PSL dataset.
-
-% Noise free video segment extract table
-\begin{table}[!htb]
-\captionsetup{font=footnotesize}
-\centering
-\begin{tabular}{lrrc}
-\toprule
-\multicolumn{1}{c}{\textbf{Video}} & 
-	\multicolumn{1}{c}{\textbf{Start}} &
-	\multicolumn{1}{c}{\textbf{End}} &
-	\multicolumn{1}{c}{\textbf{Alignment}}\\
-\midrule
-\multirow{5}{7.5em}{\textbf{consultant-01-session-01-part-01.mp4}} & 00:30 & 00:55 & center\\
-& 01:15 & 01:29 & center\\
-& 00:53 & 01:07 & center\\
-& 08:12 & 09:01 & center\\
-\midrule
-\multirow{5}{7.5em}{\textbf{consultant-02-session-01-part-01.mp4}} & 00:15 & 00:21 & center\\
-& 00:15 & 00:21 & center\\
-& 00:53 & 01:07 & center\\
-& 02:43 & 02:47 & center\\
-& 17:33 & 18:01 & left\\
-\bottomrule
-\end{tabular}
-\captionsetup{font=footnotesize}
-\caption{Noise free video segments extract} \label{table:noise-free-video-segments}
-\end{table}
-
-% Informat aligment
-\begin{table}[!htb]
-\captionsetup{font=footnotesize}
-\centering
-\begin{tabular}{cc}
-\toprule
-\multicolumn{1}{c}{\textbf{Center Aligned}} & 
-	\multicolumn{1}{c}{\textbf{Left Aligned}}\\
-\midrule
-\includegraphics{images/informant-center-alignment.png}& \includegraphics{images/informant-left-alignment.png}\\
-\bottomrule
-\end{tabular}
-\caption{Informant Alignment} \label{table:informant-alignment}
-\end{table}
-
-\subsubsection{Video Pre-Processing}\label{video-pre-processing}
-Non-annotated PSL videos require processing before any metadata can be extracted, we propose a sequence of pre-processing tasks that take advantage of the annotation generated on \ref{manual-video-cleanup}.
-A \textbf{video splitting processor} generates a set of video chunks using the \textit{ffmpeg} multimedia framework and stores produced video chunks in Amazon S3 for later usage.
-Audio within video chunks is then transcribed by an \textbf{audio transcription processor}, using the Amazon Transcription service, we selected the Amazon Transcription service because it provides an accurate mapping between audio participants and transcribed words along with useful metadata that describes the start and end time when words are pronounced by the translator.
-
-At the moment of writing this work Amazon Transcription service only supported Spain and US Spanish.
-This caused certain words that are specific for Peruvian Spanish not being fully recognized, in order to improve transcription accuracy we built a custom vocabulary containing Peruvian expressions which improved  Peruvian words recognition, for the matters of this work Peruvian words that remained unrecognized were omitted and not processed.
-
-\subsubsection{Audio Transcription Analysis}\label{audio-transcription-analysis}
-Audio transcription requires additional processing in order to produce useful information that leads to a successful PSL signs detection.
-Bag of Embedding Words (BoEW) is a widely used technique on Natural Language Processing tasks providing a easy and flexible way to list the most relevant words based on frequency.
-This work is focused on detecting nouns and numbers (our method is designed to be progressively improved to handle a wider set of PSL elements) assuming that nouns (numbers are a subset of nouns) suffer less variations in spoken Spanish than verbs, pronouns, adverbs and adjectives, and provide more semantic value than conjunctions, prepositions and interjections.
-
-\begin{table}[!htb]
-\captionsetup{size=footnotesize}
-\begin{tabular}{ p{16em} c}
-\toprule
-\multicolumn{1}{c}{\textbf{Token}} & 
-	\multicolumn{1}{c}{\textbf{Frequency}}\\
-\midrule
-\textbf{pareja}&	40\\
-\textbf{cosas}&	30\\
-\textbf{cine}&	20\\
-\textbf{noche}&	20\\
-\textbf{terror}&	10\\
-\textbf{parque}&	10\\
-\textbf{casa}&	10\\
-\textbf{montón}&	10\\
-\textbf{apariciones}&	10\\
-\textbf{fantasmas}&	10\\
-\textbf{dos}&	10\\
-\bottomrule
-\end{tabular}
-\caption{Most relevant tokens detection frequency in the PSL dataset} \label{tab:token-freq}
-\end{table}
-
-We used Amazon comprehend for text analysis, specifically the syntax detection functionality which will provide a comprehensive list of detected language elements along with a score from 0.0 to 1.0 indicating the detection accuracy, we have selected the ones that have at least a 0.8 accuracy score and omitted the rest, this process was automated using a \textbf{transcription detection processor} which uses BoEW to provide a list of most relevant nouns and numbers based on appearance frequency.
-Table \ref{tab:token-freq} shows a list of nouns and numbers and their frequencies in the PSL dataset.
-
-Once a weighted list of nouns and numbers is generated a mapping showing when nouns and numbers appear in videos is required, moving forward called \textbf{Samples Metadata}.
-Table \ref{tab:token-video-mapping} shows mapping metadata extracted from PSL.
-
-\subsubsection{Samples Generation}\label{samples-generation} 
-Our method requires PSL elements to be represented as a set of RGB frames and a calculated Optical Flow using the Lucas-Kanade method, both representations are inputs of two different models as presented on \ref{nsdm}.
-
-\textbf{Translation Delay Factor: } The difference in time between the instant when a sign is emitted and when a translation for that given sign is delivered is uncertain, we are calling that uncertainty \textit{the translation delay factor}, we are trying to approximate it using a constant value, we chose a three seconds translation delay factor assuming that most of the translations will occur between three seconds after a sign is emitted.
-
-A \textbf{RGB Samples generation processor} uses samples metadata in combination with the translation delay factor to determine frames that represent a given PSL element.
-We use \textit{OpenCV} to extract frames and store them following a hierarchical folder structure (listing \ref{list:rgb-samples-folders}) that nSDm data loaders will use to feed data into the RGB branch in the nSDm model architecture \ref{nsdm-architecture} during training and testing.
-
-An \textbf{Optical Flow Samples generation processor} uses video frames and the hierarchical folder structure generated by the RGB samples generation processor to calculate an Optical Flow representation for PSL elements and store them in a hierarchical folder structure that will also be used by the nSDm data loaders to feed the optical flow branch on the nSDm model architecture \ref{nsdm-architecture} during training and testing.
-We selected optical flow as a samples generation strategy due to its ability to represent movement traces from previous frames.
-It is particular useful for representing body movement patterns executed by informant while emitting a PSL sign.
-A PSL sign is made up of different body movements including: elbow, arms, neck, eyes, shoulders and hands, which are performed quickly, a way to detect movement traces between frames allows to generate a single image representation of all movement involved on a sign emission(Figure \ref{fig:opticalflow-two}).
-
-\begin{table}[!htb]
-\captionsetup{font=footnotesize}
-\centering
-\begin{tabular}{ l p{10em} r r }
-\toprule
-\multicolumn{1}{c}{\textbf{Token}} & 
-	\multicolumn{1}{c}{\textbf{Video}} &
-	\multicolumn{1}{c}{\textbf{Start}} &
-	\multicolumn{1}{c}{\textbf{End}}\\
-\midrule
-\textbf{cine}&	consultant-02-session-01-part-01-00.mp4&	4.19&	4.75\\
-\textbf{cine}&	consultant-02-session-01-part-01-01.mp4&	1.19&	1.75\\
-\textbf{terror}&	consultant-02-session-01-part-01-01.mp4&	3.82&	4.4\\
-\textbf{parque}&	consultant-02-session-01-part-01-03.mp4&	8.97&	9.3\\
-\textbf{casa}&	consultant-02-session-01-part-01-03.mp4&	10.12&	10.57\\
-\textbf{pareja}&	consultant-02-session-01-part-01-04.mp4&	3.91&	4.36\\
-\textbf{noche}&	consultant-02-session-01-part-01-04.mp4&	4.49&	4.92\\
-\textbf{noche}&	consultant-02-session-01-part-01-04.mp4&	7.91&	8.2\\
-\textbf{mont\'on}&	consultant-02-session-01-part-01-04.mp4&	8.5&	8.78\\
-\textbf{cosas}&	consultant-02-session-01-part-01-04.mp4&	8.88&	9.38\\
-\bottomrule
-\end{tabular}
-%\end{tabular}}
-\caption{Shows metadata extacted from the PSL dataset: (1)\textit{Token} could be a noun or a number (2)\textit{Video Path} shows the video where the token was detected (3)\textit{Start Time} time when the token reproduction starts (4)\textit{End Time} time when the token reproduction ends.}
-\label{tab:token-video-mapping}
-\end{table}
-
-\subsection{Novel Signs Detection Model (nSDm)}\label{nsdm}
-We propose a novel model for signs detection that combine two neural networks architectures, a CNN and a RNN with the objective to learn visual features like edges, corners and ridges (CNN) and at the same time patterns learned from a time based series of inputs (RNN).
-CNN network receives optical flow inputs and the RNN branch receives RGB frames extracted from the labeled PSL dataset described in \ref{videoannot}.
-
-For this work we selected the Tensorflow/Keras functional API for its ability to define combined models along with a versatile data extraction and transformation layer.
-More details and usage are provided in sections \ref{nsdm-architecture} and \ref{new-videos-testing}
-
-Implementation details can be found at https://github.com/erichuizapucp/signs-recognition
-\subsubsection{Architecture}\label{nsdm-architecture}
-A series of models stacked in a composed architecture focused on learning visual and time series features using two network architecture flavors enforced with a flexible data input pipeline for data feeding as shown in Figure \ref{fig:ndsm-architecture}, transformation and normalization.
-
-The \textbf{Input pipeline} accesses the labeled PSL samples and applies transformations preparing the data for upper layers, transformations were applied to both optical flow and RGB frames, PSL samples were resized to be compliant with ImageNet pre-trained models using a 224 by 224 shape and three channels for color images, data augmentation transformation were not applied due to the nature of the experiment where samples were captured using similar light conditions and camera orientations, PSL samples were transformed to tensors and normalized to floats in the [0, 1] interval, finally the transformed and normalized versions of optical flow and RGB samples were tight together in a tensor structure along with the label for usage at upper layers.
-\begin{table}[!htb]
-\captionsetup{size=footnotesize}
-\begin{tabular}{ p{16em} c}
-\toprule
-\multicolumn{1}{c}{\textbf{Hyper parameter}} & 
-	\multicolumn{1}{c}{\textbf{Value}}\\
-\midrule
-\textbf{Learning Rate}&	0.001\\
-\textbf{No Epochs}&	5\\
-\textbf{No Steps per Epoch}&	5\\
-\textbf{Batch Size}&	10\\
-\textbf{Shuffle Buffer Size}&	50\\
-\bottomrule
-\end{tabular}
-\caption{nSDm training hyper parameters} \label{tab:nSDm-hyperparams}
-\end{table}
-
-The \textbf{Wrapper Layer} wraps a CNN network designed for learning visual features from optical flow and a RNN network designed for learning features from series of video frames, the layer receives data from the input pipeline then splits it into optical flow and RGB samples and passes them to the CNN and RNN networks respectively.
-The logits produced are then concatenated and passed to the dense layer.
-
-The dense layer uses a softmax activation function to convert logits into probabilities used for a correct sign classification.
-Hyper parameters used during the training process are listed on Table \ref{tab:nSDm-hyperparams}
-
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/nsdm-model-architecture.png}
-\caption{nSDm model architecture}
-\label{fig:ndsm-architecture}
-\end{figure}
-
-The \textbf{Opticalflow model} described in Figure \ref{fig:opticalflow-architecture} receives decoded opticalflow samples, it uses a Resnet152 backbone pre-trained with ImageNet, we used a fine tuning transfer learning approach, the backbone produces a 7x7x2048 output that then is passed to a Global Average Pooling layer for obtaining a flattened output of 1x1x2048 which is then passed to a dense layer for logits computation.
-
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/opticalflow-model-architecture.png}
-\caption{Opticalflow model architecture}
-\label{fig:opticalflow-architecture}
-\end{figure}
-
-The \textbf{Recurrent RGB model} described in Figure \ref{fig:rgb-architecture} receives a sequence of decoded video frames, it uses a \textit{many-to-one} architecture with LSTM cells that hold state of 2048 units length, the output produced by the recurrent layers is then passed to a dense layer for logits computation.
-
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/recurrent-rgb-model-architecture.png}
-\caption{Recurrent RGB model architecture}
-\label{fig:rgb-architecture}
-\end{figure}
-
-Logits produced by the opticalflow and recurrent RGB models are concatenated and then passed to a dense layer with a softmax activation function producing PSL signs detection probabilities.
-
-\subsection{Sign detection testing}\label{new-videos-testing}
-We selected videos for consultant 24 which will be described in \ref{datasetdesc} for testing nSDm (the model was not trained with consultant 24 videos), we proceed to extract RGB frames for video fragments of five seconds length moving ahead one second on each loop until the end of the video that way we can reduce the number of signs cut or incomplete between video fragments.
-Optical flow and RGB frames generators feed nSDm for detecting signs \ref{nsdm} present in the video fragment.
-
-A software component called Video Updater uses video metadata produced by the RGB Frames Generator like start and end time in seconds and detected signs for annotating the video adding a mask that highlights detected signs.
-We used OpenCV for incorporating masks to test videos, Figure \ref{fig:video-testing-architecture} shows the signs detection testing architecture.
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/new-video-test-architecture.png}
-\caption{Signs detection testing process architecture}
-\label{fig:video-testing-architecture}
-\end{figure}
-
-\section{Experimentation}\label{experimentation}
-\subsection{Dataset Description \cite{lsp_dataset}}\label{datasetdesc}
+\section{Available Datasets}\label{available-datasets}
 The PSL dataset was developed by the PUCP Grammar and Signs research group in 2014 and consists in a set of videos recorded during the interviews of 24 individuals, 12 male and 12 female informants, all of them are Lima Peru residents and reported to be born with a permanent deafness condition or acquired the condition before the acquisition of Spanish.
@@
 
-\textbf{Recording Session 3}: A PSL \cite{lsp_2015} conversation facilitated by the coordinator happening between the informant and the translator.\\
+\textbf{Recording Session 3}: A PSL \cite{lsp_2015} conversation facilitated by the coordinator happening between the informant and the translator.
 
@@
 
-\subsection{Video Annotation Results}\label{video-annotation-results}
-The video annotation pipeline described on \ref{videoannot} produced an annotated PSL dataset suitable for using it in a supervised learning experiment.
-The annotated dataset is divided in two main parts (RGB and Optical Flow samples)
-
-\textbf{RGB Samples folder structure} is a hierarchical folder structure where each detected noun or number is represented as a first level folder, all instances of a detected sign received an identifier which is an auto incremental integer, detected instances represent the second level folders, detected signs video represent the third level, Figure \ref{fig:rgb-two}.
-\begin{lstlisting}[caption=RGB Samples Folder Structure example, basicstyle=\ttfamily\small]
-L1:dos
-	L2:1
-		L3:rgb-frame01.jpg
-		L3:rgb-frame02.jpg
-		L3:rgb-frame03.jpg
-		...
-		L3:rgb-frame08.jpg
-L1:cine
-	L2:1
-		L3:rgb-frame01.jpg
-		L3:rgb-frame02.jpg
-		L3:rgb-frame03.jpg
-		...
-		L3:rgb-frame15.jpg
-	L2:2
-		L3:rgb-frame01.jpg
-		L3:rgb-frame02.jpg
-		L3:rgb-frame03.jpg
-		...
-		L3:rgb-frame11.jpg
-\end{lstlisting}\label{list:rgb-samples-folders}
-
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/dos-rgb-frames.png}
-\caption{PSL number "Two" RGB representation}
-\label{fig:rgb-two}
-\end{figure}
-
-\textbf{Optical Flow samples folder structure} is a hierarchical folder structure based on the RGB samples folder structure, it is a more simple based on two levels instead of three, the nature of Optical Flow of tracing movement between frames allow to produce a single image for each detected PSL element instance (listing \ref{list:opticalflow-samples-folders}), Figure \ref{fig:opticalflow-two} shows an example of an optical flow generated sample
-\begin{lstlisting}[caption=Optical Flow Samples Folder Structure example, basicstyle=\ttfamily\small]
-L1:dos
-	L2:oflow-dos-01.jpg
-L1:cine
-	L2:oflow-cine-01.jpg
-	L2:oflow-cine-02.jpg
-\end{lstlisting}\label{list:opticalflow-samples-folders}
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/dos-opticalflow.jpg}
-\caption{PSL number "Two" OpticalFlow representation}
-\label{fig:opticalflow-two}
-\end{figure}
-
-\subsection{Sign detection results}
-We trained nSDm with the train PSL dataset and validated it with the validation PSL dataset, we trained the model  for five epochs using five and ten percent of the dataset obtaining the results in Tables \ref{tab:signs-detection-results-5-percent} and \ref{tab:signs-detection-results-10-percent}
-
-Train results shows a trend indicating that the more PSL samples we use the nSDm performance improves highlighting the need for processing additional videos at the PSL corpus.
-% Sign detection results 5% of the dataset
-\begin{table}[!htb]
-\captionsetup{font=footnotesize}
-\centering
-\begin{tabular}{ p{16em} c }
-\toprule
-\multicolumn{1}{c}{\textbf{Epoch}} & 
-	\multicolumn{1}{c}{\textbf{AUC}}\\
-\midrule
-\textbf{Epoch 1}&	0.0708\\
-\textbf{Epoch 2}&	0.0539\\
-\textbf{Epoch 3}&	0.0558\\
-\textbf{Epoch 4}&	0.1175\\
-\textbf{Epoch 5}&	0.0368\\
-\bottomrule
-\end{tabular}
-\caption{Shows results of training nSDm with the 5\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{AUC} area under the precision-recall curve (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall.}
-\label{tab:signs-detection-results-5-percent}
-\end{table}
-
-% Sign detection results 10% of the dataset
-\begin{table}[!htb]
-\captionsetup{font=footnotesize}
-\centering
-\begin{tabular}{ p{16em} c }
-\toprule
-\multicolumn{1}{c}{\textbf{Epoch}} & 
-	\multicolumn{1}{c}{\textbf{AUC}}\\
-\midrule
-\textbf{Epoch 1}&	0.3686\\
-\textbf{Epoch 2}&	0.5994\\
-\textbf{Epoch 3}&	0.1412\\
-\textbf{Epoch 4}&	0.1703\\
-\textbf{Epoch 5}&	0.1680\\
-\bottomrule
-\end{tabular}
-\caption{Shows results of training nSDm with the 10\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{AUC} area under the precision-recall curve (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall.}
-\label{tab:signs-detection-results-10-percent}
-\end{table}
-
-Results of infer PSL samples are presented on Figure \ref{fig:rgb-pareja-sample} we observed the noun \textit{Pareja} being correctly classified, in the other hand we observe that silent periods are being incorrectly classified as shown on Figure \ref{fig:rgb-pareja-wrong-sample}, adding one additional class for \textit{silent} could improve our classifier allowing it to infer when a consultant is to performing any action.
-
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/pareja-rgb-frames.png}
-\caption{PSL noun "Pareja" RGB Frames correctly classified}
-\label{fig:rgb-pareja-sample}
-\end{figure}
-
-\begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/pareja-rgb-frame-wrong.png}
-\caption{PSL noun "Pareja" RGB Frames incorrectly recognized. The classifier is incorrectly classifying silent periods where the consultant is not performing any action}
-\label{fig:rgb-pareja-wrong-sample}
-\end{figure}
-
-\section{Discussion and Future Work}\label{discussion-future-work}
-\begin{enumerate}[•]
-\item The ability to detect all participants in recording sessions individually is mandatory for a successful video annotation pipelines based on video transcriptions as the one we presented at \ref{videoannot}.
-Future work about annotating videos based on patterns recognized in video content itself using non supervised learning approaches may improve the generated dataset driving to better results.
-\item Existing video annotation pipelines require large amounts of manual annotation work, generally for cleaning and validation tasks.
-Even though several steps within the overall annotation process can be fully automated human intervention for ensuring the quality of the generated dataset.
-Future work that use Reinforcement Learning for training an agent trained to clean and validate non-annotated videos before and after they are passed through the annotation pipeline may minimize or remove the need of human intervention during video annotation tasks.
-\item Not all the corporal movements executed by a PSL consultant are valid PSL signs or elements.
-Not handling these exceptions correctly could affect greatly the nSDm performance.
-Future work that add support for an additional class called \textit{invalid} and labeling invalid signs or corporal movements could help nSDm to learn invalid PSL actions and obtain better results.
-\item Optical flow represents movement patterns between video frames allowing models to learn these given movement patterns.
-However, Optical flow could consider camera unintentional movements and camera focus switches incorrectly as valid movement patterns between frames.
-Future work that adds Motion Boundary Histogram (MBH) \cite{wang2013dense} support to the optical flow gradient could add robustness to the ability to detect camera movements driving to a better optical flow samples generation.
-\item nSDm internal models (CNN and RNN models) as described on \ref{nsdm-architecture} can be trained independently using different hyper parameters, optimizers and loss functions.
-Pre-trained models will boost the nSDm training process enabling future works to use PSL to train nSDm using more epochs, data and being more flexible on using single models (CNN or RNN) or combining them as in nSDm.
-\item Improving they way people with deafness or hearing disabilities interact with computers requires delivering an output after PSL signs are detected.
-Future work that use Auto Encoder or Adversarial Generative Networks can be used to generate output in the form of signs that are recognizable completing the cycle of a bidirectional communication.
-\item Training nSDm with more PSL samples greatly improves its performance. Future work on annotating more videos including PSL content will increase nSDm performance.
-
-\end{enumerate}
-
-\section{Conclusion}\label{conclusion}
-\begin{enumerate}[•]
-\item Video annotation pipelines based on video transcriptions require recognizing all video participants for a correct video annotation.
-\item The \textit{Delay Factor} introduced noise to the generated labeled PSL dataset. Post annotation human intervention is needed to reduce the delay factor effect.
-\item Annotating more PSL videos and training nSDm with these new generated samples increases the model performance.
-\item Optical flow successfully represents PSL consultant corporal movements over time.
-\item Transfering weights learned a ResNet152 model trained with ImageNet influence positively Optical flow patterns learning.
-\item Using optical flow for movement pattern representation allow capturing all actions and movements executed by a PSL consultant.
-\item RGB video frames extracted from video fragments allow networks to learn sequences of movement over time.
-\item Using LSTM cells allow successfully usage of larger PSL sequences (a larger number of frames) due to its better handling of the vanishing gradients.
-\item Combining a CNN and a RNN in a composite model improves signs detection performance, while CNNs learns visual features in optical flow samples RNNs learns sequences of actions, networks outputs can be pondered for training a better classifier.
-\item The area under the precision-recall curve allow measuring how well is nSDm detecting because it summarizes the trade-off between the true positive signs rate and the predicted signs.
-\end{enumerate}
-
-\bibliographystyle{ieeetr}
-\bibliography{References}
-
-\end{document}
+Other existing Sing Language Recognition are publicly available.
+These datasets can be characterized as isolated or continuous, taking into account whether annotation are provided at the gloss (fine-grained) or the sentence (coarse-grained) levels.
+Additionally, they can be divided into Signer Dependent (SD) and Signer Independent (SI) ones, based on the defined evaluation scheme.
+In particular, in the SI datasets a signer cannot be present in both the training and the test set.
+In Table \ref{tab:public-large-datasets}, the following most widely known public SLR datasets, along with their main characteristics.

### 2021-04-18 phd/thesis proposal/Preamble/mydefinitions.tex
fixed references issues and added thesis proposal Latex files

%----------------------------------------------------------------------------------------
%	VALUES FOR THE THESIS
%----------------------------------------------------------------------------------------

\newcommand{\name}{Eric Raphael Huiza Pereyra} % Author name
\newcommand{\thesistitle}{Self supervised intelligent user experiences for an improved human computer interaction applied to Peruvian deaf population} % Title of the thesis
\newcommand{\submissiondate}{February, 2021} % Submission date "Month, year"
\newcommand{\supervisor}{César Armando Beltrán Castañon} % Supervisor name
%\newcommand{\cosupervisor}{C.~O'Supervisor} % Co-Supervisor name, comment this line if there is none%

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY STYLE (pick the style you want)
%----------------------------------------------------------------------------------------

\usepackage[square, numbers, sort&compress]{natbib} % for bibliography - Square brackets, citing references with numbers, citations sorted by appearance in the text and compressed (as in [4-7])
%\usepackage[longnamesfirst,round]{natbib} % Natural Sciences bibliography

\bibliographystyle{Preamble/physics_bibstyle} % You may use a different style adapted to your field
%\bibliographystyle{unsrtnat} % You may use a different style adapted to your field

%----------------------------------------------------------------------------------------
%	YOUR PACKAGES (be careful of package interaction)
%----------------------------------------------------------------------------------------

\usepackage{amsthm,amsmath,amssymb,amsfonts,bbm}% Math symbols

%----------------------------------------------------------------------------------------
%	YOUR DEFINITIONS AND COMMANDS
%----------------------------------------------------------------------------------------

% New Commands
\newcommand{\bea}{\begin{eqnarray}} % Shortcut for equation arrays
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\e}[1]{\times 10^{#1}}  % Powers of 10 notation

% Defining a theorem box for Criteria
\newtheorem{critere}{Criterion}
\newcommand{\crit}[2]{
\begin{center}  
\fbox{ \begin{minipage}[c]{0.9 \textwidth}
\begin{critere}
\textbf{\textup{ #1}} --- #2
\end{critere}
\end{minipage}  } \end{center}
}

### 2021-04-18 phd/thesis proposal/Thesis_proposal.tex
fixed references issues and added thesis proposal Latex files

\documentclass[12pt, twoside]{book} % 12 pt font, two-sided book style
\usepackage[a4paper, includehead, headheight=0.6cm, inner=3cm ,outer=2.5cm, top=2.5 cm, bottom=2.5cm]{geometry}  % Changing size of document
\usepackage[english]{babel} % The document is in English
\usepackage[utf8]{inputenc} % UTF8 encoding
\usepackage[T1]{fontenc} % Font encoding

\usepackage{graphicx} % For including images
\graphicspath{{./Images/}} % Specifies the directory where pictures are stored
\usepackage{longtable} % tables that can span several pages
\usepackage[bf]{caption} % caption: FIG in bold
\usepackage{fancyhdr} % For the headers

\newcommand{\numberedchapter}{ % Preparation for numbered chapters
	\cleardoublepage % To make sure the previous headers are passed
	\fancyhead[RE]{{\bfseries \leftmark}}% Headers for left pages
	\fancyhead[LO]{{\bfseries \rightmark}}}% Headers for right pages
\newcommand{\unnumberedchapter}[1]{ % Preparation for unnumbered chapters
	\cleardoublepage % To make sure the previous headers are passed
	\addcontentsline{toc}{chapter}{#1} % Also adds the chapter name to the Contents
	\fancyhead[RE]{{\bfseries #1}} % Headers for left pages
	\fancyhead[LO]{}}%Headers for right pages

\usepackage{emptypage} % No headers on an empty page

\usepackage{eso-pic} % For the background picture on the title page

\usepackage{hyperref} % Adds clickable links at references
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{array}
\usepackage{colortbl}

%----------------------------------------------------------------------------------------
%	ADD YOUR CUSTOM VALUES, COMMANDS AND PACKAGES
%----------------------------------------------------------------------------------------

% Open Preamble/mydefinitions.tex and enter some values (name, thesis title...) 
% and include your own custom LaTeX functions and packages

\input{Preamble/mydefinitions}

\begin{document}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\pagestyle{empty} % No page numbers
\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the preamble pages

\begin{titlepage}
\begin{center}
\vfill
{\large \scshape Pontifical Catholic University of Peru\\Postgraduate School}\\[1.4cm]
{\Large Doctoral Thesis Proposal}\\[0.5cm]
\rule{\textwidth}{1.5pt}\\[0cm]
{\huge \bfseries \thesistitle \par \ }\\[-0.5cm]
\rule{\textwidth}{1.5pt}\\[2.5cm]
by\\[1cm]
{\large \bfseries\name}\\[2cm]
Supervisor\\[1cm]
{\large \textbf{\supervisor}} \\ 
\vfill
%\ifx\cosupervisor\undefined\else{\hfill \large Co-Supervisor: \textbf{\cosupervisor}} \\ \fi%
\vspace{1cm}
\submissiondate
\end{center}
\end{titlepage}

%----------------------------------------------------------------------------------------
%	PREAMBLE PAGES (comment out unnecessary pages)
%----------------------------------------------------------------------------------------

\pagestyle{fancy} % Changes the headers
\renewcommand{\chaptermark}[1]{ \markboth{#1}{}} % Getting the chapter name right
\renewcommand{\sectionmark}[1]{\markright{\thesection\; #1}} % Getting the section name right
\fancyhf{}% Clears header and footer
\fancyhead[RO,LE]{\thepage} % page number on the outside of headers

\input{Preamble/abstract}

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES
%----------------------------------------------------------------------------------------

\unnumberedchapter{Contents}
\tableofcontents % Write out the Table of Contents

%----------------------------------------------------------------------------------------
%	THESIS MAIN TEXT - CHAPTERS
%----------------------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\mainmatter % Begin numeric (1,2,3...) page numbering

\numberedchapter
\input{MainText/chapter1} % Import your chapters here
\input{MainText/chapter2}
\pagestyle{plain}
\input{MainText/chapter3} 
%\input{MainText/chapter4} 
%\input{MainText/chapter5}

%----------------------------------------------------------------------------------------
%	APPENDICES
%----------------------------------------------------------------------------------------

%\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
%\appendix % Starts of appendices
%
%\numberedchapter
%\input{MainText/appendixA}
%\input{MainText/appendixB}
%\input{MainText/appendixC}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\unnumberedchapter{Bibliography} % Title of the unnumbered chapter
\bibliography{Preamble/Thesis_bibliography} % The references information are stored in the file named "Thesis_bibliography.bib"

\end{document}

### 2021-04-18 src/main/app/video/video_evaluator.py

@@ -6,7 +6,7 @@ import tensorflow as tf
 from pre_processing.video.opticalflow_samples_extractor import OpticalflowSamplesExtractor
 from pre_processing.video.rgb_samples_extractor import RGBSamplesExtractor
 
-from legacy.nsdm_v2_builder import NSDMV2ModelBuilder
+from learning.model.legacy.nsdm_v2_builder import NSDMV2ModelBuilder
 from learning.dataset.prepare.combined_dataset_preparer import CombinedDatasetPreparer
 
 from learning.common.labels import SIGNS_CLASSES

### 2021-04-18 src/main/learning/model/custom/nsdm_v3.py

@@ -1,0 +1,12 @@
+import tensorflow as tf
+
+
+class NSDMV3(tf.keras.Model):
+    def __init__(self):
+        super(NSDMV3, self).__init__()
+
+    def get_config(self):
+        pass
+
+    def call(self, inputs, training=None, mask=None):
+        pass

### 2021-04-18 src/main/learning/model/legacy/nsdm_builder.py

@@ -1,7 +1,7 @@
 import logging
 
 from tensorflow.keras.models import Model
-from legacy.base_model_builder import BaseModelBuilder
+from learning.model.legacy.base_model_builder import BaseModelBuilder
 from tensorflow.keras import Input
 from tensorflow.keras.layers import Average
 from learning.common.model_type import NSDM

### 2021-04-18 src/main/learning/model/legacy/nsdm_v2_builder.py

@@ -1,7 +1,7 @@
 import logging
 
 from tensorflow.keras.models import Model, Sequential
-from legacy.nsdm_builder import NSDMModelBuilder
+from learning.model.legacy.nsdm_builder import NSDMModelBuilder
 from tensorflow.keras import Input
 from tensorflow.keras.layers import Dense, Concatenate
 from learning.common.model_type import NSDMV2

### 2021-04-18 src/main/learning/model/legacy/opticalflow_model_builder.py

@@ -1,6 +1,6 @@
 import logging
 
-from legacy.base_model_builder import BaseModelBuilder
+from learning.model.legacy.base_model_builder import BaseModelBuilder
 from tensorflow.keras.applications.resnet_v2 import ResNet152V2
 from tensorflow.keras.models import Model
 from tensorflow.keras.layers import Dense, GlobalAveragePooling2D

### 2021-04-18 src/main/learning/model/legacy/rgb_recurrent_model_builder.py

@@ -1,6 +1,6 @@
 import logging
 
-from legacy.base_model_builder import BaseModelBuilder
+from learning.model.legacy.base_model_builder import BaseModelBuilder
 from tensorflow.keras.models import Model
 from tensorflow.keras import Input
 from tensorflow.keras.layers import Masking, Bidirectional, Dense, LSTM

### 2021-05-09 phd/thesis plan/MainText/chapter1.tex
made updates for better LaTeX files organization and some minor updates on the opticalflow and rgb datasets preparers

\chapter{Introduction and planning} \label{ch-1}

\section{Introduction}

The World Health Organization (WHO) stated that 466 million people worldwide have disabling hearing loss posing a global cost of 750 million dollars annually for unaddressed cases, estimating that by 2050 over 900 million people will have this condition. Current estimates suggest 83\% gap in hearing aid needed and use i.e., only 17\% of those who could benefit from using some kind of hearing aid\cite{deafness_and_hearing_loss_2020}, configuring a real need for investing on inclusive computing system that help on decreasing the gap and enable people with the tools to successfully interact with computer systems.

The Peruvian Institute of Informatics and Statistics (INEI) conducted a national disabilities survey with the objective of developing a better understanding about disabilities that affect the Peruvian population as well as their prevalence and posterior segmentation\cite{disabilities_survey_2012}. Results showed that 1.8\% of the Peruvian population suffer at least partially when not permanent deafness or hearing limitations indicating that approximately half a million people in the country is subject to have impediments or difficulties while performing day to day tasks that require the assistance of computer based systems. Peruvians with deafness or hearing limitations use the Peruvian Signs Language (PSL) as their main communication medium. PSL is mandatory at universities and certain public institutions as well as some official acts but is not common or nonexistent in the majority of venues and acts, henceforth the importance of designing systems that are capable to synthesize PSL inputs and simulate or generate PSL compatible outputs in a natural way. Furthermore, it is worth to indicate that in the same way as spoken languages, signs languages present local variations i.e. people living in Lima metropolitan area are not expected to use the same set of signs as people in other parts of the territory thereby it is correct to assume the existence of local PSL variations that enrich the language but at the same time add complexity towards to standardization, description and processing. This work will use a dataset containing PSL recordings with the variation used in the Lima metropolitan area due to the difficulty or inability to find datasets for other PSL variations.

The Grammar and Signs research group of the Pontifical Catholic University of Peru (PUCP) built the first PSL dataset  which is publicly available at the university digital archives. The dataset consists of a set of recordings following a method that involves a signer, a translator and a coordinator. It is important to highlight that the dataset is neither labeled or annotated and cannot be used as it is for training or testing a machine learning model. This work elaborates a method for extracting features and learning PSL elements using a self supervised approach.

A grammatically description of the PSL elements including but not limited to phonology, dactylology, numbers, common expressions, gender, verb tenses, question forms, nouns and adjectives then continued with a definition of PSL manual and non manual features where manual features refers to hand gestures and non manual refers to other body parts sections like neck, shoulders and the head. It is important to establish the relation between grammatical rules and manual or non manual features for which we will use the HamNoSys / SiGML notation in the context of an isolated signs detection method which is then extended to a continuous signs detection method approached as a self supervised learning task focused on learning features from the PSL dataset using representational autoencoders and energy functions (LeCun et al. 2020) while reconstructing PSL elements.

A conversational user interface between deaf and a computing system enabled by a graphical gestural user interface powered by HAmin for a 3D human simulation achieved by synthesizing PSL to Spanish and then Spanish to a PSL bundled into a mobile application that can be installed as a standalone application or an extension to existing application or services.

This doctoral project pursues the main objective of building "A novel solution for deaf users who aim to improve their human-computer interaction enabled by state of art self supervised learning and computer vision methods." and the following specific objectives:

\begin{itemize}
\item Produce an efficient self supervised method for annotating small non annotated signs language datasets applied to the Peruvian signs language corpus produced by the Grammar and Signs research group of the Pontifical Catholic University of Peru generalizable to other sign languages datasets.
\item Produce a Nielsen heuristics aware graphical gestural user interface capable of enabling an effective bidirectional human computer interaction between deaf users and a device powered by an algorithm trained with the dataset produced in the objective above.
\item Produce a prototype installable and distributable application powered by the algorithm and user interface produced in objectives 1 and 2.
\end{itemize}

The rest of the document is organized as follows. In section \ref{literature-review} we review the related work and state-of-art on sign languages recognition, self supervised learning applied to images and videos and graphical gestural user interfaces. In sections \ref{research-schedule} and \ref{research-budget} we describe the schedule and budget plan for executing the set of tasks needed to complete the doctoral project and finally in section \ref{available-datasets} we present the available datasets for signs language recognition and their details.

\section{Literature Review}\label{literature-review}
\subsection{Signs Language Recognition and annotation}
Signs Language Recognition (SLR) are roughly divided in two categories Isolated SLR and Continuous SLR\cite{adaloglou2020comprehensive} figure \ref{fig-slr} shows tasks involved on SLR.

\subsubsection{Isolated Signs Language Recognition}
Methods using this approach will assume individual and well defined glosses that are recognized in a series of video frames. Recognizing gestures is a difficult task, due to intrapersonal and interpersonal variations in performing them. In \cite{kennaway2015avatar} a new descriptive language called SiGML (Signing Gesture Markup Language) is presented, SiGML was developed from HamNoSys, the Hamburg Notation System. HamNoSys is a notation for recording sign language gestures, developed by researchers on sign language at the IDGS (Institut für Deutsche Ge-bädensprache) at the University of Hamburg, it is not specific to any one sign language, but is intended to cover all signing gestures in all sign languages. It records sign languages elements in terms of hand shape, hand location and hand movement and other special terms for non manual gestures. Signs are not holistic gestures but are rather analyzable, as a combination of linguistically significant features. Similarly to spoken languages, Signs Languages are composed of the following indivisible features\cite{adaloglou2020comprehensive}:
\begin{itemize}
\item Manual features, i.e. hand shape, position, movement, orientation of the palm or fingers
\item Non-manual features, namely eye gaze, head-nods/ shakes, shoulder orientations, various kinds of facial expression as mouthing and mouth gestures.
\end{itemize}
SiGML provides a formal notation that is easily understood by computer because it is XML based as shown in the figure \ref{fig-sigml}.

\subsubsection{Continuous Signs Language Recognition}
Continuous Signs Language Recognition (CSLR) uses contextual information that is not limited to spatial information that resides around the hand shape for fixed point in time, but also temporal information that consists of hand and body movements. Signs are famously multi-channel, information is carried in the hand shape, motions, body pose and even facial gestures \cite{slimane2021context}, \cite{camgoz2017subunets}. While models that recognize text based language use punctuation marks to separate sentences, CSLR models use pauses e.g. silent regions. There have been studies in the literature addressing automatic sign segmentation. However to the best of the authors’ knowledge, there is no study which utilizes sign segmentation for realizing continuous sign language recognition for complex scenarios. Following successful segmentation, the system needs to under-stand what information is being conveyed within a sign sentence. Current approaches tackle this by recognizing sign glosses and other linguistic components. Such methods can be grouped under the banner of CSLR. From a computer vision perspective, this is the most challenging task. Considering the input of the system is high dimensional spatio-temporal data, i.e. sign videos, models are required that understand what a signer looks like and how they interact and move within their 3D signing space. Moreover, the model needs to comprehend what these aspects mean in combination. This complex modeling problem is exacerbated by the asynchronous multi-articulatory nature of sign languages. Although there have been promising results towards CSLR, the state-of-the-art can only recognize sign glosses and operate within a limited domain of discourse, namely weather forecasts \cite{camgoz2020sign}

\subsubsection{Sign Languages Recognition Approaches}
\textbf{Specialized sub-networks (SubUNets)} allow decomposing a complex signs recognition problem into a series of specialized expert systems allowing to learn both spatial and temporal features where each subunit learns an intermediate representation that is transferred to the next subunit. (Camgoz et al.) \cite{camgoz2017subunets} proposed an architecture where each SubUNet consists of three tiers, firstly Convolutional Neural Network layer to learn spatial features, then a Bidirectional Long Short Term Memory layer to learn temporal features on top of the spatial features and finally a Connectionist Temporal Classification Loos layer to allow the networks to be trained with different length videos and label sequences as presented in figure \ref{temporal-modeling}. (Huang et al.) \cite{huang2018videobased} proposed a two streams \textbf{3DCNN} architecture for learning global (entire video frame) and local (hands area crop) video representations in combination with a Hierarchical Attention Network (HAN) for latent space based recognition eliminating the need for temporal layers as proposed in other studies \cite{slimane2021context}, \cite{camgoz2017subunets}, \cite{camgoz2020sign}. HAN is an extension of LSTM which incorporates the attention mechanism based on the structure of input. It uses a latent space which maps video representations learned on the 3DCNN streams with relevant video sentences using one hot encoding vectors. The encoder in the HAN reflects the hierarchical structures in its inputs and incorporates the attention mechanism.

\subsection{Graphical Gestural User Interfaces}
A subset of Natural Language Processing called natural language generation or simply generation produces sentences from semantic information. The lack of a written form for most sign languages means that these traditional processing stages work somewhat differently. Instead of a written string, many sign language systems will create some type of script (generally in a proprietary format for each system) that specifies the movements for an animated character. Instead of speech output, sign language systems produce an animation of a humanlike character performing the sentence (based on the information encoded in the script) \cite{huenerfauth2009sign}. HamNoSys defines a notation which is independent of the Signs Languages. A HamNoSys transcription is independent of the person or avatar performing it which solves the problem of different signers styles while signing. It records only those aspects of the gesture which are significant for the correct performance of a sign. HamNoSys defines a Signing space, see figure \ref{fig-signing-space} and a set of pre-defined locations see figure \ref{fig-signing-locations} on the surface of the body figures \cite{kennaway2015avatar}

To generate motion data for a particular avatar, numerical information about the avatar must be supplied, to be combined with the avatar- independent description of the movement, to produce avatar-specific motion data. The information required divides into two classes, information about the geometry of a specific avatar, and information about body language or signing style, which can vary independently of the avatar.
The geometrical information required about the avatar consists of:
\begin{itemize}
\item The dimensions of the avatar’s skeleton: the lengths of its bones, and the positions and orientations of the bones in some standard posture.
\item For each of the non-manuals defined in HamNoSys, the mixture of avatar-specific facial morphs or bone rotations which implements it.
\item The coordinates of every location nameable in HamNoSys, relative to the bone which moves the relevant body part. In principle, a HamNoSys location should be considered as a volume having a certain extent, the choice of which point within that volume to use being dependent on the context, but for implementation we have preferred to model each location by a single fixed point.
\end{itemize}

The first of these can in principle be read automatically from whatever avatar description file is exported by the modeling software used to create it. The second requires the avatar creator to express each of the HamNoSys facial movements as a combination of the avatar’s morphs. The third requires the avatar creator to specify the coordinates of all the HamNoSys locations on the surface of the body. Some of these locations can be discovered automatically. For example, it is fairly easy to determine surface points on the palmar, dorsal, ulnar, and radial sides of each finger, level with the midpoint of the middle phalanx. Other points are best placed manually: it is easier for the modeler to say where the ears are than to determine them algorithmically from analysis of the surface mesh. Specifying the locations of these surface points should be considered a necessary part of the construction of any avatar intended for synthetic animation. Note that no other knowledge is required of the surface mesh itself, which may be arbitrarily dense, subject to the requirement of rendering it at the desired frame rate.

The “body language” class of information consists of the mapping of HamNoSys categories to numbers. It requires:
\begin{itemize}
\item A numerical definition of the size of a “large”, “medium”, or “small” movement, “near” and “far” proximities, the shape of a “deep” or “shallow” curved arc, and similarly for all the other geometric categories in HamNoSys. Distances are best given as proportions of various measurements of the avatar: for example, “far” from the chest might be defined as a certain proportion of the length of either arm.
\item A numerical specification of the time required for each type of movement, and of how that time should change when HamNoSys specifies the manner of movement as “fast”, “slow”, “tense”, etc.
\item A numerical specification of how the hands should accelerate and decelerate during movements of different types, performed with different manners.
\item The temporal trajectories that the avatar-specific morphs and bone rotations that implement non-manual movements should follow. This takes the form of attack, sustain, and release times, plus a description of how the elements of the movement ramp up from zero to the full value during the attack, and down again to zero during the release.
\end{itemize}

Body language information is specific, not to a particular avatar body, but to the personality inhabiting it. Given all of this information, generating motion data requires solving the following problems \cite{kennaway2015avatar}.

\section{Research Schedule}\label{research-schedule}
\input{Tables/researchschedule}

\section{Research Budget}\label{research-budget}
The student counts with a USD 4000 education bonus per year provided by his employer. In the event the education bonus will not suffice the student will use his own monetary resources to cover additional expenses.
\input{Tables/researchbudget}

\section{Contributions}
This study will contribute with an annotated PSL dataset generalizable to other studies focused on accessibility for minorities and deaf population. Along with an annotated PSL dataset we will produce a set of state of art DNN architectures and models for learning features from a non annotated PSL dataset using self supervised techniques. The produced models will be compared using an ablative analysis focusing on indicators for Isolated Signs Language Recognition and Continuous Sign Language Recognition.  Additionally to Isolated SLR and Continuous SLR recognition models we will produce another set of state of art architectures and models for Sign Language Generation. These models will be trained to produce (regression) SiGML or HamNoSys notation from a Isolated SLR or a Continuous  SLR input. It is important to highlight that models designed for Sign Language Recognition and Sign Language Generation can be trained and used independently as building blocks for other studies in Signs Languages related tasks but at the same time can complement each other producing a pipeline that enables generating Signs Language elements from a non annotated dataset.

Finally we will animate the generated SiGML or HamNoSys Sign Language elements into a 3D animated avatar following state of art Human-Computer interaction techniques. Sign Language animation will allow closing the bridge between deaf people and computer based systems allowing a bi-directional communication and interaction.

\section{Target Indexed Journals}
\input{Tables/targetindexedjournals}

### 2021-05-09 phd/thesis plan/MainText/chapter3.tex
made updates for better LaTeX files organization and some minor updates on the opticalflow and rgb datasets preparers

\chapter{Figures, tables and images} \label{chap-2}

\section{Figures}

\begin{figure}[hbt!]
\center
\includegraphics[width=\linewidth]{chap3/signs-language-recognition.png} 
\caption{Signs Language Recognition}
\label{fig-slr}
\end{figure}

\begin{figure}[hbt!]
\center
\includegraphics[width=\linewidth]{chap3/sigml-example.png} 
\caption{SiGML transcription of 'here'}
\label{fig-sigml}
\end{figure}

\begin{figure}[hbt!]
\center
\includegraphics[width=\linewidth]{chap3/temporal-modeling.png} 
\caption{Temporal modeling using SubUNets}
\label{temporal-modeling}
\end{figure}

\begin{figure}[hbt!]
\center
\includegraphics[width=\linewidth]{chap3/signing-space.png} 
\caption{Signing Space}
\label{fig-signing-space}
\end{figure}

\begin{figure}[hbt!]
\center
\includegraphics[width=\linewidth]{chap3/signing-locations.png} 
\caption{Signing Locations}
\label{fig-signing-locations}
\end{figure}

\begingroup
\footnotesize
\begin{landscape}
\begin{longtable}{|>{\hspace{0pt}}m{0.142\linewidth}|>{\hspace{0pt}}m{0.083\linewidth}|>{\hspace{0pt}}m{0.065\linewidth}|>{\hspace{0pt}}m{0.063\linewidth}|>{\hspace{0pt}}m{0.077\linewidth}|>{\hspace{0pt}}m{0.063\linewidth}|>{\hspace{0pt}}m{0.109\linewidth}|>{\hspace{0pt}}m{0.073\linewidth}|>{\hspace{0pt}}m{0.106\linewidth}|>{\hspace{0pt}}m{0.079\linewidth}|>{\hspace{0pt}}m{0.063\linewidth}|}
\caption{Public Available Large Scale Datasets}\\ 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} DATASET & LANG & SIG\par{}NERS & CLA\par{}SSES & VIDEO \par{}INST. & DUR.~\par{}(H) & RES. & FPS & TYPE & MODA\par{}LITIES & YEAR \endfirsthead 
\hline
Signum SI & German & 25 & 780 & 19500 & 55.3 & 776x578 & 30 & continuous & RGB & 2007 \\ 
\hline
Signum isol. & German & 25 & 455 & 11375 & 8.43 & 776x578 & 30 & both & RGB & 2007 \\ 
\hline
Signum subset & German & 1 & 780 & 2340 & 4.92 & 776x578 & 30 & both & RGB & 2007 \\ 
\hline
Phoenix SD & German & 9 & 1231 & 6841 & 10.71 & 210x260 & 25 & continuous & RGB & 2014 \\ 
\hline
Phoenix SI & German & 9 & 1117 & 4667 & 7.28 & 210x260 & 25 & continuous & RGB & 2014 \\ 
\hline
CSL SD & Chinese & 50 & 178 & 25000 & 100+ & 1920x1080 & 30 & continuous & RGB+D & 2016 \\ 
\hline
CSL SI & Chinese & 50 & 178 & 25000 & 100+ & 1920x1080 & 30 & continuous & RGB+D & 2016 \\ 
\hline
CSL isol. & Chinese & 50 & 500 & 125000 & 67.75 & 1920x1080 & 30 & isolated & RGB+D & 2016 \\ 
\hline
Phoenix-T & German & 9 & 1231 & 8257 & 10.53 & 210x260 & 25 & continuous & RGB & 2018 \\ 
\hline
ASL 100 & English & 189 & 100 & 5736 & 5.55 & varying & varying & isolated & RGB & 2019 \\ 
\hline
ASL 1000 & English & 222 & 1000 & 25513 & 24.65 & varying & varying & isolated & RGB & 2019 \\ 
\hline
GSL isol. & Greek & 7 & 310 & 40785 & 6.44 & 848x480 & 30 & isolated & RGB+D & 2019 \\ 
\hline
GSL SD & Greek & 7 & 310 & 10290 & 9.59 & 848x480 & 30 & continuous & RGB+D & 2019 \\
\hline
\end{longtable}
\label{tab:public-large-datasets}
\end{landscape}

\endgroup

### 2021-05-09 phd/thesis plan/Tables/researchbudget.tex
made updates for better LaTeX files organization and some minor updates on the opticalflow and rgb datasets preparers

\begingroup
\footnotesize
\begin{longtable}{|>{\hspace{0pt}}m{0.286\linewidth}|>{\hspace{0pt}}m{0.357\linewidth}|>{\hspace{0pt}}m{0.104\linewidth}|>{\hspace{0pt}}m{0.19\linewidth}|} 
\hline
\rowcolor[rgb]{0.502,0.502,0.502} RESOURCES & DESCRIPTION & COST & FUNDING \endfirsthead 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} SOFTWARE LICENSING &  & USD 0.00 &  \\ 
\hline
Pycharm (Student license) & IDE for python development &  &  \\ 
\hline
AppCode (Student license) & IDE for Apps development &  &  \\ 
\hline
CLion (Student license) & IDE for C and C++ development &  &  \\ 
\hline
iOS & Mobile operating system &  &  \\ 
\hline
MacOSx & Desktop operating system &  &  \\ 
\hline
TextMaker & LateX editor &  &  \\ 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} CLOUD SERVICES &  & USD 5508.00 &  \\ 
\hline
AWS Sage Maker & \textcolor[rgb]{0.086,0.098,0.122}{5 jobs per month x 1 instances }\par{}\textcolor[rgb]{0.086,0.098,0.122}{per job x 8 hours per job = 40.00 }\par{}\textcolor[rgb]{0.086,0.098,0.122}{SageMaker Training hours per month}\par{}\textcolor[rgb]{0.086,0.098,0.122}{}\textcolor[rgb]{0.086,0.098,0.122}{}\par{}\textcolor[rgb]{0.086,0.098,0.122}{40.00 hours per month x 3.825 USD }\par{}\textcolor[rgb]{0.086,0.098,0.122}{per hour instance cost = 153.00 USD }\par{}\textcolor[rgb]{0.086,0.098,0.122}{(monthly On-Demand cost)}\par{}\textcolor[rgb]{0.086,0.098,0.122}{}\textbf{Total cost for SageMaker Training (monthly): }\par{}\textbf{153.00 USD}\textcolor[rgb]{0.086,0.098,0.122}{}\textcolor[rgb]{0.086,0.098,0.122}{} & USD 5508.00 & \par{}Student Education Bonus\par{}provided by employer.\par{} \\ 
\hline
GitHub & Source code versioning and control & USD 0.00 &  \\ 
\hline
S3 (Student version) & Cloud objects storage & USD 0.00 &  \\ 
\hline
Lambda / Serverless (Student version) & Functions as a service & USD 0.00 &  \\ 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} PERSONEL &  & USD 0.00 &  \\ 
\hline
Doctoral student &  & USD 0.00 &  \\ 
\hline
Doctoral advisor &  & USD 0.00 &  \\ 
\hline
\end{longtable}
\endgroup

### 2021-05-09 phd/thesis plan/Tables/researchschedule.tex
made updates for better LaTeX files organization and some minor updates on the opticalflow and rgb datasets preparers

\begingroup
\footnotesize
\begin{longtable}{|>{\hspace{0pt}}m{0.042\linewidth}|>{\hspace{0pt}}m{0.311\linewidth}|>{\hspace{0pt}}m{0.388\linewidth}|>{\centering\hspace{0pt}}m{0.096\linewidth}|>{\centering\arraybackslash\hspace{0pt}}m{0.096\linewidth}|} 
\hline
\rowcolor[rgb]{0.502,0.502,0.502}  & \multicolumn{1}{>{\centering\hspace{0pt}}m{0.311\linewidth}|}{TASK\textbf{}} & \multicolumn{1}{>{\centering\hspace{0pt}}m{0.388\linewidth}|}{DESCRIPTION} & START \par{}DATE & END \par{}DATE \endfirsthead 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} 1 & SELF SUPERVISED CORPUS \par{}ANNOTATION & DESIGN AND IMPLEMENT A SELF \par{}SUPERVISED PSL CORPUS \par{}ANNOTATION~METHOD & 03/15/2021 & 07/14/2022 \\ 
\hline
1.1 & Initial corpus clean up. & Perform initial corpus cleaning tasks \par{}including\par{}1. Video organization (titles, dates, \par{}duration, signer).\par{}2. Video storage in a cloud system. & 03/15/2021 & 03/26/2021 \\ 
\hline
1.2 & Peruvian sings language grammatical \par{}rules description. & Perform a detailed description of the \par{}Peruvian signs language elements \par{}including:\par\null\par{}1. Phonology.\par{}2. Dactylology.\par{}3. Numbers.\par{}4. Common Expressions.\par{}5. Vocabulary.\par{}6. Gender.\par{}7. Verb tenses.\par{}8. Question forms.\par{}9. Nouns.\par{}10. Adjectives. & 03/15/2021 & 23/07/2021 \\ 
\hline
1.3 & Define peruvian signs language \par{}manual and non manual features. & Create a descriptive list of manual \par{}and non manual features available in \par{}PSL e.g. hand shapes, rotation, \par{}orientation and other body parts like \par{}neck orientation, shoulders and arms \par{}movements. & 03/15/2021 & 03/15/2021 \\ 
\hline
1.4 & Establish relations between \par{}grammatical rules, manual and \par{}non manual features. & Using the HamNoSys / SiGML notation \par{}define the manual and non manual \par{}features for the most common \par{}PSL elements. & 03/15/2021 & 12/17/2021 \\ 
\hline
1.5 & Design the input preprocessing \par{}method. & Design the architecture design for the \par{}data input pipeline including data \par{}transformation and augmentation. \par{}The design should contain elements to \par{}detect the signer (object detection) and \par{}perform a more granular body parts \par{}detection using state of art mechanisms. & 04/19/2021 & 05/14/2021 \\ 
\hline
1.6 & Implement the input preprocessing \par{}method. & Implement the architecture designed \par{}in 1.5 and produce a software based \par{}solution for pre-processing raw PSL videos \par{}and generate a list of PSL manual and \par{}non manual elements. & 05/17/2021 & 09/10/2021 \\ 
\hline
1.7 & Design the feature extraction \par{}method. & Design a method that is capable to find \par{}relations between manual and non manual \par{}features using Word2Vec as a \par{}starting point but extended to a Visual words. & 06/07/2021 & 07/05/2021 \\ 
\hline
1.8 & Implement the feature extraction \par{}method. & Implement a software solution based on \par{}1.7 that is capable to find relations between \par{}manual and non manual features on the \par{}same glose (isolated signs detection). & 07/06/2021 & 10/11/2021 \\ 
\hline
1.9 & Design the temporal sequencing \par{}method. & Once an isolated signs detection method \par{}is implemented, extend it to a method to \par{}determine temporal sequencing between \par{}isolated signs in order to detect more complex \par{}PSL elements (e.g expressions). & 10/12/2021 & 11/12/2021 \\ 
\hline
1.10 & Implement the temporal sequencing \par{}method. & Implement a software solution that is \par{}capable to perform a continuous \par{}signs detection using temporal \par{}sequencing based on the design in 1.9. & 11/15/2021 & 12/17/2021 \\ 
\hline
1.11 & Isolated Signs Language \par{}Recognition. & Based on 1.8 create an isolated signs \par{}language dataset. & 20/12/2021 & 01/14/2022 \\ 
\hline
1.12 & Continuous Signs Language \par{}Recognition. & Based on 1.10 create a continuous signs \par{}language dataset. & 01/17/2022 & 01/31/2022 \\ 
\hline
1.13 & Design the self and continuously \par{}improving signs detection method. & Design a solution based on Energy functions \par{}and Representational Autoencoders to allow \par{}predicting and reconstructing PSL elements \par{}in order to enter into a continuous improvement \par{}executing tasks 1.5 to 1.12. & 01/02/2022 & 03/31/2022 \\ 
\hline
1.14 & Implement the self and continuously \par{}improving signs detection method. & Based on the design 1.13 implement a \par{}network architecture based on representational \par{}auto encoders with the objective to \par{}optimize an energy function. & 04/01/2022 & 07/14/2022 \\ 
\hline
1.15 & Estimate the average duration \par{}of a gloss. & Based on the design 1.13 implement a \par{}network architecture that is based on \par{}representational auto encoders and \par{}energy functions optimization to \par{}progressively detect the average \par{}duration of a PSL gloss. & 01/07/2022 & 07/14/2022 \\ 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} 2 & GRAPHICAL GESTURAL USER \par{}INTERFACE & DESIGN AND IMPLEMENT A GESTURAL\par{}USER INTERFACE FOR PSL SIGNERS & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{07/18/2022} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{07/14/2023} \\ 
\hline
2.1 & Design the conversational user \par{}interface. & Design a user interface to enable \par{}conversations between a signer and \par{}a computer. For this task we propose \par{}designing a chatbot using today’s available \par{}cloud options (e.g. Alexa) by extending it \par{}with the models implemented in 1. & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{07/18/2022} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{08/26/2022} \\ 
\hline
2.2 & Implement the conversational user \par{}interface. & Implement the user interface designed in 2.1, \par{}it will use 3D modeling using a Human \par{}simulation approach instead of concatenating \par{}a series of videos. Human simulation modeling \par{}is achievable using current technologies like \par{}Humanoid Animation (HAmin). & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{08/29/2022} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{11/30/2022} \\ 
\hline
2.3 & Signs Language synthesis to Spanish. & Using the algorithms approximated in 1\textbf{ }as well \par{}as the grammatical rules defined and created \par{}in 1.1, 1.2, 1.3 and 1.4 implement a software \par{}solution for synthesizing PSL elements to \par{}Spanish. & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{12/01/2022} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{02/01/2023} \\ 
\hline
2.4 & Spanish synthesis to Signs Language. & Using the algorithms approximated in 1\textbf{ }as well \par{}as the grammatical rules defined and created \par{}in 1.1, 1.2, 1.3 and 1.4 implement a software \par{}solution for synthesizing Spanish to \par{}PSL elements. & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{02/01/2023} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{04/01/2023} \\ 
\hline
2.5 & Signs Language Animation. & Implement the signs language animation or \par{}Human simulation modeling which is \par{}achievable using current technologies \par{}like Humanoid Animation (HAmin). & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{04/04/2023} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{07/14/2023} \\ 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} 3 & PROTOTYPE APPLICATION & DESIGN AND IMPLEMENT AN INSTALLABLE \par{}APPLICATION TO ENABLE INTERACTION\par{}BETWEEN SIGNERS AND COMPUTERS & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{07/17/2023} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{12/15/2023} \\ 
\hline
3.1 & Design an application that can be \par{}installed in devices \par{}(e.g. mobile phones and tablets) \par{}for signers human computer \par{}interaction. & Design an iOS application that can be installed \par{}in apple devices, the application to be capable \par{}to use the set of algorithms implemented in 1 \par{}and the user gestural user interface \par{}implemented in 2. & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{07/17/2023} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{08/21/2023} \\ 
\hline
3.2 & Implement an application that can \par{}be installed in devices and enable \par{}signers human computer interaction. & Implement an iOS application that use the \par{}design defined in 3.1. & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{08/22/2023} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{12/15/2023} \\ 
\hline
\rowcolor[rgb]{0.753,0.753,0.753} 4 & GENERAL TASKS & COMMON GENERAL AND ADMINISTRATIVE\par{}TASKS & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{03/15/2021} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{12/15/2023} \\ 
\hline
4.1 & Weekly meetings with thesis advisor. &  & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{03/15/2021} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{12/12/2023} \\ 
\hline
4.2 & Weekly visits to the PUCP campus. &  & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{03/15/2021} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{12/12/2023} \\ 
\hline
4.3 & Doctoral thesis plan. & Design and writing of the doctoral thesis plan. & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{03/15/2021} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{07/16/2023} \\ 
\hline
4.4 & Write and publish First paper. & Write and publishing of the first paper required \par{}for the doctoral program. & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{08/02/2021} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{07/15/2022} \\ 
\hline
4.5 & Write and publish Second paper. & Write and publishing of the second paper \par{}required for the doctoral program. & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{08/15/2022} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{07/17/2023} \\ 
\hline
4.6 & Doctoral thesis. & Design and writing of the doctoral thesis. & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{03/15/2021} & \multicolumn{1}{>{\hspace{0pt}}m{0.096\linewidth}|}{12/15/2023} \\
\hline
\end{longtable}
\endgroup

### 2021-05-09 phd/thesis plan/Tables/targetindexedjournals.tex
made updates for better LaTeX files organization and some minor updates on the opticalflow and rgb datasets preparers

\begingroup
\footnotesize
\begin{longtable}{|l|l|l|}
\caption{List of target indexed journals}\\ 
\hline
\rowcolor[rgb]{0.502,0.502,0.502} NAME & OVERVIEW & URL \endfirsthead 
\hline
\begin{tabular}[c]{@{}l@{}}International Journal \\of Machine Learning \\and Cybernetics\end{tabular} & \begin{tabular}[c]{@{}l@{}}The International Journal of Machine Learning\\~and Cybernetics (IJMLC) focuses on the key \\research problems emerging at the junction of \\machine learning and cybernetics and serves\\~as a broad forum for rapid dissemination of the \\latest advancements in the area. The emphasis \\of IJMLC is on the hybrid development of \\machine learning and cybernetics schemes \\inspired by different contributing disciplines \\such as engineering, mathematics, cognitive \\sciences, and applications. New ideas, design \\alternatives, implementations and case studies \\pertaining to all the aspects of machine learning\\~and cybernetics fall within the scope of the IJMLC.\end{tabular} & \begin{tabular}[c]{@{}l@{}}https://www.springer\\.com/journal/\\13042\end{tabular} \\ 
\hline
\begin{tabular}[c]{@{}l@{}}Machine Learning \\with Applications\end{tabular} & \begin{tabular}[c]{@{}l@{}}Machine Learning with Applications (MLWA) is a \\peer reviewed, open access journal focused on \\research related to machine learning. The journal \\encompasses all aspects of research and \\development in ML, including but not limited \\to data mining, computer vision, natural language \\processing (NLP), intelligent systems, neural networks, \\AI-based software engineering, bioinformatics and their \\applications in the areas of engineering, medicine, \\biology, education, business and social sciences. \\It covers a broad spectrum of applications in the \\community, from industry, government, and academia.\end{tabular} & \begin{tabular}[c]{@{}l@{}}https://www.journals.\\elsevier.com/\\machine-learning\\-with-applications\end{tabular} \\ 
\hline
\begin{tabular}[c]{@{}l@{}}\textcolor[rgb]{0.133,0.133,0.133}{Computers \& Electrical }\\\textcolor[rgb]{0.133,0.133,0.133}{Engineering}\end{tabular} & \begin{tabular}[c]{@{}l@{}}Published since 1973, Computers \& Electrical \\Engineering provides rapid publication of topical \\research into the integration of computer technology \\and computational techniques with electrical and \\electronic systems. The journal publishes papers \\featuring novel implementations of computers and \\computational techniques in areas like signal \\processing, high-performance computing, \\artificial intelligence, and communications.\end{tabular} & \begin{tabular}[c]{@{}l@{}}https://www.journals.\\elsevier.com/\\computers-and-\\electrical-engineering\end{tabular} \\
\hline
\end{longtable}

\endgroup

### 2021-05-09 src/main/learning/dataset/prepare/combined_dataset_preparer.py

@@ -34,9 +34,9 @@ class CombinedDatasetPreparer(BaseDatasetPreparer):
 
         return transformed_img, transformed_frames_seq
 
-    def _prepare_sample2(self, opticalflow_feature, rgb_feature, label):
-        transformed_img = self._transform_image(opticalflow_feature)
-        transformed_frames_seq = tf.py_function(self._py_transform_frame_seq, [rgb_feature], tf.float32)
+    def _prepare_sample2(self, opticalflow_sample, rgb_sample, label):
+        transformed_img = self._transform_image(opticalflow_sample)
+        transformed_frames_seq = tf.py_function(self._py_transform_frame_seq, [rgb_sample], tf.float32)
 
         return transformed_img, transformed_frames_seq, label
 

### 2021-05-09 src/main/learning/dataset/prepare/opticalflow_dataset_preparer.py

@@ -25,8 +25,8 @@ class OpticalflowDatasetPreparer(BaseDatasetPreparer):
     def transform_feature_for_predict(self, **kwargs):
         raise NotImplementedError('This method is not supported for the Opticalflow dataset')
 
-    def _prepare_sample(self, feature, label):
-        transformed_img = self._transform_image(feature)
+    def _prepare_sample(self, opticalflow_sample, label):
+        transformed_img = self._transform_image(opticalflow_sample)
         return transformed_img, label
 
     def _get_dataset_type(self):

### 2021-05-09 src/main/learning/dataset/prepare/rgb_dataset_preparer.py

@@ -24,9 +24,9 @@ class RGBDatasetPreparer(BaseDatasetPreparer):
     def transform_feature_for_predict(self, **kwargs):
         raise NotImplementedError('This method is not supported for the RGB dataset')
 
-    def _prepare_sample(self, opticalflow_feature, label):
+    def _prepare_sample(self, rgb_sample, label):
         # RNN feature
-        transformed_frames_seq = tf.py_function(self._py_transform_frame_seq, [opticalflow_feature], tf.float32)
+        transformed_frames_seq = tf.py_function(self._py_transform_frame_seq, [rgb_sample], tf.float32)
         # RNN cells require a 3D input (batch, steps, feature)
         return transformed_frames_seq, label
 

### 2021-06-26 phd/thesis plan/MainText/chapter1.tex
added a SwAV dataset preparer

@@
 \section{Literature Review}\label{literature-review}
+\subsection{Self Supervised Learning}
+Predicting the future or reconstructing the past is a way to learn and understanding the hidden patterns and distributions in data which is not necessarily annotated or pre-processed.
+Schmidhuber proposed an early self supervised method \cite{schmidhuber1990making} where two fully self supervised recurrent neural networks predicted future reinforcement based on external interactions and state sent to the next network cell is calculated on differences on successive predictions achieving a system capable to generate supervision signals.
+Unlike supervised tasks where predictions are influenced by target and observed variables, self supervised tasks are influenced in a greater extent by hidden or invisible variables also known as latent variables
+
+\subsubsection{Pretext Tasks}
+\subsubsection{Self Supervised Representation Learning}
 \subsection{Signs Language Recognition and annotation}

### 2021-06-26 src/main/app/video/video_evaluator.py

@@ -3,11 +3,11 @@ import os
 import cv2 as cv
 import tensorflow as tf
 
-from pre_processing.video.opticalflow_samples_extractor import OpticalflowSamplesExtractor
-from pre_processing.video.rgb_samples_extractor import RGBSamplesExtractor
+from opticalflow_samples_extractor import OpticalflowSamplesExtractor
+from rgb_samples_extractor import RGBSamplesExtractor
 
 from learning.model.legacy.nsdm_v2_builder import NSDMV2ModelBuilder
-from learning.dataset.prepare.combined_dataset_preparer import CombinedDatasetPreparer
+from legacy.combined_dataset_preparer import CombinedDatasetPreparer
 
 from learning.common.labels import SIGNS_CLASSES
 
@@ -35,7 +35,7 @@ class VideoEvaluator:
 
         start_time = self.start_time
         end_time = start_time + self.extract_length
-        # extract samples for 1 second video and move 0.5 seconds through the end of the video.
+        # extract samples for 1 second samples_generation and move 0.5 seconds through the end of the samples_generation.
         while end_time < video_duration:
             rgb_sample = rgb_samples_extractor.extract_sample(VideoPath=video_path, StartTime=start_time,
                                                               EndTime=end_time)

### 2021-06-26 src/main/learning/dataset/prepare/base_dataset_preparer.py

@@ -1,17 +1,15 @@
-import os
 import logging
 import tensorflow as tf
 
-from learning.dataset.tfrecord.tf_record_dataset_reader import TFRecordDatasetReader
 from learning.common.common_config import IMAGENET_CONFIG, FRAMES_SEQ_CONFIG, DATASET_BATCH_SIZE
 
 
 class BaseDatasetPreparer:
-    def __init__(self):
+    def __init__(self, train_dataset_path, test_dataset_path):
         self.logger = logging.getLogger(__name__)
 
-        self.working_dir = os.getenv('WORK_DIR', './')
-        self.dataset_dir = 'serialized-dataset'
+        self.train_dataset_path = train_dataset_path
+        self.test_dataset_path = test_dataset_path
 
         self.imagenet_img_width = IMAGENET_CONFIG['img_width']
         self.imagenet_img_height = IMAGENET_CONFIG['img_height']
@@ -26,23 +24,10 @@ class BaseDatasetPreparer:
         self.dataset_batch_size = DATASET_BATCH_SIZE
 
     def prepare_train_dataset(self):
-        return self._prepare(self._get_train_dataset_path)
+        return self._prepare(self.train_dataset_path)
 
     def prepare_test_dataset(self):
-        return self._prepare(self._get_test_dataset_path)
-
-    def _prepare(self, get_dataset_path_func):
-        dataset_path = get_dataset_path_func()
-        dataset_reader = TFRecordDatasetReader(self._get_dataset_type(), dataset_path)
-        dataset = dataset_reader.read()
-        return dataset
-
-    def _get_train_dataset_path(self):
-        return os.path.join(self.working_dir, self.dataset_dir, self._get_dataset_type())
-
-    # TODO: for now the test dataset path is the same as the train dataset path
-    def _get_test_dataset_path(self):
-        return self._get_train_dataset_path()
+        return self._prepare(self.test_dataset_path)
 
     def _transform_image(self, img, resize_shape=None):
         if resize_shape is None:
@@ -63,6 +48,13 @@ class BaseDatasetPreparer:
         img = tf.image.convert_image_dtype(img, tf.float32)
         # resize the image to the desired size.
         return tf.image.resize(img, resize_shape)
+
+    @staticmethod
+    def _transform(input_data, trans_ops):
+        transformed_input = input_data
+        for op in trans_ops:
+            transformed_input = op(transformed_input)
+        return transformed_input
 
     def _py_transform_frame_seq(self, sample):
         transformed_images = []
[…]

### 2021-06-26 src/main/learning/dataset/prepare/legacy/combined_dataset_preparer.py

@@ -1,13 +1,13 @@
 import logging
 import tensorflow as tf
 
-from learning.dataset.prepare.base_dataset_preparer import BaseDatasetPreparer
+from learning.dataset.prepare.serialized_dataset_preparer import SerializedDatasetPreparer
 from learning.common.dataset_type import COMBINED
 
 
-class CombinedDatasetPreparer(BaseDatasetPreparer):
-    def __init__(self):
-        super().__init__()
+class CombinedDatasetPreparer(SerializedDatasetPreparer):
+    def __init__(self, train_dataset_path, test_dataset_path):
+        super().__init__(train_dataset_path, test_dataset_path)
         self.logger = logging.getLogger(__name__)
 
     def prepare_train_dataset(self):
@@ -22,6 +22,21 @@ class CombinedDatasetPreparer(BaseDatasetPreparer):
     def prepare_test_dataset(self):
         return self.prepare_train_dataset()
 
+    def _prepare_sample(self, feature, sample):
+        raise NotImplementedError('This method is not supported for the Combined dataset')
+
+    def _prepare_sample2(self, opticalflow_sample, rgb_sample, label):
+        transformed_img = self._transform_image(opticalflow_sample)
+        transformed_frames_seq = tf.py_function(self._py_transform_frame_seq, [rgb_sample], tf.float32)
+
+        return transformed_img, transformed_frames_seq, label
+
+    def _prepare_sample3(self, feature):
+        raise NotImplementedError('This method is not supported for the Combined dataset')
+
+    def _get_dataset_type(self):
+        return COMBINED
+
     def transform_feature_for_predict(self, **kwargs):
         opticalflow_feature = kwargs['OpticalflowFeature']
         rgb_feature = kwargs['RGBFeature']
@@ -33,15 +48,3 @@ class CombinedDatasetPreparer(BaseDatasetPreparer):
         transformed_frames_seq = tf.expand_dims(self._py_transform_decoded_frame_seq(rgb_feature), axis=0)
 
         return transformed_img, transformed_frames_seq
-
-    def _prepare_sample2(self, opticalflow_sample, rgb_sample, label):
-        transformed_img = self._transform_image(opticalflow_sample)
-        transformed_frames_seq = tf.py_function(self._py_transform_frame_seq, [rgb_sample], tf.float32)
-
-        return transformed_img, transformed_frames_seq, label
-
-    def _get_dataset_type(self):
-        return COMBINED
-
[…]

### 2021-06-26 src/main/learning/dataset/prepare/legacy/opticalflow_dataset_preparer.py

@@ -1,13 +1,13 @@
 import logging
 import tensorflow as tf
 
-from learning.dataset.prepare.base_dataset_preparer import BaseDatasetPreparer
+from learning.dataset.prepare.serialized_dataset_preparer import SerializedDatasetPreparer
 from learning.common.dataset_type import OPTICAL_FLOW
 
 
-class OpticalflowDatasetPreparer(BaseDatasetPreparer):
-    def __init__(self):
-        super().__init__()
+class OpticalflowDatasetPreparer(SerializedDatasetPreparer):
+    def __init__(self, train_dataset_path, test_dataset_path):
+        super().__init__(train_dataset_path, test_dataset_path)
         self.logger = logging.getLogger(__name__)
 
     def prepare_train_dataset(self):
@@ -22,15 +22,18 @@ class OpticalflowDatasetPreparer(BaseDatasetPreparer):
     def prepare_test_dataset(self):
         return self.prepare_train_dataset()
 
-    def transform_feature_for_predict(self, **kwargs):
-        raise NotImplementedError('This method is not supported for the Opticalflow dataset')
-
     def _prepare_sample(self, opticalflow_sample, label):
         transformed_img = self._transform_image(opticalflow_sample)
         return transformed_img, label
 
-    def _get_dataset_type(self):
-        return OPTICAL_FLOW
-
     def _prepare_sample2(self, feature1, feature2, label):
         raise NotImplementedError('This method is not supported for the Opticalflow dataset')
+
+    def _prepare_sample3(self, feature):
+        raise NotImplementedError('This method is not supported for the Opticalflow dataset')
+
+    def transform_feature_for_predict(self, **kwargs):
+        raise NotImplementedError('This method is not supported for the Opticalflow dataset')
+
+    def _get_dataset_type(self):
+        return OPTICAL_FLOW

### 2021-06-26 src/main/learning/dataset/prepare/legacy/rgb_dataset_preparer.py

@@ -1,13 +1,13 @@
 import logging
 import tensorflow as tf
 
-from learning.dataset.prepare.base_dataset_preparer import BaseDatasetPreparer
+from learning.dataset.prepare.serialized_dataset_preparer import SerializedDatasetPreparer
 from learning.common.dataset_type import RGB
 
 
-class RGBDatasetPreparer(BaseDatasetPreparer):
-    def __init__(self):
-        super().__init__()
+class RGBDatasetPreparer(SerializedDatasetPreparer):
+    def __init__(self, train_dataset_path, test_dataset_path):
+        super().__init__(train_dataset_path, test_dataset_path)
         self.logger = logging.getLogger(__name__)
 
     def prepare_train_dataset(self):
@@ -21,17 +21,20 @@ class RGBDatasetPreparer(BaseDatasetPreparer):
     def prepare_test_dataset(self):
         return self.prepare_train_dataset()
 
-    def transform_feature_for_predict(self, **kwargs):
-        raise NotImplementedError('This method is not supported for the RGB dataset')
-
     def _prepare_sample(self, rgb_sample, label):
         # RNN feature
         transformed_frames_seq = tf.py_function(self._py_transform_frame_seq, [rgb_sample], tf.float32)
         # RNN cells require a 3D input (batch, steps, feature)
         return transformed_frames_seq, label
 
+    def _prepare_sample2(self, feature1, feature2, label):
+        raise NotImplementedError('This method is not supported for the RGB dataset')
+
+    def _prepare_sample3(self, feature):
+        raise NotImplementedError('This method is not supported for the RGB dataset')
+
     def _get_dataset_type(self):
         return RGB
 
-    def _prepare_sample2(self, feature1, feature2, label):
+    def transform_feature_for_predict(self, **kwargs):
         raise NotImplementedError('This method is not supported for the RGB dataset')

### 2021-06-26 src/main/learning/dataset/prepare/raw_dataset_preparer.py

@@ -1,0 +1,41 @@
+import tensorflow as tf
+
+from learning.dataset.prepare.base_dataset_preparer import BaseDatasetPreparer
+from pathlib import Path
+
+
+class RawDatasetPreparer(BaseDatasetPreparer):
+    def __init__(self, train_dataset_path, test_dataset_path):
+        super().__init__(train_dataset_path, test_dataset_path)
+
+    def _prepare(self, get_dataset_path_func):
+        dataset_path = get_dataset_path_func()
+        raw_file_list = self._get_raw_file_list(dataset_path)
+        dataset: tf.data.Dataset = tf.data.Dataset.from_generator(self._data_generator, args=[raw_file_list])
+        dataset = dataset.map(self._prepare_sample3, num_parallel_calls=tf.data.AUTOTUNE, deterministic=False)
+
+        return dataset
+
+    def _get_raw_file_list(self, dataset_path):
+        file_types = self._get_raw_file_types()
+        raw_file_list = []
+
+        for file_type in file_types:
+            raw_file_list.extend([str(file_path) for file_path in Path(dataset_path).rglob(file_type)])
+
+        return raw_file_list
+
+    def _prepare_sample(self, feature, label):
+        raise NotImplementedError('_prepare_sample method not implemented.')
+
+    def _prepare_sample2(self, feature1, feature2, label):
+        raise NotImplementedError('_prepare_sample2 method not implemented.')
+
+    def _prepare_sample3(self, feature):
+        raise NotImplementedError('_prepare_sample3 method not implemented.')
+
+    def _get_raw_file_types(self):
+        raise NotImplementedError('_get_raw_file_types method not implemented.')
+
+    def _data_generator(self, list_video_path):
+        raise NotImplementedError('_data_generator method not implemented.')

### 2021-06-26 src/main/learning/dataset/prepare/serialized_dataset_preparer.py

@@ -1,0 +1,30 @@
+from learning.dataset.prepare.base_dataset_preparer import BaseDatasetPreparer
+from learning.dataset.tfrecord.tf_record_dataset_reader import TFRecordDatasetReader
+
+
+class SerializedDatasetPreparer(BaseDatasetPreparer):
+    def __init__(self, train_dataset_path, test_dataset_path):
+        super().__init__(train_dataset_path, test_dataset_path)
+
+    def _prepare(self, get_dataset_path_func):
+        dataset_path = get_dataset_path_func()
+        dataset_reader = TFRecordDatasetReader(self._get_dataset_type(), dataset_path)
+        dataset = dataset_reader.read()
+        return dataset
+
+    def _prepare_sample(self, feature, label):
+        pass
+
+    def _prepare_sample2(self, feature1, feature2, label):
+        pass
+
+    def _prepare_sample3(self, feature):
+        raise NotImplementedError('_prepare_sample3 method not implemented.')
+
+    def _get_dataset_type(self):
+        raise NotImplementedError('_get_dataset_type method not implemented.')
+
+    def transform_feature_for_predict(self):
+        raise NotImplementedError('transform_feature_for_prediction method not implemented.')
+
+

### 2021-06-26 src/main/learning/dataset/prepare/swav/multi_crop.py

@@ -1,0 +1,146 @@
+import tensorflow as tf
+
+
+class MultiCropTransformer:
+    def tie_together(self, video_fragment, min_scale, max_scale, crop_size):
+        video_fragment = self.decode(video_fragment)
+        video_fragment = self.scale(video_fragment)
+
+        # Random resized crops
+        video_fragment = self.random_resize_crop(video_fragment, min_scale, max_scale, crop_size)
+
+        # Color distortions & Gaussian blur
+        video_fragment = self.custom_augment(video_fragment)
+
+        return video_fragment
+
+    @staticmethod
+    def random_resize_crop(video_fragment, min_scale, max_scale, crop_size):
+        crop_resized_images = []
+
+        for image in video_fragment:
+            # Conditional resizing
+            if crop_size == 224:
+                image_shape = 260
+                image = tf.image.resize(image, (image_shape, image_shape))
+            else:
+                image_shape = 160
+                image = tf.image.resize(image, (image_shape, image_shape))
+
+            # Get the crop size for given min and max scale
+            size = tf.random.uniform(shape=(1,), minval=min_scale * image_shape, maxval=max_scale * image_shape,
+                                     dtype=tf.float32)
+            size = tf.cast(size, tf.int32)[0]
+            # Get the crop from the image
+            crop = tf.image.random_crop(image, (size, size, 3))
+            crop_resize = tf.image.resize(crop, (crop_size, crop_size))
+            crop_resized_images.append(crop_resize)
+
+        return crop_resized_images
+
+    def custom_augment(self, video_fragment):
+        video_fragment = self.random_apply(tf.image.flip_left_right, video_fragment, p=0.5)
+
+        # Randomly apply Gaussian blur
+        video_fragment = self.random_apply(self.gaussian_blur, video_fragment, p=0.5)
+
+        # Randomly apply transformation (color distortions) with probability p.
+        video_fragment = self.random_apply(self.color_jitter, video_fragment, p=0.8)
+
+        # Randomly apply grayscale
+        video_fragment = self.random_apply(self.color_drop, video_fragment, p=0.2)
+
+        return video_fragment
+
+    @staticmethod
+    def gaussian_blur(video_fragment, kernel_size=23, padding='SAME'):
+        transformed_video_fragment = []
+
+        for image in video_fragment:
[…]

### 2021-06-26 src/main/learning/dataset/prepare/swav/swav_video_dataset_preparer.py

@@ -1,0 +1,75 @@
+import cv2
+import tensorflow as tf
+
+from learning.dataset.prepare.base_dataset_preparer import BaseDatasetPreparer
+from pre_processing.isolated_detection.samples_generation.rgb_samples_extractor import RGBSamplesExtractor
+from learning.dataset.prepare.swav.multi_crop import MultiCropTransformer
+
+
+class SwAVDatasetPreparer(BaseDatasetPreparer):
+    def __init__(self, train_dataset_path, test_dataset_path):
+        super().__init__(train_dataset_path, test_dataset_path)
+
+        self.crop_sizes = [224, 96]
+        self.num_crops = [2, 3]
+        self.min_scale = [0.5, 0.14]
+        self.max_scale = [1., 0.5]
+
+        self.multi_crop = MultiCropTransformer()
+
+    def _prepare(self, get_dataset_path_func):
+        return super()._prepare(get_dataset_path_func)
+
+    def _data_generator(self, list_video_path):
+        extractor = RGBSamplesExtractor()
+
+        for video_path in list_video_path:
+            duration = self.__get_video_duration(video_path)
+            start_time = 0.0
+            end_time = self.__get_next_end_time(start_time=start_time)
+
+            while end_time < duration:
+                fragment_frames = extractor.extract_sample(VideoPath=video_path, StartTime=start_time, EndTime=end_time)
+
+                end_time = self.__get_next_end_time(start_time=end_time)
+                start_time += end_time
+
+                yield fragment_frames
+
+    @staticmethod
+    def __get_next_end_time(start_time):
+        fragment_duration = tf.random.uniform(shape=[], minval=0.3, maxval=0.5, dtype=tf.dtypes.float32)
+        end_time = start_time + fragment_duration
+        return end_time
+
+    def _prepare_sample3(self, video_fragment):
+        self.crop_sizes = [224, 96]
+        self.num_crops = [2, 3]
+        self.min_scale = [0.5, 0.14]
+        self.max_scale = [1., 0.5]
+
+        crops = tuple()
+        for idx, num_crop in enumerate(self.num_crops):
+            for _ in range(num_crop):
+                transformed_video_fragment = self.multi_crop.tie_together(video_fragment,
+                                                                          self.min_scale[idx],
+                                                                          self.max_scale[idx],
+                                                                          self.crop_sizes[idx])
+                crops += (transformed_video_fragment,)
+        return crops
+
[…]

### 2021-06-26 src/main/learning/execution/legacy/nsdm_executor.py

@@ -4,7 +4,7 @@ import tensorflow as tf
 from tensorflow.keras.models import Model
 from base_model_executor import BaseModelExecutor
 from learning.common import model_type
-from learning.dataset.prepare.combined_dataset_preparer import CombinedDatasetPreparer
+from legacy.combined_dataset_preparer import CombinedDatasetPreparer
 
 
 class NSDMExecutor(BaseModelExecutor):

### 2021-06-26 src/main/learning/execution/legacy/opticalflow_executor.py

@@ -3,7 +3,7 @@ import logging
 from tensorflow.keras.models import Model
 from base_model_executor import BaseModelExecutor
 from learning.common import model_type
-from learning.dataset.prepare.opticalflow_dataset_preparer import OpticalflowDatasetPreparer
+from legacy.opticalflow_dataset_preparer import OpticalflowDatasetPreparer
 
 
 class OpticalflowExecutor(BaseModelExecutor):

### 2021-06-26 src/main/learning/execution/legacy/rgb_executor.py

@@ -3,7 +3,7 @@ import logging
 from tensorflow.keras.models import Model
 from base_model_executor import BaseModelExecutor
 from learning.common import model_type
-from learning.dataset.prepare.rgb_dataset_preparer import RGBDatasetPreparer
+from legacy.rgb_dataset_preparer import RGBDatasetPreparer
 
 
 class RGBExecutor(BaseModelExecutor):

### 2021-06-26 src/main/learning/model/legacy/nsdm_builder.py

@@ -1,20 +1,21 @@
 import logging
 
 from tensorflow.keras.models import Model
-from learning.model.legacy.base_model_builder import BaseModelBuilder
+from learning.model.legacy.signs_detection_base_model import SignsDetectionBaseModelBuilder
 from tensorflow.keras import Input
 from tensorflow.keras.layers import Average
 from learning.common.model_type import NSDM
 
 
-class NSDMModelBuilder(BaseModelBuilder):
+class NSDMModelBuilder(SignsDetectionBaseModelBuilder):
     def __init__(self):
         super().__init__()
 
         # input shape 224x224x3
         self.opticalflow_input_shape = (self.imagenet_img_width, self.imagenet_img_height, self.imagenet_rgb_no_channels)
         # feature dim calculated from 224 * 224 * 3
-        self.rgb_input_shape = (None, self.frames_seq_img_width * self.frames_seq_img_height * self.frames_seq_no_channels)
+        self.rgb_input_shape = (None,
+                                self.frames_seq_img_width * self.frames_seq_img_height * self.frames_seq_no_channels)
 
         self.logger = logging.getLogger(__name__)
 
@@ -37,3 +38,9 @@ class NSDMModelBuilder(BaseModelBuilder):
 
     def get_model_type(self):
         return NSDM
+
+    def build2(self, **kwargs) -> Model:
+        raise NotImplementedError('build2 method not implemented.')
+
+    def build3(self, **kwargs) -> Model:
+        raise NotImplementedError('build3 method not implemented.')

### 2021-06-26 src/main/learning/model/legacy/nsdm_v2_builder.py

@@ -39,3 +39,9 @@ class NSDMV2ModelBuilder(NSDMModelBuilder):
 
     def get_model_type(self):
         return NSDMV2
+
+    def build2(self, **kwargs) -> Model:
+        raise NotImplementedError('build2 method not implemented.')
+
+    def build3(self, **kwargs) -> Model:
+        raise NotImplementedError('build3 method not implemented.')

### 2021-06-26 src/main/learning/model/legacy/opticalflow_model_builder.py

@@ -1,6 +1,6 @@
 import logging
 
-from learning.model.legacy.base_model_builder import BaseModelBuilder
+from learning.model.legacy.signs_detection_base_model import SignsDetectionBaseModelBuilder
 from tensorflow.keras.applications.resnet_v2 import ResNet152V2
 from tensorflow.keras.models import Model
 from tensorflow.keras.layers import Dense, GlobalAveragePooling2D
@@ -8,7 +8,7 @@ from tensorflow.keras import Input
 from learning.common.model_utility import OPTICAL_FLOW
 
 
-class OpticalFlowModelBuilder(BaseModelBuilder):
+class OpticalFlowModelBuilder(SignsDetectionBaseModelBuilder):
     def __init__(self):
         super().__init__()
         self.logger = logging.getLogger(__name__)
@@ -36,3 +36,9 @@ class OpticalFlowModelBuilder(BaseModelBuilder):
 
     def get_model_type(self):
         return OPTICAL_FLOW
+
+    def build2(self, **kwargs) -> Model:
+        raise NotImplementedError('build2 method not implemented.')
+
+    def build3(self, **kwargs) -> Model:
+        raise NotImplementedError('build3 method not implemented.')

### 2021-06-26 src/main/learning/model/legacy/rgb_recurrent_model_builder.py

@@ -1,13 +1,13 @@
 import logging
 
-from learning.model.legacy.base_model_builder import BaseModelBuilder
+from learning.model.legacy.signs_detection_base_model import SignsDetectionBaseModelBuilder
 from tensorflow.keras.models import Model
 from tensorflow.keras import Input
 from tensorflow.keras.layers import Masking, Bidirectional, Dense, LSTM
 from learning.common.model_type import RGB
 
 
-class RGBRecurrentModelBuilder(BaseModelBuilder):
+class RGBRecurrentModelBuilder(SignsDetectionBaseModelBuilder):
     def __init__(self):
         super().__init__()
         self.logger = logging.getLogger(__name__)
@@ -30,3 +30,9 @@ class RGBRecurrentModelBuilder(BaseModelBuilder):
 
     def get_model_type(self):
         return RGB
+
+    def build2(self, **kwargs) -> Model:
+        raise NotImplementedError('build2 method not implemented.')
+
+    def build3(self, **kwargs) -> Model:
+        raise NotImplementedError('build3 method not implemented.')

### 2021-06-26 src/main/learning/model/legacy/signs_detection_base_model.py

@@ -1,0 +1,33 @@
+from tensorflow.python.keras import Model
+
+from learning.model.base_model_builder import BaseModelBuilder
+from learning.common.labels import SIGNS_CLASSES
+from learning.common.common_config import IMAGENET_CONFIG, FRAMES_SEQ_CONFIG
+
+
+class SignsDetectionBaseModelBuilder(BaseModelBuilder):
+    def __init__(self):
+        super().__init__()
+
+        self.classes = SIGNS_CLASSES
+        self.no_classes = len(self.classes)
+
+        self.imagenet_img_width = IMAGENET_CONFIG['img_width']
+        self.imagenet_img_height = IMAGENET_CONFIG['img_height']
+        self.imagenet_rgb_no_channels = IMAGENET_CONFIG['rgb_no_channels']
+
+        self.frames_seq_img_width = FRAMES_SEQ_CONFIG['img_width']
+        self.frames_seq_img_height = FRAMES_SEQ_CONFIG['img_height']
+        self.frames_seq_no_channels = FRAMES_SEQ_CONFIG['rgb_no_channels']
+
+    def build(self, **kwargs) -> Model:
+        raise NotImplementedError('build1 method not implemented.')
+
+    def build2(self, **kwargs) -> Model:
+        raise NotImplementedError('build2 method not implemented.')
+
+    def build3(self, **kwargs) -> Model:
+        raise NotImplementedError('build3 method not implemented.')
+
+    def get_model_type(self):
+        raise NotImplementedError('get_model_type method not implemented.')

### 2021-06-26 src/main/learning/model/swav/swav_builder.py

@@ -1,0 +1,19 @@
+from tensorflow.python.keras import Model
+from learning.model.base_model_builder import BaseModelBuilder
+
+
+class SwAVModelBuilder(BaseModelBuilder):
+    def __init__(self):
+        super().__init__()
+
+    def build(self, **kwargs) -> Model:
+        pass
+
+    def build2(self, **kwargs) -> Model:
+        pass
+
+    def build3(self, **kwargs) -> Model:
+        pass
+
+    def get_model_type(self):
+        pass

### 2021-06-26 src/main/pre_processing/audio/audio_processor.py

@@ -1,44 +1,0 @@
-import os
-from moviepy.editor import VideoFileClip
-from pre_processing.common.downloader import Downloader
-
-
-class AudioProcessor(Downloader):
-    __audio_clips_base = 'audio-clips'
-    __audio_clip_extension = '.wav'
-
-    def __init__(self):
-        super().__init__()
-
-    def process(self, data):
-        super().process(data)
-
-        # extract the audio from video
-        audio_file_path = self.__generate_audio_clip_file_path(self._video_file_path)
-        if not os.path.exists(audio_file_path):
-            os.makedirs(os.path.dirname(audio_file_path), exist_ok=True)
-
-            self.__extract_audio(self._video_file_path, audio_file_path)
-            print('the audio clip is available at: {}'.format(audio_file_path))
-        else:
-            print('the audio clip is already available at: {}'.format(audio_file_path))
-
-        # check if the the audio clip exist in s3
-        audio_key = self.__generate_audio_clip_key(audio_file_path)
-        self.upload_file(audio_file_path, audio_key)
-
-    @staticmethod
-    def __extract_audio(video_file_path, audio_file_path):
-        video_clip = VideoFileClip(video_file_path)
-        audio_clip = video_clip.audio
-
-        # save to temp file
-        audio_clip.write_audiofile(audio_file_path)
-
-    def __generate_audio_clip_file_path(self, video_file_path: str):
-        relative_path, _ = os.path.splitext(video_file_path.replace(self._source_videos_base, self.__audio_clips_base))
-        return os.path. os.path.join(self._binary_folder, relative_path + self.__audio_clip_extension)
-
-    def __generate_audio_clip_key(self, audio_clip_file_path: str):
-        path = audio_clip_file_path.replace(self._binary_folder, '')
-        return path[1:] if path.startswith('/') else path

### 2021-06-26 src/main/pre_processing/common/downloader.py

@@ -9,13 +9,13 @@ class Downloader(Processor):
     def process(self, data):
         super().process(data)
 
-        # download the video clip from aws s3 and save to temp file
+        # download the samples_generation clip from aws s3 and save to temp file
         self._video_file_path = self._generate_video_file_path(self._video_key)
 
         if not os.path.exists(self._video_file_path):
             os.makedirs(os.path.dirname(self._video_file_path), exist_ok=True)
 
             self._s3.download_file(self._bucketName, self._video_key, self._video_file_path)
-            print('the video clip is available at: {}'.format(self._video_file_path))
+            print('the samples_generation clip is available at: {}'.format(self._video_file_path))
         else:
-            print('the video is already available at: {}'.format(self._video_file_path))
+            print('the samples_generation is already available at: {}'.format(self._video_file_path))

### 2021-06-26 src/main/pre_processing/file_processing_handler.py

@@ -1,13 +1,14 @@
 import logging
 
-from audio.audio_processor import AudioProcessor
-from text_analysis.keywords_detection_processor import KeywordsDetectionProcessor
-from text_analysis.transcription_syntax_detection_processor import TranscriptionSyntaxDetectionProcessor
-from transcription.transcription_processor import TranscriptionProcessor
-from video.video_splitting_processor import VideoSplittingProcessor
-from text_analysis.samples_metadata_generation_processor import SamplesMetadataGenerationProcessor
-from video.rgb_samples_processor import RGBSamplesProcessor
-from video.opticalflow_samples_processor import OpticalflowSamplesProcessor
+from common.enums import PreProcessingOperation as PreProcOp
+from isolated_detection.text_analysis.keywords_detection_processor import KeywordsDetectionProcessor
+from isolated_detection.text_analysis.transcription_syntax_detection_processor \
+    import TranscriptionSyntaxDetectionProcessor
+from isolated_detection.video_transcription.transcription_processor import TranscriptionProcessor
+from isolated_detection.samples_generation.video_splitting_processor import VideoSplittingProcessor
+from isolated_detection.text_analysis.samples_metadata_generation_processor import SamplesMetadataGenerationProcessor
+from isolated_detection.samples_generation.rgb_samples_processor import RGBSamplesProcessor
+from isolated_detection.samples_generation.opticalflow_samples_processor import OpticalflowSamplesProcessor
 
 
 class FileProcessingHandler:
@@ -18,14 +19,13 @@ class FileProcessingHandler:
         operation = data['operation']
 
         operations = {
-            'extract-audio': lambda x: AudioProcessor().process(x),
-            'video-split': lambda x: VideoSplittingProcessor().process(x),
-            'transcribe-audio': lambda x: TranscriptionProcessor().process(x),
-            'transcription-syntax-detection': lambda x: TranscriptionSyntaxDetectionProcessor().process(x),
-            'keywords-detection': lambda x: KeywordsDetectionProcessor().process(x),
-            'samples-metadata-generation': lambda x: SamplesMetadataGenerationProcessor().process(x),
-            'rgb-samples-generation': lambda x: RGBSamplesProcessor().process(x),
[…]

### 2021-06-26 src/main/pre_processing/isolated_detection/samples_generation/opticalflow_samples_handler.py

@@ -3,7 +3,7 @@ import os
 import logging
 
 from pre_processing.common.samples_handler import SamplesHandler
-from pre_processing.video.opticalflow_samples_extractor import OpticalflowSamplesExtractor
+from opticalflow_samples_extractor import OpticalflowSamplesExtractor
 
 
 class OpticalflowSamplesHandler(SamplesHandler):

### 2021-06-26 src/main/pre_processing/isolated_detection/samples_generation/opticalflow_samples_processor.py

@@ -4,8 +4,8 @@ import logging
 import glob
 
 from pre_processing.common.processor import Processor
-from pre_processing.video.opticalflow_samples_handler import OpticalflowSamplesHandler
 from pre_processing.common import io_utils
+from opticalflow_samples_handler import OpticalflowSamplesHandler
 
 
 class OpticalflowSamplesProcessor(Processor):

### 2021-06-26 src/main/pre_processing/isolated_detection/samples_generation/rgb_samples_handler.py

@@ -2,7 +2,7 @@ import cv2
 import os
 
 from pre_processing.common.samples_handler import SamplesHandler
-from pre_processing.video.rgb_samples_extractor import RGBSamplesExtractor
+from rgb_samples_extractor import RGBSamplesExtractor
 
 
 class RGBSamplesHandler(SamplesHandler):

### 2021-06-26 src/main/pre_processing/isolated_detection/samples_generation/rgb_samples_processor.py

@@ -4,8 +4,8 @@ import logging
 import csv
 
 from pre_processing.common.processor import Processor
-from pre_processing.video.rgb_samples_handler import RGBSamplesHandler
 from pre_processing.common import io_utils
+from rgb_samples_handler import RGBSamplesHandler
 
 
 class RGBSamplesProcessor(Processor):

### 2021-06-26 src/main/pre_processing/isolated_detection/samples_generation/video_splitting_processor.py

@@ -7,7 +7,7 @@ import io_utils
 from subprocess import check_call
 from datetime import datetime
 
-from downloader import Downloader
+from pre_processing.common.downloader import Downloader
 
 
 class VideoSplittingProcessor(Downloader):
@@ -31,7 +31,7 @@ class VideoSplittingProcessor(Downloader):
                 self.logger.error(e)
 
         exec_end = time.time()
-        self.logger.debug('Video splitting completed processing %s video chunks in %s seconds',
+        self.logger.debug('Video splitting completed processing %s samples_generation chunks in %s seconds',
                           len(seqs), round(exec_end - exec_start, 2))
 
     def __split_video(self, start, end, idx):

### 2021-06-26 src/main/pre_processing/isolated_detection/text_analysis/keywords_detection_processor.py

@@ -4,7 +4,7 @@ import time
 import os
 
 from datetime import datetime
-from processor import Processor
+from pre_processing.common.processor import Processor
 
 
 class KeywordsDetectionProcessor(Processor):

### 2021-06-26 src/main/pre_processing/isolated_detection/text_analysis/samples_metadata_generation_processor.py

@@ -26,13 +26,13 @@ class SamplesMetadataGenerationProcessor(Processor):
         self.__syntax_local_folder_path = data['syntax']['localFolderPath']
         self.__syntax_s3_folder_path = data['syntax']['s3Path']
 
-        self.__transcription_local_folder_path = data['transcription']['localFolderPath']
-        self.__transcription_s3_folder_path = data['transcription']['s3Path']
+        self.__transcription_local_folder_path = data['video_transcription']['localFolderPath']
+        self.__transcription_s3_folder_path = data['video_transcription']['s3Path']
 
         if self.__override_local_files:
             # download syntax detection files (e.g. nouns and numbers)
             self.s3_batch_download_files(self.__syntax_s3_folder_path, allowed_extensions=['.csv'])
-            # download transcription files
+            # download video_transcription files
             self.s3_batch_download_files(self.__transcription_s3_folder_path, allowed_extensions=['.json'])
 
         # load syntax tokens (nouns and numbers) in a list
@@ -43,7 +43,7 @@ class SamplesMetadataGenerationProcessor(Processor):
         metadata_file_path = self.__create_metadata_file(metadata_file_columns)
         self.logger.debug('Samples Metadata File created at: %s', metadata_file_path)
 
-        # find syntax occurrence within transcription files
+        # find syntax occurrence within video_transcription files
         transcription_paths = io_utils.get_files_in_folder(self._work_dir,
                                                            self.__transcription_local_folder_path, "*.json")
         for transcription_file_path in transcription_paths:

### 2021-06-26 src/main/pre_processing/isolated_detection/text_analysis/transcription_syntax_detection_processor.py

@@ -4,7 +4,7 @@ import os
 import boto3
 import csv
 
-from processor import Processor
+from pre_processing.common.processor import Processor
 from pre_processing.common import io_utils
 from collections import Counter
 
@@ -34,7 +34,7 @@ class TranscriptionSyntaxDetectionProcessor(Processor):
             if not transcription_key.endswith('.json'):
                 continue
 
-            self.logger.debug('Extracting audio transcription from: %s', transcription_key)
+            self.logger.debug('Extracting audio video_transcription from: %s', transcription_key)
 
             file_obj = self._s3.get_object(Bucket=self._bucketName, Key=transcription_key)
             file_content = file_obj['Body'].read().decode('utf-8')
@@ -66,10 +66,10 @@ class TranscriptionSyntaxDetectionProcessor(Processor):
         s3_key = os.path.join(s3_path, syntax_file_name)
 
         if self.s3_exist_object(s3_key):
-            self.logger.debug('audio transcription transcription-syntax-detection already available in s3 at: %s',
+            self.logger.debug('audio video_transcription video_transcription-syntax-detection already available in s3 at: %s',
                               s3_key)
             self.s3_delete_object(s3_key)
-            self.logger.debug('audio transcription %s deleted', s3_key)
+            self.logger.debug('audio video_transcription %s deleted', s3_key)
 
         if os.path.exists(local_file_path):
             self.logger.debug('local file %s exist', local_file_path)
@@ -113,8 +113,8 @@ class TranscriptionSyntaxDetectionProcessor(Processor):
             for detected_token in detected_tokens:
                 wr.writerow({'token': detected_token[0], 'count': detected_token[1]})
 
-        self.logger.debug('audio transcription syntax is available locally at: %s',
+        self.logger.debug('audio video_transcription syntax is available locally at: %s',
                           local_file_path)
 
         self.upload_file(local_file_path, s3_key)
-        self.logger.debug('audio transcription syntax uploaded to: %s', s3_key)
+        self.logger.debug('audio video_transcription syntax uploaded to: %s', s3_key)

### 2021-06-26 src/main/pre_processing/isolated_detection/video_transcription/transcription_processor.py

@@ -23,15 +23,16 @@ class TranscriptionProcessor(Processor):
         super().process(data)
 
         prefix = io_utils.get_video_chunk_base_key(self._video_key)
-        self.logger.debug('using %s for video chunks prefix', prefix)
+        self.logger.debug('using %s for samples_generation chunks prefix', prefix)
 
         video_chunks = self._s3.list_objects_v2(Bucket=self._bucketName, Prefix=prefix)
         if not (video_chunks and 'Contents' in video_chunks):
-            raise RuntimeError('There are not video chunks available for processing at the following location: {}'
+            raise RuntimeError('There are not samples_generation chunks available for processing at the following '
+                               'location: {} '
                                .format(prefix))
 
-        self.logger.debug('%d video chunks were found', len(video_chunks['Contents']))
-        self.logger.debug('This operation will process %d transcription job', max_no_jobs)
+        self.logger.debug('%d samples_generation chunks were found', len(video_chunks['Contents']))
+        self.logger.debug('This operation will process %d video_transcription job', max_no_jobs)
 
         for index, video_chunk in enumerate(video_chunks['Contents']):
             try:
@@ -40,7 +41,8 @@ class TranscriptionProcessor(Processor):
                 video_key = video_chunk['Key']
                 transcription_destination_key = self.__get_transcription_destination_key(video_key)
                 if self.s3_exist_object(transcription_destination_key):
-                    self.logger.debug('The video %s has been already transcribed, this transcription will be skipped',
+                    self.logger.debug('The samples_generation %s has been already transcribed, '
+                                      'this video_transcription will be skipped',
                                       video_key)
                     continue
 
@@ -56,7 +58,7 @@ class TranscriptionProcessor(Processor):
                 self.s3_delete_object(transcription_source_key)
                 self._transcribe.delete_transcription_job(TranscriptionJobName=job_name)
 
-                self.logger.debug('Audio transcription for %s is available at %s', job_name, transcription_source_key)
+                self.logger.debug('Audio video_transcription for %s is available at %s', job_name, transcription_source_key)
             except Exception as e:
                 self.logger.error(e)
 
@@ -76,7 +78,7 @@ class TranscriptionProcessor(Processor):
             }
         )
 
-        self.logger.debug('A new audio transcription job was started: \tJob Name: %s \tMedia Format: %s \tMedia '
[…]

### 2021-10-06 src/main/learning/dataset/prepare/base_dataset_preparer.py
initial version that is able to train SwAV for videos

@@ -1,7 +1,7 @@
 import logging
 import tensorflow as tf
 
-from learning.common.common_config import IMAGENET_CONFIG, FRAMES_SEQ_CONFIG, DATASET_BATCH_SIZE
+from learning.common.common_config import IMAGENET_CONFIG, FRAMES_SEQ_CONFIG
 
 
 class BaseDatasetPreparer:
@@ -21,13 +21,11 @@ class BaseDatasetPreparer:
 
         self.rgb_feature_dim = self.frames_seq_img_width * self.frames_seq_img_height * self.frames_seq_no_channels
 
-        self.dataset_batch_size = DATASET_BATCH_SIZE
+    def prepare_train_dataset(self, batch_size):
+        return self._prepare(self.train_dataset_path, batch_size)
 
-    def prepare_train_dataset(self):
-        return self._prepare(self.train_dataset_path)
-
-    def prepare_test_dataset(self):
-        return self._prepare(self.test_dataset_path)
+    def prepare_test_dataset(self, batch_size):
+        return self._prepare(self.test_dataset_path, batch_size)
 
     def _transform_image(self, img, resize_shape=None):
         if resize_shape is None:
@@ -81,7 +79,7 @@ class BaseDatasetPreparer:
             transformed_images.extend([transformed_img])
         return tf.stack(transformed_images)  # shape (no_frames, flattened_dim)
 
-    def _prepare(self, get_dataset_path_func):
+    def _prepare(self, dataset_path, batch_size):
         raise NotImplementedError('_prepare method not implemented.')
 
     def _prepare_sample(self, feature, label):

### 2021-10-06 src/main/learning/dataset/prepare/legacy/combined_dataset_preparer.py

@@ -10,17 +10,19 @@ class CombinedDatasetPreparer(SerializedDatasetPreparer):
         super().__init__(train_dataset_path, test_dataset_path)
         self.logger = logging.getLogger(__name__)
 
-    def prepare_train_dataset(self):
-        dataset = super().prepare_train_dataset()
+    def prepare_train_dataset(self, batch_size):
+        # don't batch the resulting dataset
+        dataset = super().prepare_train_dataset(batch_size=None)
+
         # apply transformations
         dataset = dataset.map(self._prepare_sample2, num_parallel_calls=tf.data.experimental.AUTOTUNE)
-        dataset = dataset.padded_batch(self.dataset_batch_size,
+        dataset = dataset.padded_batch(batch_size,
                                        padded_shapes=([None, None, None], [None, None], [None]))
         return dataset
 
     # TODO: implement prepare_test_dataset
-    def prepare_test_dataset(self):
-        return self.prepare_train_dataset()
+    def prepare_test_dataset(self, batch_size):
+        return self.prepare_train_dataset(batch_size)
 
     def _prepare_sample(self, feature, sample):
         raise NotImplementedError('This method is not supported for the Combined dataset')

### 2021-10-06 src/main/learning/dataset/prepare/legacy/opticalflow_dataset_preparer.py

@@ -10,17 +10,18 @@ class OpticalflowDatasetPreparer(SerializedDatasetPreparer):
         super().__init__(train_dataset_path, test_dataset_path)
         self.logger = logging.getLogger(__name__)
 
-    def prepare_train_dataset(self):
-        dataset = super().prepare_train_dataset()
+    def prepare_train_dataset(self, batch_size):
+        # batch_size -1 -> don't batch the resulting dataset
+        dataset = super().prepare_train_dataset(batch_size=None)
 
         # apply image transformation and data augmentation
         dataset = dataset.map(self._prepare_sample, num_parallel_calls=tf.data.experimental.AUTOTUNE)
-        dataset = dataset.batch(self.dataset_batch_size)
+        dataset = dataset.batch(batch_size)
         return dataset
 
     # TODO: implement prepare_test_dataset
-    def prepare_test_dataset(self):
-        return self.prepare_train_dataset()
+    def prepare_test_dataset(self, batch_size):
+        return self.prepare_train_dataset(batch_size)
 
     def _prepare_sample(self, opticalflow_sample, label):
         transformed_img = self._transform_image(opticalflow_sample)

### 2021-10-06 src/main/learning/dataset/prepare/legacy/rgb_dataset_preparer.py

@@ -10,16 +10,16 @@ class RGBDatasetPreparer(SerializedDatasetPreparer):
         super().__init__(train_dataset_path, test_dataset_path)
         self.logger = logging.getLogger(__name__)
 
-    def prepare_train_dataset(self):
-        dataset = super().prepare_train_dataset()
+    def prepare_train_dataset(self, batch_size):
+        dataset = super().prepare_train_dataset(batch_size=None)
         # apply image transformation and data augmentation
         dataset = dataset.map(self._prepare_sample, num_parallel_calls=tf.data.experimental.AUTOTUNE)
-        dataset = dataset.padded_batch(self.dataset_batch_size, padded_shapes=([None, None], [None]))
+        dataset = dataset.padded_batch(batch_size, padded_shapes=([None, None], [None]))
         return dataset
 
     # TODO: implement prepare_test_dataset
-    def prepare_test_dataset(self):
-        return self.prepare_train_dataset()
+    def prepare_test_dataset(self, batch_size):
+        return self.prepare_train_dataset(batch_size)
 
     def _prepare_sample(self, rgb_sample, label):
         # RNN feature

### 2021-10-06 src/main/learning/dataset/prepare/raw_dataset_preparer.py

@@ -8,14 +8,14 @@ class RawDatasetPreparer(BaseDatasetPreparer):
     def __init__(self, train_dataset_path, test_dataset_path):
         super().__init__(train_dataset_path, test_dataset_path)
 
-    def _prepare(self, dataset_path):
+    def _prepare(self, dataset_path, batch_size):
         raw_file_list = self._get_raw_file_list(dataset_path)
         gen_out_signature = tf.TensorSpec(shape=(None, None, None, 3), dtype=tf.int32)
         dataset: tf.data.Dataset = tf.data.Dataset.from_generator(self._data_generator,
                                                                   args=[raw_file_list],
                                                                   output_signature=gen_out_signature)
         dataset = dataset.map(self._prepare_sample3, num_parallel_calls=tf.data.AUTOTUNE, deterministic=False)
-        dataset = dataset.batch(self.dataset_batch_size)
+        # dataset = dataset.batch(batch_size)
 
         return dataset
 

### 2021-10-06 src/main/learning/dataset/prepare/serialized_dataset_preparer.py

@@ -6,10 +6,12 @@ class SerializedDatasetPreparer(BaseDatasetPreparer):
     def __init__(self, train_dataset_path, test_dataset_path):
         super().__init__(train_dataset_path, test_dataset_path)
 
-    def _prepare(self, get_dataset_path_func):
-        dataset_path = get_dataset_path_func()
+    def _prepare(self, dataset_path, batch_size):
         dataset_reader = TFRecordDatasetReader(self._get_dataset_type(), dataset_path)
         dataset = dataset_reader.read()
+        if batch_size:
+            dataset = dataset.batch(batch_size)
+
         return dataset
 
     def _prepare_sample(self, feature, label):

### 2021-10-06 src/main/learning/dataset/prepare/swav/swav_video_dataset_preparer.py

@@ -24,8 +24,8 @@ class SwAVDatasetPreparer(RawDatasetPreparer):
 
         self.object_detection_model = object_detection_model
 
-    def _prepare(self, dataset_path):
-        return super()._prepare(dataset_path)
+    def _prepare(self, dataset_path, batch_size):
+        return super()._prepare(dataset_path, batch_size)
 
     def _data_generator(self, list_video_path):
         extractor = RGBSamplesExtractor()

### 2021-10-06 src/main/learning/execution/base_model_executor.py

@@ -32,8 +32,8 @@ class BaseModelExecutor:
 
         self.model1.compile(optimizer=optimizer, loss=loss, metrics=metrics)
 
-    def train_model(self, no_epochs, no_steps_per_epoch=None):
-        dataset = self._get_train_dataset()
+    def train_model(self, batch_size, no_epochs, no_steps_per_epoch=None):
+        dataset = self._get_train_dataset(batch_size)
         history = self.model1.fit(dataset, epochs=no_epochs, steps_per_epoch=no_steps_per_epoch, verbose=2)
 
         # save training history in fs for further review
@@ -44,12 +44,12 @@ class BaseModelExecutor:
         model_serialization_path = self._get_model_serialization_path()
         self.model1.save(model_serialization_path)
 
-    def evaluate_model(self):
-        dataset = self._get_test_dataset()
+    def evaluate_model(self, batch_size):
+        dataset = self._get_test_dataset(batch_size)
         return self.model1.evaluate(dataset)
 
-    def predict_with_model(self):
-        dataset = self._get_test_dataset()
+    def predict_with_model(self, batch_size):
+        dataset = self._get_test_dataset(batch_size)
         return self.model1.predict(dataset)
 
     def _get_optimizer(self):
@@ -69,12 +69,12 @@ class BaseModelExecutor:
     def _get_model_history_serialization_path(self):
         return self.model_utility.get_model_history_serialization_path(self._get_model_type())
 
-    def _get_train_dataset(self):
-        dataset = self.dataset_preparer.prepare_train_dataset()
+    def _get_train_dataset(self, batch_size):
+        dataset = self.dataset_preparer.prepare_train_dataset(batch_size)
         return dataset
 
-    def _get_test_dataset(self):
-        dataset = self.dataset_preparer.prepare_test_dataset()
+    def _get_test_dataset(self, batch_size):
+        dataset = self.dataset_preparer.prepare_test_dataset(batch_size)
         return dataset
 
     def _get_model_type(self):

### 2021-10-06 src/main/learning/execution/legacy/nsdm_executor.py

@@ -14,14 +14,14 @@ class NSDMExecutor(BaseModelExecutor):
 
         self.dataset_preparer = CombinedDatasetPreparer(train_dataset_path, test_dataset_path)
 
-    def _get_train_dataset(self):
-        dataset = self.dataset_preparer.prepare_train_dataset()
+    def _get_train_dataset(self, batch_size):
+        dataset = self.dataset_preparer.prepare_train_dataset(batch_size)
         dataset = dataset.map(self.__map_combined_sample,
                               num_parallel_calls=tf.data.experimental.AUTOTUNE)
         return dataset
 
-    def _get_test_dataset(self):
-        dataset = self.dataset_preparer.prepare_test_dataset()
+    def _get_test_dataset(self, batch_size):
+        dataset = self.dataset_preparer.prepare_test_dataset(batch_size)
         dataset = dataset.map(self.__map_combined_sample,
                               num_parallel_calls=tf.data.experimental.AUTOTUNE)
         return dataset

### 2021-10-06 src/main/learning/execution/swav/swav_executor.py

@@ -24,14 +24,14 @@ class SwAVExecutor(BaseModelExecutor):
         self.crops_for_assign = [0, 1]
         self.temperature = 0.1
 
-    def train_model(self, no_epochs, no_steps_per_epoch=None):
+    def train_model(self, batch_size, no_epochs, no_steps_per_epoch=None):
         step_wise_loss = []
         epoch_wise_loss = []
 
         feature_backbone_model = self.__get_feature_backbone_model()
         prototype_projection_model = self.__get_prototype_projection_model()
 
-        dataset = self._get_train_dataset()
+        dataset = self._get_train_dataset(batch_size)
 
         for epoch in range(no_epochs):
             w = prototype_projection_model.get_layer('prototype').get_weights()
@@ -56,10 +56,14 @@ class SwAVExecutor(BaseModelExecutor):
 
     def train_step(self, input_views, feature_backbone, projection_prototype, optimizer, crops_for_assign, temperature):
         clip1, clip2, clip3, clip4, clip5 = input_views
-        inputs = [clip1, clip2, clip3, clip4, clip5]
+        inputs = [tf.expand_dims(clip1, axis=0),
+                  tf.expand_dims(clip2, axis=0),
+                  tf.expand_dims(clip3, axis=0),
+                  tf.expand_dims(clip4, axis=0),
+                  tf.expand_dims(clip5, axis=0)]
         batch_size = inputs[0].shape[0]
 
-        crop_sizes = [inp.shape[1] for inp in inputs]  # list of crop size of views
+        crop_sizes = [inp.shape[2] for inp in inputs]  # list of crop size of views
         unique_consecutive_count = [len([elem for elem in g]) for _, g in groupby(crop_sizes)]
         idx_crops = tf.cumsum(unique_consecutive_count)
 

### 2021-10-06 src/main/learning/model/swav/swav_builder.py

@@ -4,7 +4,8 @@ from learning.model.base_model_builder import BaseModelBuilder
 from tensorflow.keras.applications.resnet50 import ResNet50
 from tensorflow.keras import Input
 from tensorflow.keras.models import Model
-from tensorflow.keras.layers import Dense, GlobalAveragePooling2D, BatchNormalization, Activation
+from tensorflow.keras.layers import Dense, GlobalAveragePooling2D, BatchNormalization, Activation, Masking, \
+    Bidirectional, LSTM, TimeDistributed
 from learning.common.model_utility import SWAV
 
 
@@ -15,36 +16,35 @@ class SwAVModelBuilder(BaseModelBuilder):
         self.no_projection_1_neurons = 1024
         self.no_projection_2_neurons = 96
         self.prototype_vector_dim = 15
+        self.num_channels = 3
+        self.lstm_cells = 2048
+        self.embedding_size = 4096
 
-    # Features detection (Embeddings)
+    # Features detection (Embeddings) / CNN-LSTM model
     def build(self, **kwargs) -> Model:
-        input_shape = (None, None, 3)
-
-        backbone_model: Model = ResNet50(include_top=False, weights=None, input_shape=(None, None, 3))
+        backbone_model: Model = ResNet50(include_top=False, weights=None, input_shape=(None, None, self.num_channels))
         backbone_model.trainable = True
 
-        inputs = Input(shape=input_shape)
-        x = backbone_model(inputs, training=True)
+        features_input_shape = (None, None, self.num_channels)
+        features_inputs = Input(shape=features_input_shape)
+        features_x = backbone_model(features_inputs, training=True)
+        # TODO: evaluate if using Flatten instead of GlobalAveragePooling2D
+        features_x = GlobalAveragePooling2D()(features_x)
+        features_model = Model(features_inputs, features_x)
 
+        seq_input_shape = (None, None, None, self.num_channels)
+        seq_inputs = Input(shape=seq_input_shape)
+        seq_x = TimeDistributed(features_model)(seq_inputs)
+        seq_x = Masking()(seq_x)
+        seq_x = Bidirectional(LSTM(self.lstm_cells))(seq_x)
 
-        # # (no_steps, 224x224x3), no_steps is None because it will be determined at runtime by Keras
-        # input_shape = (None, self.feature_dim)
-        # inputs = Input(shape=input_shape, name='rgb_inputs')
[…]

### 2021-10-06 src/main/learning/train_model.py

@@ -23,6 +23,8 @@ from learning.common.model_utility import ModelUtility
 
 DEFAULT_NO_EPOCHS = 5
 DEFAULT_NO_STEPS_EPOCHS = None
+DEFAULT_NO_BATCH_SIZE = 1
+# DEFAULT_NO_BATCH_SIZE = 64
 
 
 def get_cmd_args():
@@ -32,6 +34,7 @@ def get_cmd_args():
     parser.add_argument('-tr', '--train_dataset_path', help='Train Dataset Path', required=True)
     parser.add_argument('-ne', '--no_epochs', help='Number of epochs', default=DEFAULT_NO_EPOCHS)
     parser.add_argument('-ns', '--no_steps', help='Number of steps per epoch', default=DEFAULT_NO_STEPS_EPOCHS)
+    parser.add_argument('-bs', '--batch_size', help='Dataset batch size', default=DEFAULT_NO_BATCH_SIZE)
     parser.add_argument('-odm', '--object_detection_model_name', help='Object Detection Model Name', required=False)
     parser.add_argument('-odcp', '--object_detection_checkout_prefix', help='Object Detection Checkout Prefix',
                         required=False)
@@ -99,7 +102,7 @@ def get_model(model_name):
     return model
 
 
-def get_executor(executor_name, model, train_dataset_path, **kwargs):
+def get_executor(executor_name, model, train_dataset_path, test_dataset_path, **kwargs):
     executors = {
         OPTICAL_FLOW: lambda: OpticalflowExecutor(model=model, train_dataset_path=train_dataset_path),
         RGB: lambda: RGBExecutor(model=model, train_dataset_path=train_dataset_path),
@@ -108,6 +111,7 @@ def get_executor(executor_name, model, train_dataset_path, **kwargs):
         SWAV: lambda: SwAVExecutor(feature_detection_model=model[0],
                                    projection_model=model[1],
                                    train_dataset_path=train_dataset_path,
+                                   test_dataset_path=test_dataset_path,
                                    object_detection_model=kwargs['ObjectDetectionModel'])
     }
     executor = executors[executor_name]()
@@ -125,6 +129,8 @@ def main():
     args = get_cmd_args()
     no_epochs = int(args.no_epochs)
     no_steps_per_epoch = args.no_steps
+    batch_size = int(args.batch_size)
+
     model_name = args.model
     train_dataset_path = args.train_dataset_path
     executor_name = model_name
@@ -138,8 +144,8 @@ def main():
     model_utility = ModelUtility()
[…]

### 2021-10-06 articles/2020/simple-method-to-detect-nouns-and-numbers/article-tesis.tex

--- 2020-07-06 article/article-tesis.tex
@@
 We produced a set of video annotations indicating signs appearances for a small set of nouns and numbers along with a \textit{labeled PSL dataset} (PSL dataset).
-A composite model obtained from the combination of a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade optical flow, and a RNN with LSTM cells trained with raw rgb frames extracted from the PSL dataset achieved 0.5994 AUC for the Precision-Recall curve on signs classification tasks, reporting state-of-art results over the PSL dataset.
+A model obtained after ensemble a 2D CNN trained with movement patterns extracted from the PSL dataset using Lucas Kanade Opticalflow, and a RNN with LSTM cells trained with raw RGB frames extracted from the PSL dataset reporting state-of-art results over the PSL dataset on signs classification tasks in terms of AUC, Precision and Recall.
 \\
@@
 \begin{figure*}[hbt!]
-\includegraphics[width=\linewidth]{images/opticalflow-model-architecture.png}
+\includegraphics[scale=0.8]{images/opticalflow-model-architecture.png}
 \caption{Opticalflow model architecture}
@@
 \subsection{Novel Signs Detection Model (nSDm)}\label{nsdm}
-We propose a novel model for signs detection that assemble the two neural networks architectures described in sections \ref{opticalflowmodel-architecture} and \ref{rgbrecurrentmodel-architecture} with the objective to learn visual features like edges, corners and ridges (CNN) and at the same time patterns learned from a time based series of inputs (RNN) to boost the performance on detecting PSL elements.
+We propose a novel model for signs detection that ensemble the two neural networks architectures described in sections \ref{opticalflowmodel-architecture} and \ref{rgbrecurrentmodel-architecture} with the objective to learn visual features like edges, corners and ridges (CNN) and at the same time patterns learned from a time based series of inputs (RNN) to boost the performance on detecting PSL elements.
 CNN network receives optical flow inputs and the RNN branch receives RGB frames extracted from the labeled PSL dataset described in \ref{videoannot}.
 
-We designed two neural network architectures for nSDm, both architectures use pre trained Opticalflow and RGB models as base models and applies different model assemble techniques on top of them.
+We designed two neural network architectures for nSDm, both architectures use pre trained Opticalflow and RGB models as base models and applies different model ensemble techniques on top of them.
 This architectures are described in detail in sections \ref{nsdmv1-architecture} and \ref{nsdmv2-architecture}.
@@
 \subsubsection{nSDmV1 Model Architecture}\label{nsdmv1-architecture}
-Pre-trained Opticalflow and RGB recurrent models are assembled using transfer learning with all layers freeze along with a flexible data input pipeline for data feeding, transformation and normalization.
+Pre-trained Opticalflow and RGB recurrent models are ensemble using transfer learning with all layers freeze along with a flexible data input pipeline for data feeding, transformation and normalization.
 
@@
 \subsection{Sign detection testing}\label{new-videos-testing}
-We selected videos in \ref{datasetdesc} that were not used for training nSDm models for testing nSDm models (models were not trained with the 24 consultant videos).
+We selected videos in section \ref{datasetdesc} that were not used for training nSDm models for testing nSDm models (models were not trained with the 24 consultant videos).
 We proceed the input video to extract RGB frames for video fragments of 0.5 seconds length moving ahead 0.1 seconds on each loop until the end of the video that way we can reduce the number of cut or incomplete signs between video fragments.
@@
 
-We selected nSDmV2 for signs detection on new videos because we obtained better results with nSDmV2 than with nSDmV1. Experimentation results are described on \ref{experimentation}
+We selected nSDmV2 for signs detection on new videos because we obtained better results with nSDmV2 than with nSDmV1. Experimentation results are described on section \ref{experimentation}
 
@@
 \begin{figure}[hbt!]
-\includegraphics[width=\linewidth]{images/dos-opticalflow.jpg}
+\centering
+\includegraphics[scale=0.3]{images/dos-opticalflow.jpg}
 \caption{PSL number "Two" OpticalFlow representation}
@@
 % Opticalflow results 5% of the dataset
-\begin{table*}[!htb]
+\begin{table*}
 \captionsetup{font=footnotesize}
@@
 % RGB Recurrent model results 5% of the dataset
-\begin{table*}[!htb]
+\begin{table*}
 \captionsetup{font=footnotesize}
@@
 % nSDmV1 model results 5% of the dataset
-\begin{table*}[!htb]
+\begin{table*}
 \captionsetup{font=footnotesize}
@@
 % nSDmV2 model results 5% of the dataset
-\begin{table*}[!htb]
+\begin{table*}
 \captionsetup{font=footnotesize}
@@
 
-Assembled models (nSDmV1 and nSDmV2) show the highest performance where nSDmV1 reports a loss equals to 0.627 a AUC equals to 1.0000 a Precision equals to 1.0000 and a Recall equals to 0.7273 and where nSDmV2 reports a loss equals to 0.1651 a Precision equals to 1.0000 and Recall equals to 1.000 and a AUC equals to 1.0000.
+Ensemble models (nSDmV1 and nSDmV2) show the highest performance where nSDmV1 reports a loss equals to 0.627 a AUC equals to 1.0000 a Precision equals to 1.0000 and a Recall equals to 0.7273 and where nSDmV2 reports a loss equals to 0.1651 a Precision equals to 1.0000 and Recall equals to 1.000 and a AUC equals to 1.0000.
 
-The results indicate assembled models perform better than individual models justifying the effort to design models that combine 2D CNN and RNN architectures.
+The results indicate ensemble models perform better than individual models justifying the effort to design models that combine 2D CNN and RNN architectures.
 nSDmV1 combines the two individual models with a simple outputs average showing showing a better recall than the Opticalflow and RGB recurrent models.
@@
 
-\textbf{PSL samples inference results} are presented on Figure \ref{fig:rgb-pareja-sample} we observed the noun \textit{Pareja} being correctly classified, in the other hand we observe that silent periods are being incorrectly classified as shown on Figure \ref{fig:rgb-pareja-wrong-sample}, adding one additional class for \textit{silent} could improve our classifier allowing it to infer when a consultant is performing any action.
+The nSDmV2 model was used for inference using the proposed Signs detection testing architecture described in Figure \ref{fig:video-testing-architecture} due to the performance presented.
+We successfully inferred samples as shown in  Figure \ref{fig:rgb-pareja-sample} where the sign for the noun "Pareja" was detected.
+In the other hand we observe that silent periods are being incorrectly classified as shown on Figure \ref{fig:rgb-pareja-wrong-sample}, adding one additional class for \textit{silent} could improve our classifier allowing it to infer when a consultant is performing any action.
 
@@
 \section{Discussion and Future Work}\label{discussion-future-work}
-\begin{enumerate}[•]
-\item The ability to detect all participants in recording sessions individually is mandatory for a successful video annotation pipelines based on video transcriptions as the one we presented at \ref{videoannot}.
-Future work about annotating videos based on patterns recognized in video content itself using non supervised learning approaches may improve the generated dataset driving to better results.
-\item Existing video annotation pipelines require large amounts of manual annotation work, generally for cleaning and validation tasks.
-Even though several steps within the overall annotation process can be fully automated human intervention for ensuring the quality of the generated dataset.
-Future work that use Reinforcement Learning for training an agent trained to clean and validate non-annotated videos before and after they are passed through the annotation pipeline may minimize or remove the need of human intervention during video annotation tasks.
-\item Not all the corporal movements executed by a PSL consultant are valid PSL signs or elements.
-Not handling these exceptions correctly could affect greatly the nSDm performance.
-Future work that add support for an additional class called \textit{invalid} and labeling invalid signs or corporal movements could help nSDm to learn invalid PSL actions and obtain better results.
-\item Optical flow represents movement patterns between video frames allowing models to learn these given movement patterns.
-However, Optical flow could consider camera unintentional movements and camera focus switches incorrectly as valid movement patterns between frames.
-Future work that adds Motion Boundary Histogram (MBH) \cite{wang2013dense} support to the optical flow gradient could add robustness to the ability to detect camera movements driving to a better optical flow samples generation.
-\item nSDm internal models (CNN and RNN models) as described on \ref{nsdm-architecture} can be trained independently using different hyper parameters, optimizers and loss functions.
-Pre-trained models will boost the nSDm training process enabling future works to use PSL to train nSDm using more epochs, data and being more flexible on using single models (CNN or RNN) or combining them as in nSDm.
-\item Improving they way people with deafness or hearing disabilities interact with computers requires delivering an output after PSL signs are detected.
-Future work that use Auto Encoder or Adversarial Generative Networks can be used to generate output in the form of signs that are recognizable completing the cycle of a bidirectional communication.
-\item Training nSDm with more PSL samples greatly improves its performance. Future work on annotating more videos including PSL content will increase nSDm performance.
 
-\end{enumerate}
+We processed the five percent of the PSL dataset with the proposed video annotation pipeline producing PSL samples for nouns and numbers using the Lucas Kanade Opticalflow representation and sequential RGB frames respectively.
+We trained four models described on sections \ref{opticalflowmodel}, \ref{rgbrecurrentmodel} and \ref{nsdm} obtaining results presented on section \ref{experimentation}.
+Results shown over fitting due to number of samples used to train the models.
+As a continuation of this work we will continue processing the rest of the PSL dataset and train models to improve their robustness.
+
+A successful supervised learning task requires a labeled dataset where samples are carefully produced and annotated.
+The video annotation pipeline described on section \ref{videoannot} requires a significant amount of human intervention to find video segments where signs are followed by a translation delivered after a delay factor that varies between translations.
+In this work we have estimated a delay factor of 3 seconds to ensure extracted frames contain the target sign but a the same time it introduces additional frames requiring human intervention to remove frames that are not relevant to the target sign.
+Applying self supervised learning techniques to avoid or minimize the need for human intervention while labeling the PSL dataset and other external PSL datasets available like the "Aprendo en Casa" dataset (Gisella Bejarano et al.) is the natural next step for this work where pre-trained nSDm models enriched with an auto-encoder architecture can be used to remove the need to human intervention on the proposed video annotation pipeline.
+
+State of art on pose estimation and body expressions detection are based on key points, joints and heat maps regression.
+The method described in this work is a supervised learning task for signs classification, converting a classification problem into a regression one seems to be a good option that could be beneficial.
+Movement across frames is captured with Opticalflow showing the body parts a PSL consultant moved to emit a sign.
+We are looking for a method for calculating key points and joint coordinates from Opticalflow samples, caluculated key points and joint coordinates which are inputs for a 2D CNN (dowsampling) and 2D Transposed CNN (upsampling) for heat maps regression that will finally be used to detect PSL elements.
 
 \section{Conclusion}\label{conclusion}
-\begin{enumerate}[•]
-\item Video annotation pipelines based on video transcriptions require recognizing all video participants for a correct video annotation.
-\item The \textit{Delay Factor} introduced noise to the generated labeled PSL dataset. Post annotation human intervention is needed to reduce the delay factor effect.
-\item Annotating more PSL videos and training nSDm with these new generated samples increases the model performance.
-\item Optical flow successfully represents PSL consultant corporal movements over time.
-\item Transfering weights learned a ResNet152 model trained with ImageNet influence positively Optical flow patterns learning.
-\item Using optical flow for movement pattern representation allow capturing all actions and movements executed by a PSL consultant.
-\item RGB video frames extracted from video fragments allow networks to learn sequences of movement over time.
-\item Using LSTM cells allow successfully usage of larger PSL sequences (a larger number of frames) due to its better handling of the vanishing gradients.
-\item Combining a CNN and a RNN in a composite model improves signs detection performance, while CNNs learns visual features in optical flow samples RNNs learns sequences of actions, networks outputs can be pondered for training a better classifier.
-\item The area under the precision-recall curve allow measuring how well is nSDm detecting because it summarizes the trade-off between the true positive signs rate and the predicted signs.
-\end{enumerate}
+Human intervention was required for cleaning and pre-processing input videos before they can passed to the proposed video annotation pipeline.
+It positively affected produced samples quality because video segments containing noise and non relevant frames can be easily removed in advance.
+A \textit{Delay Factor} between signs emitting and signs translation introduces noise because it varies on each produced sample requiring additional human intervention to post-process produced samples to remove non relevant frames.
+
+Lucas Kanade Opticalflow feature tracking method successfully represented movement that occurred during signs emitting, it is important to note that when a sign is emitted many body parts are moved including arms, hands, head, neck and eyes, Opticalflow is capable to capture movement patterns for the entire body configuring an excellent tool for visual features representation in PSL elements.
+It is a very CPU inexpensive algorithm that can be applied as a data augmentation/transformation in data input pipelines for both training and test allowing to expand its utilization to a wide range of datasets.
+
+Opticalflow model shown better performance than the RGB recurrent model in terms of AUC, Precision and Recall, the Opticalflow model uses a pre-trained RestNet152 base model with transfer learning (freezing) indicating that using a pre-trained base model positively affect the model performance.
+RGB recurrent model performance is subject to improve as we train with more PSL samples.
+
+Ensemble models shown better performance than Opticalflow and RGB recurrent models where nSDmV2 shown the highest performance.
+The nSDmV2 novel architecture where pre-trained base models were popped and then concatenated allowing adding additional layers for learning features after based model were concatenated and subsequent classifier.
+
+The area under the precision-recall curve allow measuring how well is nSDm detecting because it summarizes the trade-off between the true positive signs rate and the predicted signs.
 

### 2021-10-06 phd/thesis plan/MainText/chapter1.tex

@@
 Schmidhuber proposed an early self supervised method \cite{schmidhuber1990making} where two fully self supervised recurrent neural networks predicted future reinforcement based on external interactions and state sent to the next network cell is calculated on differences on successive predictions achieving a system capable to generate supervision signals.
-Unlike supervised tasks where predictions are influenced by target and observed variables, self supervised tasks are influenced in a greater extent by hidden or invisible variables also known as latent variables
+Unlike supervised tasks where predictions are influenced by target and observed variables, self supervised tasks are influenced in a greater extent by hidden or invisible variables also known as latent variables.
+By capturing those dependencies, a model can be used to answer questions about the values of unknown variables given the values of known variables.
+Energy-Based Models (EBMs) capture dependencies by associating a scalar energy (a measure of compatibility) to each configuration of the variables.
+Inference, i.e., making a prediction or decision, consists in setting the value of observed variables and finding values of the remaining variables that minimize the energy.
+Learning consists in finding an energy function that associates low energies to correct values of the remaining variables, and higher energies to incorrect values.
+A loss functional, minimized during learning, is used to measure the quality of the available energy functions.
+Within this common inference/learning framework, the wide choice of energy functions and loss functionals allows for the design of many types of statistical models, both probabilistic and non-probabilistic \cite{lecun2006tutorial}.
 
-\subsubsection{Pretext Tasks}
-\subsubsection{Self Supervised Representation Learning}
+Unsupervised visual representation learning, or self-supervised learning, aims at obtaining features without using manual annotations and is rapidly closing the performance gap with supervised pre- training in computer vision.
+Many recent state-of-the-art methods build upon the instance discrimination task that considers each image of the dataset (or “instance”) and its transformations as a separate class.
+This task yields representations that are able to discriminate between different images, while achieving some invariance (latent variables) to image transformations.
+Recent self-supervised methods that use instance discrimination rely on a combination of two elements: a contrastive loss and a set of image transformations.
+The contrastive loss removes the notion of instance classes by directly comparing image features while the image transformations define the invariances encoded in the features.
+Both elements are essential to the quality of the resulting networks \cite{caron2020unsupervised}.
+Caron et al 2020 proposed a cluster based (non constrastive or architectural) approach for learning latent variables or invariance from non annotated datasets that is extendable to video datasets called SwAV (Swapping Assignments between views).
+SwAV uses an online clustering mechanism that is capable to scale well to large datasets including video by solving an optimal transport problem using the Sinkhorn distance.
+
 \subsection{Signs Language Recognition and annotation}

### 2021-10-06 phd/thesis plan/Preamble/abstract.tex

\unnumberedchapter{Abstract} 
\chapter*{Abstract} 
\subsection*{\thesistitle}

People with temporal or permanent disabilities who aim to interact with computer based system faces a series of difficulties finding technology that fit their needs. Deafness is known to be the most prevalent disability in the Peruvian population who represents a minority with no equal access to technology and services.

Recent advances in user experience, usability standards and regulations supported the emergence of inclusive computer systems. However very little has been done to enable user experiences specialized for deaf. This doctoral thesis proposal describes a future effort that combines state of art self supervised learning techniques for unlabeled corpus annotation,  graphical gestural user interfaces to implement functional prototype user interface that enables a fluent communication between deaf and a computing system.

Self supervised learning techniques will be applied to the Peruvian signs language corpus developed by the grammar and signs research group of the Pontifical Catholic University of Peru contributing with future researches on the Signs Language recognition field. Human modeling animation will be applied to the self supervised annotation output for building the graphical gestural user interface.

We believe this work will tremendously contribute with the scientific community specially with the Peruvian community who is actively working on improving and enabling technology access to minorities.

### 2022-08-17 src/main/learning/execution/swav/swav_executor.py
additional updates to the swav executor

@@ -133,8 +133,7 @@ class SwAVExecutor(BaseModelExecutor):
                             crop_to_compare_prototype_start = per_replica_batch_size * crop_to_compare_index
                             crop_to_compare_prototype_end = per_replica_batch_size * (crop_to_compare_index + 1)
 
-                        crop_to_compare_prototype = tf.gather(prototype, tf.range(crop_to_compare_prototype_start,
-                                                                                  crop_to_compare_prototype_end))
+                        crop_to_compare_prototype = prototype[crop_to_compare_prototype_start:crop_to_compare_prototype_end]
                         crop_probability = tf.nn.softmax(crop_to_compare_prototype / self.temperature)
                         cross_entropy_loss = self.loss(crop_prototype_q_code, crop_probability)
 
@@ -149,6 +148,7 @@ class SwAVExecutor(BaseModelExecutor):
             # back propagation
             variables = self.feature_backbone_model.trainable_variables + \
                 self.prototype_projection_model.trainable_variables
+
             gradients = tape.gradient(loss, variables)
             self.optimizer.apply_gradients(zip(gradients, variables))
 

### 2022-08-17 src/main/learning/model/swav/swav_builder.py

@@ -8,12 +8,12 @@ class SwAVModelBuilder(BaseModelBuilder):
     def __init__(self):
         super().__init__()
 
-        self.no_projection_1_neurons = 1024
-        self.no_projection_2_neurons = 512
-        self.prototype_vector_dim = 128
+        self.no_projection_1_neurons = 256
+        self.no_projection_2_neurons = 128
+        self.prototype_vector_dim = 64
         self.num_channels = 3
-        self.lstm_cells = 512  # 2048
-        self.embedding_size = 1024  # 4096
+        self.lstm_cells = 256  # 2048
+        self.embedding_size = 512  # 4096
 
     # Features detection (Embeddings) / CNN-LSTM model
     def build(self, **kwargs) -> tf.keras.models.Model:

### 2022-08-17 src/main/learning/sage_maker_training.py

@@ -25,7 +25,7 @@ role = os.getenv('SAGEMAKER_ROLE')
 hyper_parameters = {
     'detect_person': False,
     'mirrored_training': True,
-    'batch_size': 4,
+    'batch_size': 3,
     'no_replicas': 4,
     'no_epochs': 20,
 }

### 2022-08-18 articles/2021/self-supervised-method-to-detect-nouns-and-numbers/article-tesis.tex

\documentclass[twocolumn,conference]{article}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{inputenc}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{enumerate}
\usepackage{caption}

\usepackage{array,booktabs,longtable,tabularx}
\usepackage{ltablex}
\usepackage{wrapfig,lipsum,booktabs}
\usepackage{multirow}

\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

\begin{document}
\author[1]{Eric Raphael Huiza Pereyra}
\affil[1]{Artificial Intelligence Research Group (IAPUCP) - Post Graduate Program - Pontificia Universidad Católica del Perú}
\affil[]{\textit{eric.huiza@pucp.edu.pe}}

\author[2]{Cesar Armando Beltrán Castañon}
\affil[2]{IA-PUCP}
\affil[]{\textit{cbeltran@pucp.edu.pe}}

\title{%
	\vspace{-2.0cm}
	\textbf{Talking with signs} \\	
	\Large \textbf{A self supervised method to detect nouns and numbers in a non-annotated signs language corpus}
}

\maketitle

\begin{abstract}
Self supervised learning is a current topic that allows transferring learned latent variables on pre-text tasks into downstream tasks achieving notable increases in terms of model performance. In this work we present a self supervised approach for learning latent variables in a non annotated Peruvian Signs Language dataset using a SwAV based approach applied to video datasets. We present a novel signs detection model (nSDmV3) which takes advantage of a pre-trained self supervised features detection model. obtaining XX\% AUC To our knowledge it is the first work that uses SwAV as a baseline to detect signs in a non-annotated dataset. 
\\
\keywords{self supervised learning, SwAV, latent variables,Peruvian signs language}
\end{abstract}
\section{Introduction}\label{Introduction}
Starting a model architecture design by considering the usage of pre-trained models is always a good starting point,  even when using a ResNet50 network backbone pre-trained with ImageNet the resulting model will leverage a set of fine tuned layers decreasing hours and data volume required for training. There have been some efforts to build an annotated Peruvian Signs Language (PSL) dataset (Bejarano et al) with promising but still small results. In this paper we are using Self Supervised techniques to overcome the current low annotated PSL dataset availability problems towards to obtain improved results of a previous work that was focused on Supervised learning and classic movement and pattern recognition techniques \cite{Pereyra_2021_CVPR}.

Signs languages can be described from spatial-temporal features extracted from isolated glosses and a large set of continuous streams of contextually related frames.  Similar to other languages, context is extremely important to define what is semantically correct and to provide meaning to a set of interconnected glosses which relations can easily become extremely complex and can only be described from an also extremely large set of spatial and temporal features but most importantly latent variables intrinsic to the data.

This work is focused on non contrastive clustered methods in order to decrease the computation needed to compare feature by feature required for constrastive methods. Our method is based on Swapping Assignments between multiple Views (SwAV) \cite{caron2020unsupervised} as a backbone and extended it to design a set of model architectures that can learn latent variables from a set of non annotated PSL videos, training our model with the PSL dataset generated by the department of grammar at PUCP.

Sign Language corpus annotation is a tedious and complex task where a set of automatic and manual tasks need to be orchestrated together in order to produce a quality corpus with statistical value. In this work we solve the problem of lack of annotated PSL datasets availability by proposing a self supervised method where a features detection model can be trained to learn latent variables while solving a set of pre-task in our case finding compatible augmented video fragments across the dataset in order to solve a downstream nouns and numbers classification task.

Finally we extended model architectures that were previously created as part of \cite{Pereyra_2021_CVPR} to use our SwAV based pre-trained model as a backbone and to improve the AUC of PR metric in these models. This work is divided on Related Work \ref{related-work},  Experimentation \ref{experimentation},  Discussion \ref{discussion} and Conclusion \ref{conclusion}

\section{Related Work}\label{related-work}
\subsection{Self Supervised Learning}
Predicting the future or reconstructing the past is a way to learn and understanding the hidden patterns and distributions in data which is not necessarily annotated or pre-processed. Schmidhuber proposed an early self supervised method \cite{schmidhuber1990making} where two fully self supervised recurrent neural networks predicted future reinforcement based on external interactions and state sent to the next network cell is calculated on differences on successive predictions achieving a system capable to generate supervision signals. Unlike supervised tasks where predictions are influenced by target and observed variables, self supervised tasks are influenced in a greater extent by hidden or invisible variables also known as latent variables. By capturing those dependencies, a model can be used to answer questions about the values of unknown variables given the values of known variables. Energy-Based Models (EBMs) capture dependencies by associating a scalar energy (a measure of compatibility) to each configuration of the variables. Inference, i.e., making a prediction or decision, consists in setting the value of observed variables and finding values of the remaining variables that minimize the energy. Learning consists in finding an energy function that associates low energies to correct values of the remaining variables, and higher energies to incorrect values. A loss functional, minimized during learning, is used to measure the quality of the available energy functions. Within this common inference/learning framework, the wide choice of energy functions and loss functionals allows for the design of many types of statistical models, both probabilistic and non-probabilistic \cite{lecun2006tutorial}.

Unsupervised visual representation learning, or self-supervised learning, aims at obtaining features without using manual annotations and is rapidly closing the performance gap with supervised pre- training in computer vision. Many recent state-of-the-art methods build upon the instance discrimination task that considers each image of the dataset (or “instance”) and its transformations as a separate class. This task yields representations that are able to discriminate between different images, while achieving some invariance (latent variables) to image transformations. Recent self-supervised methods that use instance discrimination rely on a combination of two elements: a contrastive loss and a set of image transformations. The contrastive loss removes the notion of instance classes by directly comparing image features while the image transformations define the invariances encoded in the features. Both elements are essential to the quality of the resulting networks \cite{caron2020unsupervised}. Caron et al 2020 proposed a cluster based (non constrastive or architectural) approach for learning latent variables or invariance from non annotated datasets that is extendable to video datasets called SwAV (Swapping Assignments between views). SwAV uses an online clustering mechanism that is capable to scale well to large datasets including video by solving an optimal transport problem using the Sinkhorn distance.

\subsection{Signs Language Recognition and annotation}
Signs Language Recognition (SLR) are roughly divided in two categories Isolated SLR and Continuous SLR\cite{adaloglou2020comprehensive}.

\subsubsection{Isolated Signs Language Recognition}
Methods using this approach will assume individual and well defined glosses that are recognized in a series of video frames. Recognizing gestures is a difficult task, due to intrapersonal and interpersonal variations in performing them. In \cite{kennaway2015avatar} a new descriptive language called SiGML (Signing Gesture Markup Language) is presented, SiGML was developed from HamNoSys, the Hamburg Notation System. HamNoSys is a notation for recording sign language gestures, developed by researchers on sign language at the IDGS (Institut für Deutsche Ge-bädensprache) at the University of Hamburg, it is not specific to any one sign language, but is intended to cover all signing gestures in all sign languages. It records sign languages elements in terms of hand shape, hand location and hand movement and other special terms for non manual gestures. Signs are not holistic gestures but are rather analyzable, as a combination of linguistically significant features. Similarly to spoken languages, Signs Languages are composed of the following indivisible features\cite{adaloglou2020comprehensive}:
\begin{itemize}
\item Manual features, i.e. hand shape, position, movement, orientation of the palm or fingers
\item Non-manual features, namely eye gaze, head-nods/ shakes, shoulder orientations, various kinds of facial expression as mouthing and mouth gestures.
\end{itemize}
SiGML provides a formal notation that is easily understood by computer because it is XML based.

\subsubsection{Continuous Signs Language Recognition}
Continuous Signs Language Recognition (CSLR) uses contextual information that is not limited to spatial information that resides around the hand shape for fixed point in time, but also temporal information that consists of hand and body movements. Signs are famously multi-channel, information is carried in the hand shape, motions, body pose and even facial gestures \cite{slimane2021context}, \cite{camgoz2017subunets}. While models that recognize text based language use punctuation marks to separate sentences, CSLR models use pauses e.g. silent regions. There have been studies in the literature addressing automatic sign segmentation. However to the best of the authors’ knowledge, there is no study which utilizes sign segmentation for realizing continuous sign language recognition for complex scenarios. Following successful segmentation, the system needs to under-stand what information is being conveyed within a sign sentence. Current approaches tackle this by recognizing sign glosses and other linguistic components. Such methods can be grouped under the banner of CSLR. From a computer vision perspective, this is the most challenging task. Considering the input of the system is high dimensional spatio-temporal data, i.e. sign videos, models are required that understand what a signer looks like and how they interact and move within their 3D signing space. Moreover, the model needs to comprehend what these aspects mean in combination. This complex modeling problem is exacerbated by the asynchronous multi-articulatory nature of sign languages. Although there have been promising results towards CSLR, the state-of-the-art can only recognize sign glosses and operate within a limited domain of discourse, namely weather forecasts \cite{camgoz2020sign}

\subsubsection{Sign Languages Recognition Approaches}
\textbf{Specialized sub-networks (SubUNets)} allow decomposing a complex signs recognition problem into a series of specialized expert systems allowing to learn both spatial and temporal features where each subunit learns an intermediate representation that is transferred to the next subunit. (Camgoz et al.) \cite{camgoz2017subunets} proposed an architecture where each SubUNet consists of three tiers, firstly Convolutional Neural Network layer to learn spatial features, then a Bidirectional Long Short Term Memory layer to learn temporal features on top of the spatial features and finally a Connectionist Temporal Classification Loos layer to allow the networks to be trained with different length videos and label sequences. (Huang et al.) \cite{huang2018videobased} proposed a two streams \textbf{3DCNN} architecture for learning global (entire video frame) and local (hands area crop) video representations in combination with a Hierarchical Attention Network (HAN) for latent space based recognition eliminating the need for temporal layers as proposed in other studies \cite{slimane2021context}, \cite{camgoz2017subunets}, \cite{camgoz2020sign}. HAN is an extension of LSTM which incorporates the attention mechanism based on the structure of input. It uses a latent space which maps video representations learned on the 3DCNN streams with relevant video sentences using one hot encoding vectors. The encoder in the HAN reflects the hierarchical structures in its inputs and incorporates the attention mechanism.
\section{Experimentation}\label{experimentation}
Our proposed self supervised signs detection approach is composed mainly for two well defined processes, a self supervised training architecture based on SwAV and a classifier that takes takes advantage of transfer learning from a fined tuned features detector from the first one.  
The dataset is a set of non annotated videos that were recorded in different sessions where twenty four signers were interviewed with the same or similar questionary. Videos in their raw format are processed through a data ingestion pipeline which converts videos in a single continuous video stream which is then passed through our self supervised architecture based in SwAV but adapted for videos using a SubUNet type of architecture. After training our self supervised system we obtained a features detector that learned latent variables from the dataset in a self supervised fashion. Finally used the features detector as a backbone model in our nSDmV3 classifier.
\subsection{Self Supervised Training}\label{self-supervised-training}
Our method uses a Multi Crop input data pipeline which receives a continuous video stream and produces augmented video fragments using a set of data augmentation transformations as described in \ref{multi-crop-data-pipeline}. We configured the input data pipeline to produce video fragments where frames were cropped in 224x224 and 96x96, in our experimentation each training step receives five video fragment transformations, two 224x224 and three 96x96.
Our self supervised training architecture uses a features detection model (nSDmV3 Features Detection Model) based on SubUNet where augmented video fragments are inputs of a RRN using LSTM cells where each frame is an input of a RestNet50 backbone to ensure learning spatial and temporal features and representing these features in a video embedding vector. These embedding vectors are then used as inputs of a Projections Model that follows a fully-connected architecture composed of two dense layers a batch normalization layer, and two dense layers. The entire self supervised training architecture is described on Figure \ref{fig:nsdmv3-self-supervised-training-model-architecture}

\begin{figure*}[hbt!]
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{images/nsdmv3-self-supervised-training-model-architecture.png}
\caption{nSDmV3 Self Supervised Training}
\label{fig:nsdmv3-self-supervised-training-model-architecture}
\end{figure*}

The projection model maps video embedding vectors to a set of clusters or prototype vectors which are then coded using the Sinkhron Algorithm to ensure a balanced vector codification. Codes and prototype vectors are then contrasted using cross entropy where one video video augmentation code is contrasted with all other video augmentation prototypes looking for compatibility towards to learn latent variables within the dataset. Finally the energy produced by each energy function in this case the cross contrasting using cross entropy is averaged with the objective to minimize the average energy across the training epochs.

\subsubsection{Multicrop Data Pipeline}\label{multi-crop-data-pipeline}
Our data input pipeline receives raw formatted videos and converts them in a continuous video stream using a \textbf{Video Stream Generator} component. Video steams are then sent to a \textbf{Data Preparer} component which is responsible to produce random duration video fragments between 0.5 and 1.5 seconds. A Human Detection Model is used to detect humans on video fragment frames getting rid of frame areas that will not contribute to learning PSL elements.

Video Fragments frames are randomly cropped using a 224x224 high resolution crops and 96x96 low resolution crops. Crops are then augmented in a sequence of transformations (Gaussian Blur, Color Jitter, Color Drop and Flip Left to Right).  Figure \ref{fig:nsdmv3-multicrop-data-pipeline-architecture} present architecture details.

\begin{figure*}[hbt!]
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{images/nsdmv3-self-supervised-training-data-input-architecture.png}
\caption{nSDmV3 Multicrop Data Pipeline Architecture}
\label{fig:nsdmv3-multicrop-data-pipeline-architecture}
\end{figure*}

\subsection{nSDmV3 Classifier}\label{nsdmv3-classifier}
We evolved a previous signs detection model architecture nSDmV2 as described on \cite{Pereyra_2021_CVPR} into a novel architecture that uses the Features Detection Model defined on \ref{fig:nsdmv3-self-supervised-training-model-architecture} as a backbone with all its layer fine tuned. The features detection backbone is concatenated with fully-connected layers using dense, batch normalization layers and finally a dense layer with a softmax activation. nSDmV3 classifier architecture is described in Figure \ref{fig:nsdmv3-classifier-architecture}

\begin{figure*}[hbt!]
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{images/nsdmv3-classifier-architecture.png}
\caption{nSDmV3 Classifier Architecture}
\label{fig:nsdmv3-classifier-architecture}
\end{figure*}

\subsection{Results}\label{results}
Our proposed self supervised architecture goal is to achieve successful metrics with a small number of annotated samples. The results presented on tables \ref{tab:nSDmV2-detection-results-5-percent} and \ref{tab:nSDmV3-detection-results-5-percent} indicate an increase on Precision, Recall and AUC on nSDmV3 compared to nSDmV2.

% nSDmV2 5% of the dataset
\begin{table*}
\captionsetup{font=footnotesize}
\centering
\begin{tabular}{ p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} }
\toprule
\multicolumn{5}{c}{\textbf{nSDmV2 Training Results with the 5\% of the dataset}} \\
\hline
\hline
\textbf{Epoch}&	\textbf{Loss}	&\textbf{Precision}	&\textbf{Recall}	&\textbf{AUC} \\
\hline
\midrule
Epoch 1&	2.7568&	0.0000e+00&	0.0000e+00&	0.0911\\
Epoch 2&	2.6647&	0.0000e+00&	0.0000e+00&	0.1008\\
Epoch 3&	2.6455&	0.0000e+00&	0.0000e+00&	0.1017\\
Epoch 4&	2.6296&	0.0000e+00&	0.0000e+00&	0.1193\\
Epoch 5&	2.6121&	0.0000e+00&	0.0000e+00&	0.1222\\
Epoch 6&	2.6032&	0.0000e+00&	0.0000e+00&	0.1222\\
Epoch 7&	2.5943&	0.0000e+00&	0.0000e+00&	0.1229\\
Epoch 8&	2.5875&	0.0000e+00&	0.0000e+00&	0.1229\\
Epoch 9&	2.5825&	0.0000e+00&	0.0000e+00&	0.1229\\
Epoch 10&	2.5790&	0.0000e+00&	0.0000e+00&	0.1229\\
\bottomrule
\end{tabular}
\caption{Shows results of training the nSDmV2 model with the 5\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{Loss} obtained loss (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall (5)\textit{AUC} area under the precision-recall curve.}
\label{tab:nSDmV2-detection-results-5-percent}
\end{table*}

% nSDmV3 5% of the dataset
\begin{table*}
\captionsetup{font=footnotesize}
\centering
\begin{tabular}{ p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} }
\toprule
\multicolumn{5}{c}{\textbf{nSDmV3 Test Results with the 5\% of the dataset}} \\
\hline
\hline
\textbf{Epoch}&	\textbf{Loss}	&\textbf{Precision}	&\textbf{Recall}	&\textbf{AUC} \\
\hline
\midrule
Epoch 1&	X.XXXX&	X.XXXX&	X.XXXX&	X.XXXX\\
Epoch 2&	X.XXXX&	X.XXXX&	X.XXXX&	X.XXXX\\
Epoch 3&	X.XXXX&	X.XXXX&	X.XXXX&	X.XXXX\\
Epoch 4&	X.XXXX&	X.XXXX&	X.XXXX&	X.XXXX\\
Epoch 5&	X.XXXX&	X.XXXX&	X.XXXX&	X.XXXX\\
Epoch 6&	X.XXXX&	X.XXXX&	X.XXXX&	X.XXXX\\
Epoch 7&	X.XXXX&	X.XXXX&	X.XXXX&	X.XXXX\\
Epoch 8&	X.XXXX&	X.XXXX&	X.XXXX&	X.XXXX\\
Epoch 9&	X.XXXX&	X.XXXX&	X.XXXX&	X.XXXX\\
Epoch 10&	X.XXXX&	X.XXXX&	X.XXXX&	X.XXXX\\
\bottomrule
\end{tabular}
\caption{Shows results of testing the nSDmV3 model with the 5\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{Loss} obtained loss (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall (5)\textit{AUC} area under the precision-recall curve.}
\label{tab:nSDmV3-detection-results-5-percent}
\end{table*}

\section{Discussion}\label{discussion}
It is possible to decrease and at some sort eliminate the need to annotate a dataset when pre-training a classifier features detection model using self-supervised techniques.  Our method approaches this problem using a SwAV based method towards to obtain a pre-trained features detection model trained with the PSL dataset leaning to a learn contextual and continuous patterns within the dataset. A classifier that use the pre-trained features detection model as a backbone is expected to have an boost on its performance in terms of AUC leaning to more simplistic and lighter video pre-processing and annotation tasks.

\section{Conclusion}\label{conclusion}
Self supervised techniques used to learn features from a non-annotated dataset stimulates and boosts positively the learning process by enabling the learning of latent variables in a non-annotated dataset. Pre-text tasks such us our SwAV based self supervised training pipeline is capable to extract and learn hidden features from the PSL dataset improving the AUC from our nSDmV3 classifier with respect to the previous version nSDmV2 in XX\%, obtaining a XX\% in comparison to the XX\% obtained with nSDmV2. The process to annotate a signs language dataset is tedious and complex requiring a great amount of manual annotation work along with a set of semi-automated pre-processing tasks. Self supervised techniques helps to decrease the amount of dataset annotation achieving successful results with low annotation levels. nSDmV3 along with our self supervised trainign pipeline are generalizable to other video related downstream tasks for the benefit of the scientific community.

\bibliographystyle{ieeetr}
\bibliography{References}

\end{document}

### 2023-06-18 src/main/learning/sage_maker_nsdmv3_training.py
improved RGB samples generation

@@ -26,7 +26,7 @@ hyper_parameters = {
     'mirrored_training': 'True',
     'batch_size': '4',
     'no_replicas': '4',
-    'no_epochs': '80',
+    'no_epochs': '60',
     'start_learning_rate': '0.0001',
     'no_dense_layer1_neurons': '256',
     'no_dense_layer2_neurons': '128',
@@ -44,12 +44,12 @@ estimator = Estimator(
     hyperparameters=hyper_parameters,
     metric_definitions=[
         {
-            'Name': 'train:EpochCrossEntropyLoss',
-            'Regex': 'epoch_loss: (.*?);',
+            'Name': 'train:EpochLoss',
+            'Regex': 'loss: \d+\.\d+',
         },
         {
-            'Name': 'train:StepCrossEntropyLoss',
-            'Regex': 'step_loss: (.*?);',
+            'Name': 'train:EpochAUC',
+            'Regex': 'auc: \d+\.\d+',
         }
     ],
     checkpoint_s3_uri=checkpoint_location,

### 2023-06-18 src/main/pre_processing/isolated_detection/samples_generation/rgb_samples_processor.py

@@ -1,52 +1,59 @@
 import os
+import csv
 import shutil
 import logging
-import csv
 
-from pre_processing.common.processor import Processor
 from pre_processing.common import io_utils
 from rgb_samples_handler import RGBSamplesHandler
+from pre_processing.common.processor import Processor
 
 
 class RGBSamplesProcessor(Processor):
-    __videos_dataset_rgb_path = 'dataset/rgb'
-
     def __init__(self):
         super().__init__()
 
         self.logger = logging.getLogger(__name__)
-        
+        self.handler = RGBSamplesHandler()
+
     def process(self, data):
         metadata_file_path = os.path.join(self.work_dir, data['metadata'])
+        dataset_path = data['dataset_path']
+
+        local_video_prefix = data['local_video_prefix']
+        s3_video_prefix = data['s3_video_prefix']
+
         if not os.path.exists(metadata_file_path):
             self.logger.debug('the %s metadata file does not exist, the process cannot continue', metadata_file_path)
             return
 
         reset_dataset = bool(data['reset_dataset'])
-        if reset_dataset and os.path.exists(os.path.join(self.work_dir, self.__videos_dataset_rgb_path)):
-            shutil.rmtree(os.path.join(self.work_dir, self.__videos_dataset_rgb_path))
+        if reset_dataset and os.path.exists(os.path.join(self.work_dir, dataset_path)):
+            shutil.rmtree(os.path.join(self.work_dir, dataset_path))
 
         delay_factor = float(data['translation_delay_factor'])
-        metadata_fields = ['token', 'video_path', 'start_time', 'end_time']
+        end_time_delay = float(data['end_time_bound_in_seconds'])
+        metadata_fields = ['token', 'video_name', 'start_time', 'end_time']
 
         with open(metadata_file_path) as metadata_file:
             reader = csv.DictReader(metadata_file, fieldnames=metadata_fields)
             next(reader, None)  # skip the csv header row
             for metadata in reader:
-                self.__handle_sample_metadata(metadata, delay_factor)
+                self.handle_sample_metadata(metadata, dataset_path, local_video_prefix, s3_video_prefix, delay_factor,
+                                            end_time_delay)
 
         logging.debug('process is completed')
 
-    def __handle_sample_metadata(self, metadata, delay_factor: float):
+    def handle_sample_metadata(self, metadata, dataset_path, local_video_prefix, s3_video_prefix, delay_factor: float,
+                               end_time_delay):
[…]

### 2023-06-18 src/main/pre_processing/isolated_detection/text_analysis/samples_metadata_generation_processor.py

@@ -17,6 +17,9 @@ class SamplesMetadataGenerationProcessor(Processor):
         self.override_local_files = False
         self.min_instances_per_sample = 10
 
+        self.transcription_path_prefix = 'transcriptions'
+        self.video_extension = '.mp4'
+
         self.logger = logging.getLogger(__name__)
 
     def process(self, data):
@@ -25,6 +28,8 @@ class SamplesMetadataGenerationProcessor(Processor):
 
         transcription_local_folder_path = os.path.join(self.work_dir, data['transcription']['localFolderPath'])
         transcription_s3_folder_path = data['transcription']['s3Path']
+
+        video_path_prefix = data['output']['video_path_prefix']
 
         if self.override_local_files:
             # download syntax detection files (e.g. nouns and numbers)
@@ -36,7 +41,7 @@ class SamplesMetadataGenerationProcessor(Processor):
         syntax_tokens = self.get_syntax_tokens(syntax_local_folder_path)
 
         # create the samples metadata file
-        metadata_file_columns = ['token', 'video_path', 'start_time', 'end_time']
+        metadata_file_columns = ['token', 'video_name', 'start_time', 'end_time']
         metadata_file_path = self.create_metadata_file(metadata_file_columns, syntax_local_folder_path)
         self.logger.debug('Samples Metadata File created at: %s', metadata_file_path)
 
@@ -50,7 +55,7 @@ class SamplesMetadataGenerationProcessor(Processor):
 
             transcription_file_relative_path = transcription_file_path.replace(self.work_dir + '/', '')
             self.handle_transcription_items(transcription_items, syntax_tokens, transcription_file_relative_path,
-                                            metadata_file_path, metadata_file_columns)
+                                            video_path_prefix, metadata_file_path, metadata_file_columns)
 
         self.upload_file(metadata_file_path, syntax_s3_folder_path + self.metadata_file_name)
         self.logger.debug('samples generation metadata uploaded to: %s', syntax_s3_folder_path + self.metadata_file_name)
@@ -93,7 +98,7 @@ class SamplesMetadataGenerationProcessor(Processor):
         return os.path.join(self.work_dir, metadata_file_path)
 
     def handle_transcription_items(self, transcription_items, syntax_tokens, transcription_file_path,
-                                   metadata_file_path, file_columns):
+                                   video_path_prefix, metadata_file_path, file_columns):
 
[…]

### 2023-06-18 articles/2021/self-supervised-method-to-detect-nouns-and-numbers/article-tesis.tex

@@
 \toprule
-\multicolumn{5}{c}{\textbf{nSDmV2 Training Results with the 5\% of the dataset}} \\
+\multicolumn{3}{c}{\textbf{nSDmV2 Training Results}} \\
 \hline
 \hline
-\textbf{Epoch}&	\textbf{Loss}	&\textbf{Precision}	&\textbf{Recall}	&\textbf{AUC} \\
+\textbf{Epoch}&	\textbf{Loss}		&\textbf{AUC} \\
 \hline
 \midrule
-Epoch 1&	2.7568&	0.0000e+00&	0.0000e+00&	0.0911\\
-Epoch 2&	2.6647&	0.0000e+00&	0.0000e+00&	0.1008\\
-Epoch 3&	2.6455&	0.0000e+00&	0.0000e+00&	0.1017\\
-Epoch 4&	2.6296&	0.0000e+00&	0.0000e+00&	0.1193\\
-Epoch 5&	2.6121&	0.0000e+00&	0.0000e+00&	0.1222\\
-Epoch 6&	2.6032&	0.0000e+00&	0.0000e+00&	0.1222\\
-Epoch 7&	2.5943&	0.0000e+00&	0.0000e+00&	0.1229\\
-Epoch 8&	2.5875&	0.0000e+00&	0.0000e+00&	0.1229\\
-Epoch 9&	2.5825&	0.0000e+00&	0.0000e+00&	0.1229\\
-Epoch 10&	2.5790&	0.0000e+00&	0.0000e+00&	0.1229\\
+Epoch 1&	2.7568&	0.0911\\
+Epoch 2&	2.6647&	0.1008\\
+Epoch 3&	2.6455&	0.1017\\
+Epoch 4&	2.6296&	0.1193\\
+Epoch 5&	2.6121&	0.1222\\
+Epoch 6&	2.6032&	0.1222\\
+Epoch 7&	2.5943&	0.1229\\
+Epoch 8&	2.5875&	0.1229\\
+Epoch 9&	2.5825&	0.1229\\
+Epoch 10&	2.5790&	0.1229\\
 \bottomrule
 \end{tabular}
-\caption{Shows results of training the nSDmV2 model with the 5\% of the labeled PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{Loss} obtained loss (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall (5)\textit{AUC} area under the precision-recall curve.}
+\caption{Shows results of training the nSDmV2 model with the labeled train PSL dataset: (1)\textit{Epoch} identifies the epoch in in the training process (2)\textit{Loss} obtained loss (3)\textit{Precision} obtained precision (4)\textit{Recall} obtained recall (5)\textit{AUC} area under the precision-recall curve.}
 \label{tab:nSDmV2-detection-results-5-percent}
@@
 \toprule
-\multicolumn{5}{c}{\textbf{nSDmV3 Test Results with the 5\% of the dataset}} \\
+\multicolumn{5}{c}{\textbf{nSDmV3 Train Results}} \\
 \hline

### 2023-06-20 articles/2021/self-supervised-method-to-detect-nouns-and-numbers/article-tesis.tex
made updates to the introduction

@@
 	\vspace{-2.0cm}
-	\textbf{Talking with signs} \\	
-	\Large \textbf{A self supervised method to detect nouns and numbers in a non-annotated signs language corpus}
+	\textbf{A self supervised method to detect nouns and numbers in a non-annotated signs language corpus} 
 }
@@
 \section{Introduction}\label{Introduction}
-Starting a model architecture design by considering the usage of pre-trained models is always a good starting point,  even when using a ResNet50 network backbone pre-trained with ImageNet the resulting model will leverage a set of fine tuned layers decreasing hours and data volume required for training.
-There have been some efforts to build an annotated Peruvian Signs Language (PSL) dataset (Bejarano et al) with promising but still small results.
-In this paper we are using Self Supervised techniques to overcome the current low annotated PSL dataset availability problems towards to obtain improved results of a previous work that was focused on Supervised learning and classic movement and pattern recognition techniques \cite{Pereyra_2021_CVPR}.
-
-Signs languages can be described from spatial-temporal features extracted from isolated glosses and a large set of continuous streams of contextually related frames.
-Similar to other languages, context is extremely important to define what is semantically correct and to provide meaning to a set of interconnected glosses which relations can easily become extremely complex and can only be described from an also extremely large set of spatial and temporal features but most importantly latent variables intrinsic to the data.
-
-This work is focused on non contrastive clustered methods in order to decrease the computation needed to compare feature by feature required for constrastive methods.
-Our method is based on Swapping Assignments between multiple Views (SwAV) \cite{caron2020unsupervised} as a backbone and extended it to design a set of model architectures that can learn latent variables from a set of non annotated PSL videos, training our model with the PSL dataset generated by the department of grammar at PUCP.
-
-Sign Language corpus annotation is a tedious and complex task where a set of automatic and manual tasks need to be orchestrated together in order to produce a quality corpus with statistical value.
-In this work we solve the problem of lack of annotated PSL datasets availability by proposing a self supervised method where a features detection model can be trained to learn latent variables while solving a set of pre-task in our case finding compatible augmented video fragments across the dataset in order to solve a downstream nouns and numbers classification task.
-
-Finally we extended model architectures that were previously created as part of \cite{Pereyra_2021_CVPR} to use our SwAV based pre-trained model as a backbone and to improve the AUC of PR metric in these models.
-This work is divided on Related Work \ref{related-work},  Experimentation \ref{experimentation},  Discussion \ref{discussion} and Conclusion \ref{conclusion}
+In recent years, there has been a significant surge in interest surrounding inclusive software.
+The aim is to enhance accessibility and improve user experience.
+Incorporating inclusivity is essential to ensure seamless interaction between individuals with temporal or permanent physical limitations and computer systems.
+Deafness is a prevalent physical limitation, according to the World Health Organization (WHO), over 466 million people worldwide experience some level of hearing loss \cite{deafness_and_hearing_loss_2020} and approximately 8\% of the Peruvian population is affected by hearing loss \cite{disabilities_survey_2012}.
+Despite that there have been great advances on using AI for sign language recognition the lack of large annotated dataset and the complexity associated to learn intrinsic relations and dependencies between continuous sign languages constructions have posed challenges \cite{adaloglou2020comprehensive}.
+These challenges  have transformed the problem into a high dimensional problem that cannot be solved solely with supervision \cite{facebook-ssl}.
+Starting a model architecture design that leverages pre-trained models is always a good starting point even when using a ResNet50 pre-trained with ImageNet backbone, the resulting model will leverage a set of fine tuned layers decreasing time and data volume required for training.
+Previous knowledge and existing skill-sets are vital elements that will indeed decrease the teaching required to learn a new task.
+Yann Lecun have used an analogy where he compares the time and skills humans require to learn to drive a car with the massive amount of time and a data required to train autonomous cars emphasizing that previous knowledge and skillsets will greatly reduce training time as well as the data require to train a model \cite{facebook-ssl}.
+In this work we use a non-annotated Peruvian Signs Language dataset created by the department of grammar at the PUCP, we will call it PSL dataset in the rest of this study.
+There have been some efforts to build an annotated Peruvian Signs Language (PSL) dataset (Bejarano et al) with promising results.
+However, there are challenges that are intrinsic to sign languages lexical and syntactical structures that require not only learning isolated glosses but continuous language constructions requiring to learn complex relationships and dependencies between actions that include several parts of the body including hands, arms, neck, head, eyes and eyebrows.
+Signs languages can be defined as a set of interrelated spatial-temporal elements composed from very large but still discrete set of isolated lexical glosses and most important from a continuous set of contextually related or interrelated streams of glosses that could be related to other streams of glosses recursively building complex continuous syntactical constructions.
+Continuous streams of glosses are high dimensional, which configures limitations when using probabilistic learning due to mathematical constraints while calculating probabilistic distributions using high dimensional data where integrals can become intractable\cite{kingma2022autoencoding}.
+In this paper we use the Sinkhorn distance which is an optimal transport distance we use to measure dissimilarity between continue streams of glosses probability distributions in an online and non-contrastive fashion, this to overcome the high dimensional curse by leveraging the geometry of the data.
+As in every language context and common sense are extremely important to define what is semantically correct and to provide a meaning to a set of interconnected streams of glosses.
+Learning latent variables intrinsic to the dataset enables the development of common sense in the context of PSL dataset.
+This, in turn, significantly reduces the size of the annotated PSL dataset required for accurate classification of PSL elements such as nouns and numbers.
+This paper employs self-supervision as a strategy to address the absence of a substantial dataset containing annotated PSL glosses and the nonexistence of a dataset comprising continuous streams of PSL language constructions during the training of our proposed novel signs language detection model nsvmv3.
+Annotating a Sign Language dataset is a laborious and intricate process that requires a combination of automatic and manual tasks to produce a high quality corpus with statistical value.
+While creating a large annotated PSL dataset is beyond the scope of this work, we propose a pretext task that leverage SwAV (Swapped assignment Views) \cite{caron2020unsupervised}.
+A self-supervised technique original designed for image data, we extended SwAV to work with the PSL dataset for training a spatial-temporal features model with the full PSL dataset.
+This approach allow us to learn latent variables from raw video files, enabling an understanding of the complex relations and dependencies between continuous PSL streams.
+This with the aim is to use our nsvmv3 model for a downstream task focused on successfully detecting and classifying signs utilizing small annotated PSL datasets for both training and inference purposes.
+To to ensure scalability when using our method with large PSL datasets, this work uses non-contrastive or online clustering methods to reduce computational load when comparing features and generating prototypes as discussed earlier.
+Specially, we utilize the Sinkhorn distance as a non-constrastive and online method to compare prototypes probability distributions \cite{caron2020unsupervised}.
+The structure of this work comprises the following sections: Related Work (Section \ref{related-work}), Experimentation (Section \ref{experimentation}), Discussion (Section \ref{discussion}), and Conclusion (Section \ref{conclusion}).
 
@@
 \subsection{Self Supervised Learning}
+Self Supervised Learning also called the dark matter of intelligence (Lecunn et all 2021) is a first step on human level AI since it is focused on learning a form of generalized common sense instead of learning limited features from a labeled datasets which leads to learn new skills without needing a great amount of teaching.
 Predicting the future or reconstructing the past is a way to learn and understanding the hidden patterns and distributions in data which is not necessarily annotated or pre-processed.

### 2023-07-03 articles/2021/self-supervised-method-to-detect-nouns-and-numbers/article-tesis.tex
updates to the paper introduction

@@
 Deafness is a prevalent physical limitation, according to the World Health Organization (WHO), over 466 million people worldwide experience some level of hearing loss \cite{deafness_and_hearing_loss_2020} and approximately 8\% of the Peruvian population is affected by hearing loss \cite{disabilities_survey_2012}.
-Despite that there have been great advances on using AI for sign language recognition the lack of large annotated dataset and the complexity associated to learn intrinsic relations and dependencies between continuous sign languages constructions have posed challenges \cite{adaloglou2020comprehensive}.
+Despite that there have been great advances on using AI for sign languages recognition, the absence of large annotated dataset and the complexity associated to learn intrinsic relations and dependencies between continuous sign languages constructions have posed challenges \cite{adaloglou2020comprehensive}.
 These challenges  have transformed the problem into a high dimensional problem that cannot be solved solely with supervision \cite{facebook-ssl}.
